### 2024-10-16 paper/main.tex
Back in it

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Seismica Submission Template
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Options available: 
%%					anonymous
%%					breakmath
%%					languages
%%					preprint
%%					report

%% anonymous: produces an anonymous PDF for double-blind review. Will NOT print authors information and acknowledgements, but WILL PRINT data availibility section, so be careful!
%% breakmath: for manuscripts with many long formulas, you can specify the breakmath option (loads the package breqn and uses the dmath environment)
%% languages: see below, use to add abstract(s) in additional languages
%% preprint: removes line numbers, switch to two-columns
%% report: if this is a report

% \documentclass[breakmath,report]{seismica}
\documentclass[preprint]{seismica}

% import packages
\usepackage{gensymb, nicefrac}

\title{Single-seismometer Focal Mechanism and Uncertainty Estimation from Body Waves}
\shorttitle{Single-seismometer Focal Mechanism and Uncertainty Estimation} % used for header, not mandatory

%% If this is a REPORT, select a report type within: 
%\reporttype{Null Results Report} % (null-results/failed experiments)
%\reporttype{Software Report}
%\reporttype{Data Report} %  (e.g., Large Community dataset initiatives, Instrument Deployments, and Field Campaigns)
%\reporttype{Fast Report}

%% Will not be printed if anonymous option ON
\author[1]{Victor Agaba
	% \orcid{0000-0000-0000-0000}
	\thanks{Corresponding author: victoragaba2025@u.northwestern.edu}
}
\author[1]{Suzan van der Lee
	\orcid{0000-0003-1884-1185}
}
\author[2]{Madelyn Sita
	\orcid{0000-0002-7214-7058}
}
\author[3]{Caio Ciardelli}
\author[4]{Paula Babirye}
\affil[1]{Department of Earth and Planetary Sciences, Northwestern University, Evanston, IL, USA}
\affil[2]{Department of Chemistry, University of Virginia, Charlottesville, VS, USA}

%% Author CRediT roles
%% Please use the CRediT roles as defined at https://casrai.org/credit
%% Use as many roles as necessary; there is no requirement to use all 14 roles
\credit{Conceptualization}{SL, MS.}
\credit{Methodology}{VA, SL.}
\credit{Software}{VA.}
%\credit{Validation}{people}
\credit{Formal Analysis}{VA, SL.}
%\credit{Investigation}{people}
%\credit{Resources}{people}
\credit{Writing - Original draft}{VA, SL.}
%\credit{Writing - Review \& Editing}{people}
\credit{Visualization}{VA, SL.}
\credit{Supervision}{SL.}
%\credit{Project administration}{people}
%\credit{Funding acquisition}{people}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Abstracts in other languages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If your article includes abstracts in other languages, uncomment the lines below and fill in
%%      the appropriate sections. You will need to use the [languages] option at the top,
%%      and will need to use lualatex instead of pdflatex to compile the document.
%% We will use luatex, polyglossia and fontspec for the compilation of the accepted version. 
%% Feel free to use any polyglossia command.
%\setotherlanguages{french,thai}  % replace with your language(s), per polyglossia
%% if an additional font is needed for the abstract, load it with:
%\newfontfamily\thaifont[Script=Thai]{Noto Serif Thai}

%% If you are using Arabic, do not include it in \setotherlanguages{} as it is not supported 
%% by polyglossia. Instead, with the [languages] option at the top, you can use these commands
%% within the text:
%%\begin{Arabic} and \end{Arabic} around paragraphs in Arabic
%%\n{} to wrap any digits within Arabic text that should read left-to-right
%%\textarabic{} for Arabic text embedded in a left-to-right paragraph

%% also see https://www.overleaf.com/latex/examples/how-to-write-multilingual-text-with-different-scripts-in-latex/wfdxqhcyyjxz for reference
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Several abstracts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% the command \makeseistitle does not allow page breaks in preprint mode. If you have
%%		many abstracts, you can use the command \addsummaries. It will induce a pagebreak.

\begin{document}
    \makeseistitle{
        \begin{summary}{Abstract}
            Work on later, 200-word max.
        \end{summary}
        \begin{summary}{Non-technical summary}
            Work on later, shorter than abstract.
        \end{summary}
    }

    %%%% In preprint mode, in case you have too many abstracts and need a page break, use this:
    %\addsummaries{
        %	\begin{french}
            %	\begin{summary}{Résumé} 
                %	The text goes here. Again, no longer than 200 words.
                %	\end{summary}
            %	\end{french}
        %	\begin{summary}{Non-technical summary}
            %	The text goes here. Again, no longer than 200 words.
            %	\end{summary}
        %}  %% don't forget this one!

\section{Introduction} \label{sec:introduction}

    [Suzan] Insert words here...\\
    Talk about this being an inverse method for a highly nonlinear problem\\
    Writing tip: maximize information content of words

\section{Cosine similarity misfit criterion} \label{sec:misfit}

    \subsection{Derivation}
     When a quake event happens, a single seismometer reads $P$, $SV$ and $SH$ body waves through which the respective amplitudes $A^P$, $A^{SV}$ and $A^{SH}$ along with their noise levels $\sigma^P$, $\sigma^{SV}$ and $\sigma^{SH}$ (representing standard deviation of an assumed normal distribution centered at the amplitude value) can be extracted.
     Given a ray path's azimuth $\phi$ and take-off angles $i$, $j$ of $P$ and $S$ waves respectively, we model the amplitudes as functions of the source mechanism's fault strike $\psi$, fault dip $\delta$ and slip rake $\lambda$, further accounting for the $P$- and $S$-velocities $\alpha_h$, $\beta_h$ at source depth $h$ \citep{AkiRichards1980}. The expressions are:

     \begin{align} \label{eq:eq1}
         \nonumber &A^P \sim \nicefrac{\left( s_R(3\cos^2(i) - 1) - q_R\sin(2i) - p_R\sin^2(i) \right)}{\alpha_h^3}\\
         &A^{SV} \sim \nicefrac{\left( 1.5 s_R\sin(2j) + q_R\cos(2j) + 0.5 p_R\sin(2j) \right)}{\beta_h^3}\\
         \nonumber &A^{SH} \sim \nicefrac{\left( q_L\cos(j) + p_L\sin(j) \right)}{\beta_h^3}
     \end{align}

     where

     \begin{align} \label{eq:eq2}
         \nonumber s_R &= 0.5 \sin(\lambda) \sin(2\delta)\\
         \nonumber q_R &= \sin(\lambda) \cos(2\delta) \sin(\psi_r) + \cos(\lambda) \cos(\delta) \cos(\psi_r)\\
         p_R &= \cos(\lambda) \sin(\delta) \sin(2\psi_r) - 0.5\sin(\lambda) \sin(2\delta) \cos(2\psi_r)\\
         \nonumber p_L &= 0.5\sin(\lambda) \sin(2\delta) \sin(2\psi_r) + \cos(\lambda) \sin(\delta) \cos(2\psi_r)\\
         \nonumber q_L &= -\cos(\lambda) \cos(\delta) \sin(\psi_r) + \sin(\lambda) \cos(2\delta) \cos(\psi_r)
     \end{align}

     Our goal is two-fold:
     \begin{enumerate}
         \item [i)]
            Find estimates $\widehat{\psi}, \widehat{\delta}, \widehat{\lambda}$ for the strike, dip and rake that \textit{best} describe the collected data.

        \item [ii)]
            Estimate the joint noise distribution of fault parameters to more accurately constrain the set of \textit{acceptable} estimates for the source mechanism.

     \end{enumerate}

     We thus need a metric against which to evaluate what the best and acceptable estimates are. The misfit function used by \citet{sita_potential_2022} is defined as the angle between the vector of observed amplitudes and that synthesized from parameter estimates. In this paper, we shall use a misfit $\mu$ of cosine similarity between the two vectors which has a 1-to-1 correspondence with angle:

     \begin{align} \label{eq:eq3}
         \mu(\mathbf{p}) = \frac{\mathbf{A}_s^\top\mathbf{A}_o}{\|\mathbf{A}_s\|\|\mathbf{A}_o\|}
     \end{align}

     where $\mathbf{A}_o = (A^P, A^{SV}, A^{SH})^\top$ are observed amplitudes, $\mathbf{A}_s = (\widehat{A}^P, \widehat{A}^{SV}, \widehat{A}^{SH})^\top$ are synthetic amplitudes, and $\mathbf{p} = (\widehat{\psi}, \widehat{\delta}, \widehat{\lambda})$ are parameter estimates.

     \subsection{Justification}
     We are interested in situations where absolute amplitudes are either unknown or unnecessary, for instance if using data collected from NASA's Insight mission \citep{sita_potential_2022}.
     In this case, we cannot know the absolute velocities because they depend on an understanding of Mars' interior that is good enough to build a reliable velocity model. Alternative misfit functions can account for the relative amplitudes, such as in Equations \ref{eq:eq4} and \ref{eq:eq5}:

     \begin{align} \label{eq:eq4}
         \mu(\mathbf{p}) = \sqrt{\left( \Delta \frac{A^{P}}{A^{SV}}\right)^2 + \left( \Delta \frac{A^{SV}}{A^{SH}}\right)^2}
     \end{align}

     \begin{align} \label{eq:eq5}
         \mu(\mathbf{p}) = \sqrt{\log^2 \left( \Delta \frac{A^P}{A^{SV}} \right) + \log^2 \left( \Delta \frac{A^{SV}}{A^{SH}} \right)}
     \end{align}

     First, we note that using cosine similarity is more numerically stable in cases where one of the observed or synthetic amplitudes is close to zero since the norm of the associated vector in Equation \ref{eq:eq3} may still be large enough to make division feasible.

     Secondly, we have chosen cosine similarity over angle because it involves one less operation, which is more computationally efficient and easier to differentiate for directed search algorithms (see Section \ref{sec:algorithms}). The Gaussian function in Equation \ref{eq:eq6} is similarly easy to differentiate, but it does not account for relative amplitudes so it would exclude otherwise acceptable solutions.

     \begin{align} \label{eq:eq6}
         &\mu(\mathbf{p}) \sim \exp \left( - \frac{1}{2} (\mathbf{A}_s - \mathbf{A}_o)^\top \mathbf{\Sigma}^{-1} (\mathbf{A}_s - \mathbf{A}_o) \right)\\
         \nonumber &\text{where } \mathbf{\Sigma} = \text{diag}(\sigma^P, \sigma^{SV}, \sigma^{SH})
     \end{align}

     \subsection{Tolerance}
     Given vectors $\mathbf{A}_s$ and $\mathbf{A}_o$, we would like to know the largest cosine similarity for which $\mathbf{A}_s$ is considered an acceptable fit for the recorded amplitudes while respecting the asymmetry of error levels for different amplitudes. This is visualized in 3D space as a confidence ellipsoid centered at $\mathbf{A}_o$ (Equation \ref{eq:eq7}), for which acceptable fits are vectors that fall inside the elliptical cone with vertex at the origin and tangent to the ellipsoid.
     \begin{align} \label{eq:eq7}
         \sum_{k \in \{P, SV, SH\}} \frac{(A^k_b-A^k)^2}{(\sigma^k)^2} = 1
     \end{align}

     where $\mathbf{A} = (A^P_b, A^{SV}_b, A^{SH}_b)^\top$ describes a boundary point. The elliptical cone intersects with the ellipsoid at every boundary point for which
     \begin{align} \label{eq:eq8}
         \sum_{k \in \{P, SV, SH\}} \frac{2A^k_b(A^k_b-A^k)}{(\sigma^k)^2} = 0
     \end{align}

     Combining equations \ref{eq:eq7} and \ref{eq:eq8} gives the plane containing every point of tangency:
     \begin{align} \label{eq:eq9}
         \sum_{k \in \{P, SV, SH\}} \frac{(A^k_b-A^k)^2}{(\sigma^k)^2} - \frac{2A^k_b(A^k_b-A^k)}{(\sigma^k)^2} = 1\\
         \nonumber \sum_{k \in \{P, SV, SH\}} \frac{A^k}{(\sigma^k)^2} A^k_b = \left( \sum_{k \in \{P, SV, SH\}} \frac{(A^k)^2}{(\sigma^k)^2} \right) -1
     \end{align}

     Let $\mathbf{N}_1 = (\nicefrac{A^P}{(\sigma^P)^2}, \nicefrac{A^{SV}}{(\sigma^{SV})^2}, \nicefrac{A^{SH}}{(\sigma^{SH})^2})^\top$ be the normal of this plane. Consider also the plane containing vectors $\mathbf{A}_o$ and $\mathbf{A}_s$, with normal
     \begin{align} \label{eq:eq10}
         \mathbf{N}_2 = \mathbf{A}_o \times \mathbf{A}_s = (,,)^\top
     \end{align}

\section{Search Methods} \label{sec:algorithms}

    Insert words here...

\section{Uncertainty Quantification} \label{sec:uncertainty}

    Insert words here...

\section{Example: from Paula} \label{sec:example}

    Insert words here...

\section{Conclusions} \label{sec:conclusion}

    Insert words here...

%% Will not be printed if anonymous option ON
\begin{acknowledgements}
    [Suzan] Insert words here...
\end{acknowledgements}

\section*{Data and code availability}
Zenodo, figshare, and Dryad to archive data and code. Citations for datasets and codes should be included in the references, including citations for any seismic networks from which data was used. Github is not considered a persistent repository, and we encourage authors to archive a snapshot of any github-hosted code on zenodo.

\section*{Competing interests}
    The authors have no competing interests.

%% If the article is accepted, a separate bibfile must be uploaded along with the compiled manuscript, source file, and separate figure files.
%% When available, DOI numbers must be provided for all references, including datasets and codes.
\bibliography{references}

\end{document}

### 2024-10-16 project/function_repo.py

@@ -1,0 +1,1273 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import scipy.stats as stats
+from obspy.taup import TauPyModel
+import collections
+
+from obspy.imaging.beachball import beachball
+from obspy.imaging.beachball import beach
+
+eps = 1e-10; halfpi = np.pi/2; twopi = 2*np.pi
+i_hat = np.array([1,0,0]); j_hat = np.array([0,1,0]); k_hat = np.array([0,0,1])
+np.random.seed(1029) # I hope this carries over to other files
+# It only does for randomness associated with fr
+
+def set_axes_equal(ax): # might be useful later
+    """
+    Make axes of 3D plot have equal scale so that spheres appear as spheres,
+    cubes as cubes, etc.
+
+    Input
+      ax: a matplotlib axis, e.g., as output from plt.gca().
+    """
+
+    x_limits = ax.get_xlim3d()
+    y_limits = ax.get_ylim3d()
+    z_limits = ax.get_zlim3d()
+
+    x_range = abs(x_limits[1] - x_limits[0])
+    x_middle = np.mean(x_limits)
+    y_range = abs(y_limits[1] - y_limits[0])
+    y_middle = np.mean(y_limits)
+    z_range = abs(z_limits[1] - z_limits[0])
+    z_middle = np.mean(z_limits)
+
+    # The plot bounding box is a sphere in the sense of the infinity
+    # norm, hence I call half the max range the plot radius.
+    plot_radius = 0.5*max([x_range, y_range, z_range])
+
+    ax.set_xlim3d([x_middle - plot_radius, x_middle + plot_radius])
+    ax.set_ylim3d([y_middle - plot_radius, y_middle + plot_radius])
+    ax.set_zlim3d([z_middle - plot_radius, z_middle + plot_radius])
+
+def uniform_samples(n: int, bounds: list) -> list:
+    """
+    Divide bounds into n uniformly distributed points
+
+    Args:
+        n (int): # divisions
+        bounds (list[float]): array bounds [start, end]
+        
+    Returns:
+        list: n-size array of floats
+    """
+    start = bounds[0]
+    end = bounds[1]
+    
+    return stats.uniform.rvs(loc=start, scale=end-start, size=n)
+
+def fill_square_randomly(n: int, bounds: list) -> list:
+    """
+    Fill 2D grid with n uniformly distributed points
+    Return array of coordinates w.r.t grid boundaries
[…]

### 2024-10-16 project/seismic_model.py

@@ -1,0 +1,202 @@
+'''
+Name: Victor Agaba
+
+Date: 5th December 2023
+
+The goal of this file is to create a model of the nonlinear inverse
+problem to which we will apply the optimization algorithms.
+'''
+
+
+from numpy import *
+import matplotlib.pyplot as plt
+
+
+class InvProblem:
+    '''
+    Base class for an unconstrained optimzation problem.
+    Catch anything whose method is not implemented.
+    '''
+    def __init__(self):
+        pass
+
+    def value(self, m):
+        raise Exception('Method "value" is not implemented')
+
+    def gradient(self, m):
+        raise Exception('Method "gradient" is not implemented')
+
+    def hessian(self, m):
+        raise Exception('Method "hessian" not is implemented')
+
+    def starting_point(self):
+        raise Exception('Method "starting_point" not implemented')
+
+
+class SeismicModel(InvProblem):
+    '''
+    This class will compute the misfit function and its gradient
+    for a given model.
+    '''
+    def __init__(self, phi, i, j, alpha, beta, d):
+        '''
+        Initialize with the parts of the model that don't change.
+        Uses the same notation as in the project description.
+        ALL ANGLES ARE IN RADIANS!!!
+        '''
+        # initialize misfits and grad norms for tracking
+        self.misfits = []
+        self.grad_norms = []
+        self.iterates = []
+        
+        # initialize constant Jacobian matrix
+        factors = array([alpha, beta, beta])**3
+        self.C = array([[(3*cos(i)**2 - 1), -sin(2*i), -sin(i)**2, 0, 0],
+                      [1.5*sin(2*j), cos(2*j), 0.5*sin(2*j), 0, 0],
+                      [0, 0, 0, cos(j), sin(j)]])
+        
+        # scale each row of C by the corresponding factor
+        self.C /= factors[:,newaxis]
+        
+        # also sture J and nabla_E for later use
+        self.J = zeros((5,3))
+        self.nabla_E = zeros(3)
+        
+        # store other inputs for later use
+        self.phi = phi
+        self.d = d
+        
+        # store current value of m, g and E
+        self.m = None
+        self.g = None
+        self.E = None
[…]

### 2024-10-16 seismic_model.py

@@ -57,7 +57,7 @@ class SeismicModel(InvProblem):
         # scale each row of C by the corresponding factor
         self.C /= factors[:,newaxis]
         
-        # also sture J and nabla_E for later use
+        # also store J and nabla_E for later use
         self.J = zeros((5,3))
         self.nabla_E = zeros(3)
         

### 2024-10-16 paper/main.tex
No TexLive on farallon

@@
     \subsection{Derivation}
-     When a quake event happens, a single seismometer reads $P$, $SV$ and $SH$ body waves through which the respective amplitudes $A^P$, $A^{SV}$ and $A^{SH}$ along with their noise levels $\sigma^P$, $\sigma^{SV}$ and $\sigma^{SH}$ (representing standard deviation of an assumed normal distribution centered at the amplitude value) can be extracted.
-     Given a ray path's azimuth $\phi$ and take-off angles $i$, $j$ of $P$ and $S$ waves respectively, we model the amplitudes as functions of the source mechanism's fault strike $\psi$, fault dip $\delta$ and slip rake $\lambda$, further accounting for the $P$- and $S$-velocities $\alpha_h$, $\beta_h$ at source depth $h$ \citep{AkiRichards1980}.
-The expressions are:
+     When a quake event happens, a single seismometer reads $P$, $SV$ and $SH$ body waves through
+     which the respective amplitudes $A^P$, $A^{SV}$ and $A^{SH}$ along with their noise levels
+     $\sigma^P$, $\sigma^{SV}$ and $\sigma^{SH}$ (representing standard deviation of an assumed normal
+     distribution centered at the amplitude value) can be extracted.
+     Given a ray path's azimuth $\phi$ and take-off angles $i$, $j$ of $P$ and $S$ waves respectively,
+     we model the amplitudes as functions of the source mechanism's fault strike $\psi$, fault dip
+     $\delta$ and slip rake $\lambda$, further accounting for the $P$- and $S$-velocities $\alpha_h$,
+     $\beta_h$ at source depth $h$ \citep{AkiRichards1980}. The expressions are:
 
      \begin{align} \label{eq:eq1}
-         \nonumber &A^P \sim \nicefrac{\left( s_R(3\cos^2(i) - 1) - q_R\sin(2i) - p_R\sin^2(i) \right)}{\alpha_h^3}\\
+         \nonumber &A^P \sim \nicefrac{\left( s_R(3\cos^2(i) - 1) - q_R\sin(2i) - p_R\sin^2(i)\right)}{\alpha_h^3}\\
          &A^{SV} \sim \nicefrac{\left( 1.5 s_R\sin(2j) + q_R\cos(2j) + 0.5 p_R\sin(2j) \right)}{\beta_h^3}\\
@@
          \item [i)]
-            Find estimates $\widehat{\psi}, \widehat{\delta}, \widehat{\lambda}$ for the strike, dip and rake that \textit{best} describe the collected data.
+            Find estimates $\widehat{\psi}, \widehat{\delta}, \widehat{\lambda}$ for the strike, dip
+            and rake that \textit{best} describe the collected data.
 
         \item [ii)]
-            Estimate the joint noise distribution of fault parameters to more accurately constrain the set of \textit{acceptable} estimates for the source mechanism.
+            Estimate the joint noise distribution of fault parameters to more accurately constrain the
+            set of \textit{acceptable} estimates for the source mechanism.
 
@@
 
-     We thus need a metric against which to evaluate what the best and acceptable estimates are.
-The misfit function used by \citet{sita_potential_2022} is defined as the angle between the vector of observed amplitudes and that synthesized from parameter estimates.
-In this paper, we shall use a misfit $\mu$ of cosine similarity between the two vectors which has a 1-to-1 correspondence with angle:
+     We thus need a metric against which to evaluate what the best and acceptable estimates are. The
+     misfit function used by \citet{sita_potential_2022} is defined as the angle between the vector of
+     observed amplitudes and that synthesized from parameter estimates. In this paper, we shall use a
+     misfit $\mu$ of cosine similarity between the two vectors which has a 1-to-1 correspondence with angle:
 
@@
 
-     where $\mathbf{A}_o = (A^P, A^{SV}, A^{SH})^\top$ are observed amplitudes, $\mathbf{A}_s = (\widehat{A}^P, \widehat{A}^{SV}, \widehat{A}^{SH})^\top$ are synthetic amplitudes, and $\mathbf{p} = (\widehat{\psi}, \widehat{\delta}, \widehat{\lambda})$ are parameter estimates.
+     where $\mathbf{A}_o = (A^P, A^{SV}, A^{SH})^\top$ are observed amplitudes, $\mathbf{A}_s =
+     (\widehat{A}^P, \widehat{A}^{SV}, \widehat{A}^{SH})^\top$ are synthetic amplitudes, and
+     $\mathbf{p} = (\widehat{\psi}, \widehat{\delta}, \widehat{\lambda})$ are parameter estimates.
 
      \subsection{Justification}
-     We are interested in situations where absolute amplitudes are either unknown or unnecessary, for instance if using data collected from NASA's Insight mission \citep{sita_potential_2022}.
-     In this case, we cannot know the absolute velocities because they depend on an understanding of Mars' interior that is good enough to build a reliable velocity model.
-Alternative misfit functions can account for the relative amplitudes, such as in Equations \ref{eq:eq4} and \ref{eq:eq5}:
+     We are interested in situations where absolute amplitudes are either unknown or unnecessary, for
+     instance if using data collected from NASA's Insight mission \citep{sita_potential_2022}.
+     In this case, we cannot know the absolute velocities because they depend on an understanding of
+     Mars' interior that is good enough to build a reliable velocity model. Alternative misfit
+     functions can account for the relative amplitudes, such as in Equations \ref{eq:eq4} and
+     \ref{eq:eq5}:
 
@@
 
-     First, we note that using cosine similarity is more numerically stable in cases where one of the observed or synthetic amplitudes is close to zero since the norm of the associated vector in Equation \ref{eq:eq3} may still be large enough to make division feasible.
+     First, we note that using cosine similarity is more numerically stable in cases where one of the
+     observed or synthetic amplitudes is close to zero since the norm of the associated vector in
+     Equation \ref{eq:eq3} may still be large enough to make division feasible.
 
-     Secondly, we have chosen cosine similarity over angle because it involves one less operation, which is more computationally efficient and easier to differentiate for directed search algorithms (see Section \ref{sec:algorithms}).
-The Gaussian function in Equation \ref{eq:eq6} is similarly easy to differentiate, but it does not account for relative amplitudes so it would exclude otherwise acceptable solutions.
+     Secondly, we have chosen cosine similarity over angle because it involves one less operation,
+     which is more computationally efficient and easier to differentiate for directed search
+     algorithms (see Section \ref{sec:algorithms}). The Gaussian function in Equation \ref{eq:eq6} is
+     similarly easy to differentiate, but it does not account for relative amplitudes so it would
+     exclude otherwise acceptable solutions.
 
@@
      \subsection{Tolerance}
-     Given vectors $\mathbf{A}_s$ and $\mathbf{A}_o$, we would like to know the largest cosine similarity for which $\mathbf{A}_s$ is considered an acceptable fit for the recorded amplitudes while respecting the asymmetry of error levels for different amplitudes.
-This is visualized in 3D space as a confidence ellipsoid centered at $\mathbf{A}_o$ (Equation \ref{eq:eq7}), for which acceptable fits are vectors that fall inside the elliptical cone with vertex at the origin and tangent to the ellipsoid.
+     Given vectors $\mathbf{A}_s$ and $\mathbf{A}_o$, we would like to know the largest cosine
+     similarity for which $\mathbf{A}_s$ is considered an acceptable fit for the recorded amplitudes
+     while respecting the asymmetry of error levels for different amplitudes. This is visualized in 3D
+     space as a confidence ellipsoid centered at $\mathbf{A}_o$ (Equation \ref{eq:eq7}), for which
+     acceptable fits are vectors that fall inside the elliptical cone with vertex at the origin and
+     tangent to the ellipsoid.
      \begin{align} \label{eq:eq7}
@@
 
-     where $\mathbf{A} = (A^P_b, A^{SV}_b, A^{SH}_b)^\top$ describes a boundary point. The elliptical cone intersects with the ellipsoid at every boundary point for which
+     where $\mathbf{A} = (A^P_b, A^{SV}_b, A^{SH}_b)^\top$ describes a boundary point. The elliptical
+     cone intersects with the ellipsoid at every boundary point for which
      \begin{align} \label{eq:eq8}
@@
 
-     Combining equations \ref{eq:eq7} and \ref{eq:eq8} gives the plane containing every point of tangency:
+     Combining equations \ref{eq:eq7} and \ref{eq:eq8} gives the plane containing every point of 
+     tangency:
      \begin{align} \label{eq:eq9}
@@
 
-     Let $\mathbf{N}_1 = (\nicefrac{A^P}{(\sigma^P)^2}, \nicefrac{A^{SV}}{(\sigma^{SV})^2}, \nicefrac{A^{SH}}{(\sigma^{SH})^2})^\top$ be the normal of this plane.
-Consider also the plane containing vectors $\mathbf{A}_o$ and $\mathbf{A}_s$, with normal
+     Let $\mathbf{N}_1 = (\nicefrac{A^P}{(\sigma^P)^2}, \nicefrac{A^{SV}}{(\sigma^{SV})^2},
+     \nicefrac{A^{SH}}{(\sigma^{SH})^2})^\top$ be the normal of this plane. Consider also the plane
+     containing vectors $\mathbf{A}_o$ and $\mathbf{A}_s$, with normal
      \begin{align} \label{eq:eq10}
@@
 \section*{Data and code availability}
-Zenodo, figshare, and Dryad to archive data and code.
-Citations for datasets and codes should be included in the references, including citations for any seismic networks from which data was used.
-Github is not considered a persistent repository, and we encourage authors to archive a snapshot of any github-hosted code on zenodo.
+Zenodo, figshare, and Dryad to archive data and code. Citations for datasets and codes should be
+included in the references, including citations for any seismic networks from which data was used.
+Github is not considered a persistent repository, and we encourage authors to archive a snapshot of
+any github-hosted code on zenodo.
 

### 2024-10-29 paper/main.tex
latex updates
Done with tolerance, may revisit during uncertainty

@@
 % import packages
-\usepackage{gensymb, nicefrac}
+\usepackage{gensymb, amssymb, nicefrac, bm}
 
@@
     [Suzan] Insert words here...\\
-    Talk about this being an inverse method for a highly nonlinear problem\\
+    Talk about inverse method for a highly nonlinear problem\\
     Writing tip: maximize information content of words
@@
     which the respective amplitudes $A^P$, $A^{SV}$ and $A^{SH}$ along with their noise levels
-     $\sigma^P$, $\sigma^{SV}$ and $\sigma^{SH}$ (representing standard deviation of an assumed normal
+    $\sigma_P$, $\sigma_{SV}$ and $\sigma_{SH}$ (representing standard deviation of an assumed normal
     distribution centered at the amplitude value) can be extracted.
@@
     misfit function used by \citet{sita_potential_2022} is defined as the angle between the vector of
-     observed amplitudes and that synthesized from parameter estimates. In this paper, we shall use a
-     misfit $\mu$ of cosine similarity between the two vectors which has a 1-to-1 correspondence with angle:
+    observed amplitudes and that produced from a forward simulation of synthetic parameters.
+    In this paper, we shall use a misfit $\mu$ of cosine similarity between the two vectors which has a
+    1-to-1 correspondence with angle:
 
     \begin{align} \label{eq:eq3}
-         \mu(\mathbf{p}) = \frac{\mathbf{A}_s^\top\mathbf{A}_o}{\|\mathbf{A}_s\|\|\mathbf{A}_o\|}
+      \mu(\bm{p}) = \frac{\bm{A}_o^\top\bm{A}_s}{\|\bm{A}_o\|\|\bm{A}_s\|}
     \end{align}
 
-     where $\mathbf{A}_o = (A^P, A^{SV}, A^{SH})^\top$ are observed amplitudes, $\mathbf{A}_s =
-     (\widehat{A}^P, \widehat{A}^{SV}, \widehat{A}^{SH})^\top$ are synthetic amplitudes, and
-     $\mathbf{p} = (\widehat{\psi}, \widehat{\delta}, \widehat{\lambda})$ are parameter estimates.
+    where $\bm{A}_o = (A_o^P, A_o^{SV}, A_o^{SH})^\top$ are observed amplitudes, $\bm{A}_s =
+    (A_s^P, A_s^{SV}, A_s^{SH})^\top$ are synthetic amplitudes, and
+    $\bm{p} = (\psi, \delta, \lambda)^\top$ are the underlying parameters.
 
@@
     Mars' interior that is good enough to build a reliable velocity model. Alternative misfit
-     functions can account for the relative amplitudes, such as in Equations \ref{eq:eq4} and
+    functions can account for the relative amplitudes, such as in equations \ref{eq:eq4} and
     \ref{eq:eq5}:
@@
     \begin{align} \label{eq:eq4}
-         \mu(\mathbf{p}) = \sqrt{\left( \Delta \frac{A^{P}}{A^{SV}}\right)^2 + \left( \Delta \frac{A^{SV}}{A^{SH}}\right)^2}
+      \mu(\bm{p}) = \sqrt{\left( \Delta \frac{A^{P}}{A^{SV}}\right)^2 + \left( \Delta \frac{A^{SV}}{A^{SH}}\right)^2}
     \end{align}
@@
     \begin{align} \label{eq:eq5}
-         \mu(\mathbf{p}) = \sqrt{\log^2 \left( \Delta \frac{A^P}{A^{SV}} \right) + \log^2 \left( \Delta \frac{A^{SV}}{A^{SH}} \right)}
+      \mu(\bm{p}) = \sqrt{\log^2 \left( \Delta \frac{A^P}{A^{SV}} \right) + \log^2 \left( \Delta \frac{A^{SV}}{A^{SH}} \right)}
     \end{align}
@@
     observed or synthetic amplitudes is close to zero since the norm of the associated vector in
-     Equation \ref{eq:eq3} may still be large enough to make division feasible.
+    equation \ref{eq:eq3} may still be large enough to make division feasible.
 
@@
     which is more computationally efficient and easier to differentiate for directed search
-     algorithms (see Section \ref{sec:algorithms}). The Gaussian function in Equation \ref{eq:eq6} is
+    algorithms (see Section \ref{sec:algorithms}). The Gaussian function in equation \ref{eq:eq6} is
     similarly easy to differentiate, but it does not account for relative amplitudes so it would
@@
     \begin{align} \label{eq:eq6}
-         &\mu(\mathbf{p}) \sim \exp \left( - \frac{1}{2} (\mathbf{A}_s - \mathbf{A}_o)^\top \mathbf{\Sigma}^{-1} (\mathbf{A}_s - \mathbf{A}_o) \right)\\
-         \nonumber &\text{where } \mathbf{\Sigma} = \text{diag}(\sigma^P, \sigma^{SV}, \sigma^{SH})
+      &\mu(\bm{p}) \sim |\Sigma|^{-1/2} \exp \left( - \frac{1}{2} (\bm{A} - \bm{A}_o)^\top \bm{\Sigma}^{-1} (\bm{A} - \bm{A}_o) \right)
     \end{align}
@@
     \subsection{Tolerance}
-     Given vectors $\mathbf{A}_s$ and $\mathbf{A}_o$, we would like to know the largest cosine
-     similarity for which $\mathbf{A}_s$ is considered an acceptable fit for the recorded amplitudes
-     while respecting the asymmetry of error levels for different amplitudes. This is visualized in 3D
-     space as a confidence ellipsoid centered at $\mathbf{A}_o$ (Equation \ref{eq:eq7}), for which
-     acceptable fits are vectors that fall inside the elliptical cone with vertex at the origin and
-     tangent to the ellipsoid.
+    Given vectors $\bm{A}_s$ and $\bm{A}_o$, we would like to know the largest cosine
+    similarity for which $\bm{A}_s$ is considered an acceptable fit with respect to the recorded amplitudes
+    while accounting for the asymmetry of error levels for different amplitudes. Visualized in 3D
+    space, we have a confidence ellipsoid centered at $\bm{A}_o$ (equation \ref{eq:eq7}) such that
+    acceptable fits are vectors which fall inside the elliptical cone tangent to it with vertex at the
+    origin.
     \begin{align} \label{eq:eq7}
-         \sum_{k \in \{P, SV, SH\}} \frac{(A^k_b-A^k)^2}{(\sigma^k)^2} = 1
+      E(\bm{p}) = (\bm{A} - \bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = 1
     \end{align}
 
-     where $\mathbf{A} = (A^P_b, A^{SV}_b, A^{SH}_b)^\top$ describes a boundary point. The elliptical
+    \noindent where $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ is arbitrary
+    and $\hat{\bm{\Sigma}} = \text{diag}(\sigma_P^2, \sigma_{SV}^2, \sigma_{SH}^2)/\chi^2_{2,\alpha}$ for some $\alpha \in (0,1)$. The elliptical
     cone intersects with the ellipsoid at every boundary point for which
     \begin{align} \label{eq:eq8}
-         \sum_{k \in \{P, SV, SH\}} \frac{2A^k_b(A^k_b-A^k)}{(\sigma^k)^2} = 0
+      (\nabla_{\bm{A}} E)^\top \bm{A} = 2(\bm{A} - \bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} \bm{A} = 0
     \end{align}
@@
     \begin{align} \label{eq:eq9}
-         \sum_{k \in \{P, SV, SH\}} \frac{(A^k_b-A^k)^2}{(\sigma^k)^2} - \frac{2A^k_b(A^k_b-A^k)}{(\sigma^k)^2} = 1\\
-         \nonumber \sum_{k \in \{P, SV, SH\}} \frac{A^k}{(\sigma^k)^2} A^k_b = \left( \sum_{k \in \{P, SV, SH\}} \frac{(A^k)^2}{(\sigma^k)^2} \right) -1
+      \frac{1}{2} (\nabla_{\bm{A}} E)^\top \bm{A} - f(\bm{A}) = 
+              \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = -1
     \end{align}
 
-     Let $\mathbf{N}_1 = (\nicefrac{A^P}{(\sigma^P)^2}, \nicefrac{A^{SV}}{(\sigma^{SV})^2},
-     \nicefrac{A^{SH}}{(\sigma^{SH})^2})^\top$ be the normal of this plane. Consider also the plane
-     containing vectors $\mathbf{A}_o$ and $\mathbf{A}_s$, with normal
+    To account for asymmetry, we recognize that the orientation of $\bm{A}_s$ relative to $\bm{A}_o$
+    has an effect on the tolerance because the associated "thickness" of the ellipsoid changes.
+    This information is captured by the plane containing vectors $\bm{A}_o$ and $\bm{A}_s$:
     \begin{align} \label{eq:eq10}
-         \mathbf{N}_2 = \mathbf{A}_o \times \mathbf{A}_s = (,,)^\top
+      (\bm{A}_s \times \bm{A}_o)^\top \bm{A} = 0
     \end{align}
+
+    To find the constrained points of tangency, we can intersect the original ellipsoid with planes
+    \ref{eq:eq9} and \ref{eq:eq10}. The line of intersection is such that
+    \begin{align} \label{eq:eq11}
+      \bm{A}(t) = (1 - \nicefrac{1}{c_1}) \bm{A}_o + t\bm{D}
+    \end{align}
+
+    \noindent where $c_1$ (equation \ref{eq:eq12}) is a location parameter projecting $\bm{A}_o$ onto plane \ref{eq:eq9}
+    and $\bm{D}$ (equation \ref{eq:eq13}) is the direction perpendicular to planes \ref{eq:eq9} and \ref{eq:eq10}.
+
+    \begin{align} \label{eq:eq12}
+      \nonumber \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} ((1 - \nicefrac{1}{c_1})\bm{A}_o - \bm{A}_o)  = -1\\
+      \Rightarrow c_1 = \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_o = \|\bm{A}_o^\top, \bm{A}_o\|_{\hat{\bm{\Sigma}}^{-1}}^2
+    \end{align}
+    \begin{align} \label{eq:eq13}
+      \nonumber\bm{D} &= (\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1})^\top \times (\bm{A}_s \times \bm{A}_o)\\
+        &= c_1\bm{A}_s - c_2\bm{A}_o
+    \end{align}
+
+    \noindent where $c_2 = \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_s = \langle\bm{A}_o^\top, \bm{A}_s\rangle_{\hat{\bm{\Sigma}}^{-1}}$.
+
+    Combining equations \ref{eq:eq7} and \ref{eq:eq11} produces a quadratic parameterized by the scalar $t$,
+    from which we can obtain a quadratic equation for the points of tangency:
+
+    \begin{align} \label{eq:eq14}
+      \nonumber (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)  = 1\\
+      \nonumber ({\bm{D}^\top \hat{\bm{\Sigma}}^{-1} \bm{D}})t^2 - \nicefrac{2}{c_1}(\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{D})t + (\nicefrac{1}{c_1} - 1) = 0\\
+      % \Rightarrow t = \nicefrac{c_3}{c_1c_2} \pm \nicefrac{1}{c_1c_2} \sqrt{c_3^2 + c_1c_2(c_1 - 1)}
+    \end{align}
+
+    Denote $c_3 = \bm{A}_s^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s^\top, \bm{A}_s\|_{\hat{\bm{\Sigma}}^{-1}}^2$. Then by
+    substitution we have
+    ${\bm{D}^\top \hat{\bm{\Sigma}}^{-1} \bm{D}} = c_1(c_1c_3 - c_2^2)$ and ${\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{D}} = 0$.
+    Solving then yields the necessary points (denoted $\bm{A}_b$) as
+    \begin{align} \label{eq:eq15}
+      \bm{A}_b = \left(1 - \frac{1}{c_1}\right)\bm{A}_o \pm
+        \left(\frac{1}{c_1} \sqrt{\frac{c_1 - 1}{c_1c_3 - c_2^2}}\right)(c_1\bm{A}_s - c_2\bm{A}_o)
+    \end{align}
+
+    Equation \ref{eq:eq15} shows that for the points to be well-defined, the following conditions
+    must hold relative to the inner-product space defined by $\hat{\bm{\Sigma}}^{-1}$:
+
+    \begin{itemize}
+      \item Strict Cauchy-Schwarz inequality: $c_1c_3 > c_2^2$. This also implies that
+        $\hat{\bm{\Sigma}} \succ 0$ and $\bm{A}_s \nparallel \bm{A}_o$. Practically, it means that all
+        uncertainties are necessary and inversion methods that depend on forward simulation should
+        treat parallelism as an edge case.
+      \item Confidence ellipsoid cannot contain origin: $c_1 \geq 1$. Practically, we should expect
+      errors to be a small fraction of observed amplitudes.
+    \end{itemize}
+
+    Thus, given 2 nonparallel vectors $\bm{A}_o$ and $\bm{A}_s$ and positive definite covariance
+    matrix $\hat{\bm{\Sigma}}$, we accept the simulation if $\hat{\bm{A}}_o^\top \hat{\bm{A}_s} \geq \text{tol} 
+    = \min_b \hat{\bm{A}}_o^\top \hat{\bm{A}_b}$. The condition, more formally, is
+    \begin{align} \label{eq:eq16}
+      \max_b \hat{\bm{A}}_o^\top (\hat{\bm{A}}_b - \hat{\bm{A}}_s) \leq 0
+    \end{align}
+
+    \noindent where $\hat{\bm{A}}_o$, $\hat{\bm{A}}_s$ and $\hat{\bm{A}}_b$ are unit vectors.
 
@@
 
-    Insert words here...
+    The vector $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ can be treated as a continuously differentiable nonlinear function of
+    parameters $\bm{p} = (\psi, \delta, \lambda)^\top$, defined by equation \ref{eq:eq1}. We can thus apply
+    directed search methods to minimize the following cost function:
+    \begin{align} \label{eq:eq17}
+      \mu(\bm{p}) = \hat{\bm{A}}_o^\top \hat{\bm{A}}
+    \end{align}
+
+    TODO: draw computational graph, derive gradients
 

### 2024-10-29 paper/main.tex
Computational graph

@@
 % import packages
-\usepackage{gensymb, amssymb, nicefrac, bm}
+\usepackage{gensymb, amssymb, nicefrac, bm, tikz}
 
@@
 
-    \begin{align} \label{eq:eq1}
+    \begin{align} \label{eq:1}
       \nonumber &A^P \sim \nicefrac{\left( s_R(3\cos^2(i) - 1) - q_R\sin(2i) - p_R\sin^2(i)\right)}{\alpha_h^3}\\
@@
 
-    \begin{align} \label{eq:eq2}
+    \begin{align} \label{eq:2}
       \nonumber s_R &= 0.5 \sin(\lambda) \sin(2\delta)\\
@@
 
-    \begin{align} \label{eq:eq3}
+    \begin{align} \label{eq:3}
       \mu(\bm{p}) = \frac{\bm{A}_o^\top\bm{A}_s}{\|\bm{A}_o\|\|\bm{A}_s\|}
@@
     Mars' interior that is good enough to build a reliable velocity model. Alternative misfit
-    functions can account for the relative amplitudes, such as in equations \ref{eq:eq4} and
-    \ref{eq:eq5}:
+    functions can account for the relative amplitudes, such as in equations \ref{eq:4} and
+    \ref{eq:5}:
 
-    \begin{align} \label{eq:eq4}
+    \begin{align} \label{eq:4}
       \mu(\bm{p}) = \sqrt{\left( \Delta \frac{A^{P}}{A^{SV}}\right)^2 + \left( \Delta \frac{A^{SV}}{A^{SH}}\right)^2}
@@
 
-    \begin{align} \label{eq:eq5}
+    \begin{align} \label{eq:5}
       \mu(\bm{p}) = \sqrt{\log^2 \left( \Delta \frac{A^P}{A^{SV}} \right) + \log^2 \left( \Delta \frac{A^{SV}}{A^{SH}} \right)}
@@
     observed or synthetic amplitudes is close to zero since the norm of the associated vector in
-    equation \ref{eq:eq3} may still be large enough to make division feasible.
+    equation \ref{eq:3} may still be large enough to make division feasible.
 
@@
     which is more computationally efficient and easier to differentiate for directed search
-    algorithms (see Section \ref{sec:algorithms}). The Gaussian function in equation \ref{eq:eq6} is
+    algorithms (see Section \ref{sec:algorithms}). The Gaussian function in equation \ref{eq:6} is
     similarly easy to differentiate, but it does not account for relative amplitudes so it would
@@
 
-    \begin{align} \label{eq:eq6}
+    \begin{align} \label{eq:6}
       &\mu(\bm{p}) \sim |\Sigma|^{-1/2} \exp \left( - \frac{1}{2} (\bm{A} - \bm{A}_o)^\top \bm{\Sigma}^{-1} (\bm{A} - \bm{A}_o) \right)
@@
     while accounting for the asymmetry of error levels for different amplitudes. Visualized in 3D
-    space, we have a confidence ellipsoid centered at $\bm{A}_o$ (equation \ref{eq:eq7}) such that
+    space, we have a confidence ellipsoid centered at $\bm{A}_o$ (equation \ref{eq:7}) such that
     acceptable fits are vectors which fall inside the elliptical cone tangent to it with vertex at the
     origin.
-    \begin{align} \label{eq:eq7}
+    \begin{align} \label{eq:7}
       E(\bm{p}) = (\bm{A} - \bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = 1
@@
     cone intersects with the ellipsoid at every boundary point for which
-    \begin{align} \label{eq:eq8}
+    \begin{align} \label{eq:8}
       (\nabla_{\bm{A}} E)^\top \bm{A} = 2(\bm{A} - \bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} \bm{A} = 0
@@
 
-    Combining equations \ref{eq:eq7} and \ref{eq:eq8} gives the plane containing every point of 
+    Combining equations \ref{eq:7} and \ref{eq:8} gives the plane containing every point of 
     tangency:
-    \begin{align} \label{eq:eq9}
+    \begin{align} \label{eq:9}
       \frac{1}{2} (\nabla_{\bm{A}} E)^\top \bm{A} - f(\bm{A}) = 
@@
     This information is captured by the plane containing vectors $\bm{A}_o$ and $\bm{A}_s$:
-    \begin{align} \label{eq:eq10}
+    \begin{align} \label{eq:10}
       (\bm{A}_s \times \bm{A}_o)^\top \bm{A} = 0
@@
     To find the constrained points of tangency, we can intersect the original ellipsoid with planes
-    \ref{eq:eq9} and \ref{eq:eq10}. The line of intersection is such that
-    \begin{align} \label{eq:eq11}
+    \ref{eq:9} and \ref{eq:10}. The line of intersection is such that
+    \begin{align} \label{eq:11}
       \bm{A}(t) = (1 - \nicefrac{1}{c_1}) \bm{A}_o + t\bm{D}
@@
 
-    \noindent where $c_1$ (equation \ref{eq:eq12}) is a location parameter projecting $\bm{A}_o$ onto plane \ref{eq:eq9}
-    and $\bm{D}$ (equation \ref{eq:eq13}) is the direction perpendicular to planes \ref{eq:eq9} and \ref{eq:eq10}.
+    \noindent where $c_1$ (equation \ref{eq:12}) is a location parameter projecting $\bm{A}_o$ onto plane \ref{eq:9}
+    and $\bm{D}$ (equation \ref{eq:13}) is the direction perpendicular to planes \ref{eq:9} and \ref{eq:10}.
 
-    \begin{align} \label{eq:eq12}
+    \begin{align} \label{eq:12}
       \nonumber \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} ((1 - \nicefrac{1}{c_1})\bm{A}_o - \bm{A}_o)  = -1\\
@@
     \end{align}
-    \begin{align} \label{eq:eq13}
+    \begin{align} \label{eq:13}
       \nonumber\bm{D} &= (\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1})^\top \times (\bm{A}_s \times \bm{A}_o)\\
@@
 
-    Combining equations \ref{eq:eq7} and \ref{eq:eq11} produces a quadratic parameterized by the scalar $t$,
+    Combining equations \ref{eq:7} and \ref{eq:11} produces a quadratic parameterized by the scalar $t$,
     from which we can obtain a quadratic equation for the points of tangency:
 
-    \begin{align} \label{eq:eq14}
+    \begin{align} \label{eq:14}
       \nonumber (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)  = 1\\
-      \nonumber ({\bm{D}^\top \hat{\bm{\Sigma}}^{-1} \bm{D}})t^2 - \nicefrac{2}{c_1}(\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{D})t + (\nicefrac{1}{c_1} - 1) = 0\\
-      % \Rightarrow t = \nicefrac{c_3}{c_1c_2} \pm \nicefrac{1}{c_1c_2} \sqrt{c_3^2 + c_1c_2(c_1 - 1)}
+      ({\bm{D}^\top \hat{\bm{\Sigma}}^{-1} \bm{D}})t^2 - 
+          \nicefrac{2}{c_1}(\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{D})t + (\nicefrac{1}{c_1} - 1) = 0
     \end{align}
@@
     Solving then yields the necessary points (denoted $\bm{A}_b$) as
-    \begin{align} \label{eq:eq15}
+    \begin{align} \label{eq:15}
       \bm{A}_b = \left(1 - \frac{1}{c_1}\right)\bm{A}_o \pm
@@
 
-    Equation \ref{eq:eq15} shows that for the points to be well-defined, the following conditions
+    Equation \ref{eq:15} shows that for the points to be well-defined, the following conditions
     must hold relative to the inner-product space defined by $\hat{\bm{\Sigma}}^{-1}$:
@@
     = \min_b \hat{\bm{A}}_o^\top \hat{\bm{A}_b}$. The condition, more formally, is
-    \begin{align} \label{eq:eq16}
+    \begin{align} \label{eq:16}
       \max_b \hat{\bm{A}}_o^\top (\hat{\bm{A}}_b - \hat{\bm{A}}_s) \leq 0
@@
     The vector $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ can be treated as a continuously differentiable nonlinear function of
-    parameters $\bm{p} = (\psi, \delta, \lambda)^\top$, defined by equation \ref{eq:eq1}. We can thus apply
+    parameters $\bm{p} = (\psi, \delta, \lambda)^\top$, defined by equation \ref{eq:1}. We can thus apply
     directed search methods to minimize the following cost function:
-    \begin{align} \label{eq:eq17}
+    \begin{align} \label{eq:17}
       \mu(\bm{p}) = \hat{\bm{A}}_o^\top \hat{\bm{A}}
@@
 
-    TODO: draw computational graph, derive gradients
+    \textcolor{red}{Continue from here...}
+
+    \begin{figure}[ht!]
+      \centering
+      \begin{tikzpicture}[every node/.style={circle, draw, minimum size=.8cm}]
+        \node (P1) at (0,2) {$\psi$};
+        \node (P2) at (0,0) {$\delta$};
+        \node (P3) at (0,-2) {$\lambda$};
+
+        \node (H1) at (2.6,3) {$s_R$};
+        \node (H2) at (2.6,1.5) {$q_R$};
+        \node (H3) at (2.6,0) {$p_R$};
+        \node (H4) at (2.6,-1.5) {$q_L$};
+        \node (H5) at (2.6,-3) {$p_L$};
+
+        \node (A1) at (5,1.5) {\scriptsize $A^P$};
+        \node (A2) at (5,0) {\scriptsize $A^{SV}$};
+        \node (A3) at (5,-1.5) {\scriptsize $A^{SH}$};
+
+        \node (M) at (7,0) {$\mu$};
+
+        \foreach \p in {2,3}
+          \foreach \h in {1,2,3,4,5}
+            \draw (P\p) -- (H\h);
+
+        % \foreach \h in {1,2,3,4,5}
+        %   \foreach \a in {1,2,3}
+        %     \draw[dashed] (H\h) -- (A\a);
+
+        \foreach \a in {1,2,3}
+          \draw (A\a) -- (M);
+
+        % \draw[dashed] (P1) -- (H1);
+        \draw (P1) -- (H2);
+        \draw (P1) -- (H3);
+        \draw (P1) -- (H4);
+        \draw (P1) -- (H5);
+
+        \draw (H1) -- (A3);
+        \draw (H1) -- (A1);
+        \draw (H1) -- (A2);
+        \draw (H2) -- (A1);
+        \draw (H2) -- (A2);
+        \draw (H3) -- (A1);
+        \draw (H3) -- (A2);
+        \draw (H4) -- (A3);
+        \draw (H5) -- (A3);
+      \end{tikzpicture}
+      \caption{Computational graph of forwayd model.}
+      \label{fig:1}
+    \end{figure}
 

### 2024-10-29 paper/main.tex
Up to algorithms

@@
 % import packages
-\usepackage{gensymb, amssymb, nicefrac, bm, tikz}
+\usepackage{amssymb, nicefrac, bm, tikz}
 
@@
     [Suzan] Insert words here...\\
-    Talk about inverse method for a highly nonlinear problem\\
     Writing tip: maximize information content of words
@@
 
-    \subsection{Derivation}
+  \subsection{Background}
     When a quake event happens, a single seismometer reads $P$, $SV$ and $SH$ body waves through
@@
     \begin{align} \label{eq:1}
-      \nonumber &A^P \sim \nicefrac{\left( s_R(3\cos^2(i) - 1) - q_R\sin(2i) - p_R\sin^2(i)\right)}{\alpha_h^3}\\
-      &A^{SV} \sim \nicefrac{\left( 1.5 s_R\sin(2j) + q_R\cos(2j) + 0.5 p_R\sin(2j) \right)}{\beta_h^3}\\
-      \nonumber &A^{SH} \sim \nicefrac{\left( q_L\cos(j) + p_L\sin(j) \right)}{\beta_h^3}
+      \nonumber &A^P \sim \left( s_R(3\cos^2(i) - 1) - q_R\sin(2i) - p_R\sin^2(i)\right)/\alpha_h^3\\
+      \nonumber &A^{SV} \sim \left( 1.5 s_R\sin(2j) + q_R\cos(2j) + 0.5 p_R\sin(2j) \right)/\beta_h^3\\
+      &A^{SH} \sim \left( q_L\cos(j) + p_L\sin(j) \right)/\beta_h^3
     \end{align}
@@
       \nonumber q_R &= \sin(\lambda) \cos(2\delta) \sin(\psi_r) + \cos(\lambda) \cos(\delta) \cos(\psi_r)\\
-      p_R &= \cos(\lambda) \sin(\delta) \sin(2\psi_r) - 0.5\sin(\lambda) \sin(2\delta) \cos(2\psi_r)\\
+      \nonumber p_R &= \cos(\lambda) \sin(\delta) \sin(2\psi_r) - 0.5\sin(\lambda) \sin(2\delta) \cos(2\psi_r)\\
       \nonumber p_L &= 0.5\sin(\lambda) \sin(2\delta) \sin(2\psi_r) + \cos(\lambda) \sin(\delta) \cos(2\psi_r)\\
-      \nonumber q_L &= -\cos(\lambda) \cos(\delta) \sin(\psi_r) + \sin(\lambda) \cos(2\delta) \cos(\psi_r)
+      q_L &= -\cos(\lambda) \cos(\delta) \sin(\psi_r) + \sin(\lambda) \cos(2\delta) \cos(\psi_r)
     \end{align}
@@
     observed amplitudes and that produced from a forward simulation of synthetic parameters.
-    In this paper, we shall use a misfit $\mu$ of cosine similarity between the two vectors which has a
-    1-to-1 correspondence with angle:
+    In this paper, we shall use a misfit $\Phi$ of negative cosine similarity between the two vectors
+    which has a 1-to-1 correspondence with angle:
 
     \begin{align} \label{eq:3}
-      \mu(\bm{p}) = \frac{\bm{A}_o^\top\bm{A}_s}{\|\bm{A}_o\|\|\bm{A}_s\|}
+      \Phi(\bm{p}) = \frac{-\bm{A}_o^\top\bm{A}_s}{\|\bm{A}_o\|\|\bm{A}_s\|}
     \end{align}
@@
     \begin{align} \label{eq:4}
-      \mu(\bm{p}) = \sqrt{\left( \Delta \frac{A^{P}}{A^{SV}}\right)^2 + \left( \Delta \frac{A^{SV}}{A^{SH}}\right)^2}
+      \Phi(\bm{p}) = \sqrt{\left( \Delta \frac{A^{P}}{A^{SV}}\right)^2 + \left( \Delta \frac{A^{SV}}{A^{SH}}\right)^2}
     \end{align}
@@
     \begin{align} \label{eq:5}
-      \mu(\bm{p}) = \sqrt{\log^2 \left( \Delta \frac{A^P}{A^{SV}} \right) + \log^2 \left( \Delta \frac{A^{SV}}{A^{SH}} \right)}
+      \Phi(\bm{p}) = \sqrt{\log^2 \left( \Delta \frac{A^P}{A^{SV}} \right) + \log^2 \left( \Delta \frac{A^{SV}}{A^{SH}} \right)}
     \end{align}
@@
     \begin{align} \label{eq:6}
-      &\mu(\bm{p}) \sim |\Sigma|^{-1/2} \exp \left( - \frac{1}{2} (\bm{A} - \bm{A}_o)^\top \bm{\Sigma}^{-1} (\bm{A} - \bm{A}_o) \right)
+      &\Phi(\bm{p}) \sim |\Sigma|^{-1/2} \exp \left( - \frac{1}{2} (\bm{A} - \bm{A}_o)^\top \bm{\Sigma}^{-1} (\bm{A} - \bm{A}_o) \right)
     \end{align}
 
-    \subsection{Tolerance}
+  \subsection{Tolerance derivation}
     Given vectors $\bm{A}_s$ and $\bm{A}_o$, we would like to know the largest cosine
@@
 
-    The vector $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ can be treated as a continuously differentiable nonlinear function of
-    parameters $\bm{p} = (\psi, \delta, \lambda)^\top$, defined by equation \ref{eq:1}. We can thus apply
-    directed search methods to minimize the following cost function:
+  \subsection{Misfit gradient}
+    The vector $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ can be treated as a continuously differentiable 
+    nonlinear function of parameters $\bm{p} = (\psi, \delta, \lambda)^\top$, defined by equation 
+    \ref{eq:1}. We can thus apply directed search methods to minimize the following cost function:
     \begin{align} \label{eq:17}
-      \mu(\bm{p}) = \hat{\bm{A}}_o^\top \hat{\bm{A}}
+      \Phi(\bm{p}) = -\hat{\bm{A}}_o^\top \hat{\bm{A}}
     \end{align}
 
-    \textcolor{red}{Continue from here...}
+    To calculate the gradient, first note that $\bm{A}$ is a linear combination of nonlinear functions
+    of $\bm{p}$ (equation \ref{eq:2}).
 
@@
 
-        \node (M) at (7,0) {$\mu$};
+        \node (M) at (7,0) {$\Phi$};
 
@@
       \end{tikzpicture}
-      \caption{Computational graph of forwayd model.}
+      \caption{Computational graph of forward model (edges are non-zero partial derivatives).}
       \label{fig:1}
@@
 
+    Define a vector $\bm{\eta} = (s_R, q_R, p_R, q_L, p_L)^\top$. Then we can encode
+    the misfit as
+    \begin{align} \label{eq:18}
+      \Phi(\bm{p}) = \Phi(\bm{A}(\bm{\eta}(\bm{p})))
+    \end{align}
+
+    \noindent This composition is visualized as a computational graph in figure \ref{fig:1}.
+    Each stage of the composition has an associated Jacobian, which we can demonstrate layer by layer:
+
+    i) From $\bm{p}$ to $\bm{\eta}$, we can use equation \ref{eq:2} to define the Jacobian
+    as follows:
+    \begin{align*}
+      &a = \sin(\lambda)\;\;\;\;\;\;\;\;\;\;\;\;\; b = \cos(2\delta)\;\;\;\;\;\;\;\;\;\;\; c = \cos(\lambda)\\
+      &d = \sin(2\delta)\;\;\;\;\;\;\;\;\;\;\;\; e = \sin(\delta)\;\;\;\;\;\;\;\;\;\;\;\; f = \cos(\delta)\\
+      &g = \sin(\psi - \phi)\;\;\;\;\;\; h = \cos(\psi - \phi)\;\;\;\;\; p = cos(2(\psi - \phi))\\
+      &q = \sin(2(\psi - \phi))
+    \end{align*}
+    \begin{align} \label{eq:19}
+      \nonumber&\bm{J_p}(\bm{\eta}) = \\
+      &\begin{bmatrix}
+        0         &ab         &0.5cd\\
+        abh-cfg   &-2adg-ceh  &cbg-afh\\
+        2cep+adq  &cfq-abp    &-aeq-0.5cdp\\
+        -cfh-abg  &ceg-2adh   &afg+cbh\\
+        adp-2ceq  &abq+cfp    &0.5cdq-aep
+      \end{bmatrix}
+    \end{align}
+
+    ii) From $\bm{\eta}$ to $A$, we have a linear transformation whose Jacobian is the matrix
+    of coefficients from equation \ref{eq:1}:
+    \begin{align} \label{eq:20}
+      \nonumber&\bm{J_\eta}(\bm{A}) = \\
+      &\begin{bmatrix}
+        \frac{3\cos^2(i)-1}{\alpha^3}    &\frac{-\sin(2i)}{\alpha^3}    &\frac{-sin^2(i)}{\alpha^3}   &0    &0\\
+        \frac{1.5\sin(2j)}{\beta^3}      &\frac{\cos(2j)}{\beta^3}      &\frac{0.5\sin(2j)}{\beta^3}  &0    &0\\
+        0         &0           &0        &\frac{\cos(j)}{\beta^3}       &\frac{\sin(j)}{\beta^3}
+      \end{bmatrix}
+    \end{align}
+
+    iii) From $\bm{A}$ to $\Phi$, we have a scalar function whose Jacobian is the gradient
+    transposed.
+    \begin{align} \label{eq:21}
+      \nonumber\nabla_{\bm{A}}\Phi &= \frac{\|\bm{A}\|\nabla_{\bm{A}}(-\hat{\bm{A}}_o^\top\bm{A}) + 
+      \hat{\bm{A}}_o^\top \bm{A} \nabla_{\bm{A}}(\|\bm{A}\|)}{\|\bm{A}\|^2}\\
+            &= \frac{1}{\|\bm{A}\|}(\Phi\hat{\bm{A}} - \hat{\bm{A}}_o)
+    \end{align}
+
+    Combining \ref{eq:19}, \ref{eq:20} and \ref{eq:21} gives the overall gradient:
+    \begin{align} \label{eq:22}
+      \nabla_{\bm{p}}\Phi = \bm{J_p}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\Phi
+    \end{align}
+
+  \subsection{Optimization algorithms}
+    Something about gradient-based methods. Try different algorithms, they may produce
+    interesting results >> IEMS 450-2.
+
+    \noindent \textbf{TODO:} Create one 3D image with the results of gradient descent from random points in space.
+
 \section{Uncertainty Quantification} \label{sec:uncertainty}
 
-    Insert words here...
+    \subsection{Error propagation}
+      Say something here, composed Jacobian is important >> Earth 353, IEMS 401.
 
-\section{Example: from Paula} \label{sec:example}
+      \noindent Emphasize the interest in modality.
 
-    Insert words here...
+    \subsection{Monte Carlo methods}
+      Place to talk about different seeding stategies and their effect on uncertainty.
+      Why hybrid search was used, why systematic over random.
+
+    \subsection{Kagan angle}
+      An alternative way of quantifying model deviations.
+
+\section{Testing} \label{sec:example}
+
+    \subsection{Synthetic data}
+      Say something here about the Monte Carlo
+
+    \subsection{Earth data}
+      From Uganda (Paula)
+
+    \subsection{Mars data}
+      From Insight (Suzan/Maddy)
 
@@
 
-    Insert words here...
+    \subsection{Performance overview}
+      Put some words here...
+
+    \subsection{Growth avenues}
+      Put some more words here...
 
@@
 \section*{Data and code availability}
-Zenodo, figshare, and Dryad to archive data and code. Citations for datasets and codes should be
-included in the references, including citations for any seismic networks from which data was used.
-Github is not considered a persistent repository, and we encourage authors to archive a snapshot of
-any github-hosted code on zenodo.
+Create a new repo with beautiful code, READMEs and everything. One function/algorithm at a time.
 

### 2024-10-29 paper/main.tex
tilde lol

@@
     \begin{align} \label{eq:7}
-      E(\bm{p}) = (\bm{A} - \bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = 1
+      E(\bm{p}) = (\bm{A} - \bm{A}_o)^\top \tilde{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = 1
     \end{align}
@@
     \noindent where $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ is arbitrary
-    and $\hat{\bm{\Sigma}} = \text{diag}(\sigma_P^2, \sigma_{SV}^2, \sigma_{SH}^2)/\chi^2_{2,\alpha}$ for some $\alpha \in (0,1)$. The elliptical
+    and $\tilde{\bm{\Sigma}} = \text{diag}(\sigma_P^2, \sigma_{SV}^2, \sigma_{SH}^2)/\chi^2_{2,\alpha}$ for some $\alpha \in (0,1)$. The elliptical
     cone intersects with the ellipsoid at every boundary point for which
     \begin{align} \label{eq:8}
-      (\nabla_{\bm{A}} E)^\top \bm{A} = 2(\bm{A} - \bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} \bm{A} = 0
+      (\nabla_{\bm{A}} E)^\top \bm{A} = 2(\bm{A} - \bm{A}_o)^\top \tilde{\bm{\Sigma}}^{-1} \bm{A} = 0
     \end{align}
@@
       \frac{1}{2} (\nabla_{\bm{A}} E)^\top \bm{A} - f(\bm{A}) = 
-              \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = -1
+              \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = -1
     \end{align}
@@
     \begin{align} \label{eq:12}
-      \nonumber \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} ((1 - \nicefrac{1}{c_1})\bm{A}_o - \bm{A}_o)  = -1\\
-      \Rightarrow c_1 = \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_o = \|\bm{A}_o^\top, \bm{A}_o\|_{\hat{\bm{\Sigma}}^{-1}}^2
+      \nonumber \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} ((1 - \nicefrac{1}{c_1})\bm{A}_o - \bm{A}_o)  = -1\\
+      \Rightarrow c_1 = \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_o = \|\bm{A}_o^\top, \bm{A}_o\|_{\tilde{\bm{\Sigma}}^{-1}}^2
     \end{align}
     \begin{align} \label{eq:13}
-      \nonumber\bm{D} &= (\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1})^\top \times (\bm{A}_s \times \bm{A}_o)\\
+      \nonumber\bm{D} &= (\bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1})^\top \times (\bm{A}_s \times \bm{A}_o)\\
         &= c_1\bm{A}_s - c_2\bm{A}_o
@@
 
-    \noindent where $c_2 = \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_s = \langle\bm{A}_o^\top, \bm{A}_s\rangle_{\hat{\bm{\Sigma}}^{-1}}$.
+    \noindent where $c_2 = \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \langle\bm{A}_o^\top, \bm{A}_s\rangle_{\tilde{\bm{\Sigma}}^{-1}}$.
 
@@
     \begin{align} \label{eq:14}
-      \nonumber (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)  = 1\\
-      ({\bm{D}^\top \hat{\bm{\Sigma}}^{-1} \bm{D}})t^2 - 
-          \nicefrac{2}{c_1}(\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{D})t + (\nicefrac{1}{c_1} - 1) = 0
+      \nonumber (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)^\top \tilde{\bm{\Sigma}}^{-1} (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)  = 1\\
+      ({\bm{D}^\top \tilde{\bm{\Sigma}}^{-1} \bm{D}})t^2 - 
+          \nicefrac{2}{c_1}(\bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{D})t + (\nicefrac{1}{c_1} - 1) = 0
     \end{align}
 
-    Denote $c_3 = \bm{A}_s^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s^\top, \bm{A}_s\|_{\hat{\bm{\Sigma}}^{-1}}^2$. Then by
+    Denote $c_3 = \bm{A}_s^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s^\top, \bm{A}_s\|_{\tilde{\bm{\Sigma}}^{-1}}^2$. Then by
     substitution we have
-    ${\bm{D}^\top \hat{\bm{\Sigma}}^{-1} \bm{D}} = c_1(c_1c_3 - c_2^2)$ and ${\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{D}} = 0$.
+    ${\bm{D}^\top \tilde{\bm{\Sigma}}^{-1} \bm{D}} = c_1(c_1c_3 - c_2^2)$ and ${\bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{D}} = 0$.
     Solving then yields the necessary points (denoted $\bm{A}_b$) as
@@
     Equation \ref{eq:15} shows that for the points to be well-defined, the following conditions
-    must hold relative to the inner-product space defined by $\hat{\bm{\Sigma}}^{-1}$:
+    must hold relative to the inner-product space defined by $\tilde{\bm{\Sigma}}^{-1}$:
 
@@
       \item Strict Cauchy-Schwarz inequality: $c_1c_3 > c_2^2$. This also implies that
-        $\hat{\bm{\Sigma}} \succ 0$ and $\bm{A}_s \nparallel \bm{A}_o$. Practically, it means that all
+        $\tilde{\bm{\Sigma}} \succ 0$ and $\bm{A}_s \nparallel \bm{A}_o$. Practically, it means that all
         uncertainties are necessary and inversion methods that depend on forward simulation should
@@
     Thus, given 2 nonparallel vectors $\bm{A}_o$ and $\bm{A}_s$ and positive definite covariance
-    matrix $\hat{\bm{\Sigma}}$, we accept the simulation if $\hat{\bm{A}}_o^\top \hat{\bm{A}_s} \geq \text{tol} 
+    matrix $\tilde{\bm{\Sigma}}$, we accept the simulation if $\hat{\bm{A}}_o^\top \hat{\bm{A}_s} \geq \text{tol} 
     = \min_b \hat{\bm{A}}_o^\top \hat{\bm{A}_b}$. The condition, more formally, is

### 2024-11-03 src/functions.py
.

@@ -1,0 +1,38 @@
+# every necessary function for the paper is defined here
+
+import numpy as np
+from numpy import linalg
+
+
+def unit_vector(vector: list):
+    """ Returns the unit vector of a vector. """
+    return vector / linalg.norm(vector)
+
+
+def boundary_points(Ao: list, Uo: list, As: list = [], chi2: float = 1):
+    """ Returns the boundary points of the model.
+
+    Args:
+        Ao (list): Vector of observed amplitudes.
+        Uo (list): Vector of measured uncertainties.
+        As (list, optional): _description_. Defaults to [].
+        chi2 (float, optional): _description_. Defaults to 1.
+    """
+    if len(As) == 0:
+        As = np.zeros(len(Ao))
+        As[0] = 1
+    
+    Sigma = np.diag(Uo)/chi2
+    Sigma_inv = linalg.inv(Sigma)
+    
+        
+
+
+def main():
+    # boundary_points(np.zeros(3), np.zeros(3), np.zeros(3))
+    pass
+
+
+if __name__ == '__main__':
+    main()
+    

### 2024-11-03 src/minimize_objective.py

@@ -1,0 +1,333 @@
+#!/usr/bin/env python2
+# -*- coding: utf-8 -*-
+
+'''
+Name: Victor Agaba
+
+Date: 2023-20-11
+
+A module with 3 functions for implementing Newton method and gradient descent:
+    init_options()
+    minimize()
+    factorize_with_mod()
+'''
+
+# Import numpy for linear algebra computations
+import numpy as np
+
+# Import scipy to use in Cholesky factorization
+import scipy as sp
+
+
+def init_options():
+    '''
+    Option initialization for the gradient descent algorithm
+
+    Initialize algorithm options with default values
+
+    Return values:
+        options:
+            This is a dictionary with fields that correspond to algorithmic
+            options of our method.  In particular:
+
+            max_iter:
+                Maximum number of iterations
+            tol:
+                Convergence tolerance
+            step_type:
+                Different ways to calculate the search direction:
+                    'gradient_descent'
+                    'Newton'
+            linesearch:
+                Type of line search:
+                    'constant'
+                    'backtracking'
+            alpha_init:
+                First trial step size to attempt during line search:
+                    'constant':     Value of the constant step size
+                    'backtracking': Initial trial step size
+            suff_decrease_factor:
+                Coefficient in sufficient decrease condition for backtracking
+                    line search
+            perturb_init:
+                Initial perturbation of Hessian matrix to ensure positive
+                    definiteness
+            perturb_inc_factor:
+                Increase factor for Hessian perturbation
+            output_level:
+                Amount of output to be printed
+                0: No output
+                1: Only summary information
+                2: One line per iteration (good for debugging)
+                3: Print iterates in each iteration
+    '''
+    options = {}
+    options['max_iter'] = 1000
+    options['tol'] = 1e-6
+    options['step_type'] = 'gradient_descent'
+    options['linesearch'] = 'backtracking'
+    options['alpha_init'] = 1
+    options['suff_decrease_factor'] = 1e-4
+    options['perturb_init'] = 1e-4
+    options['perturb_inc_factor'] = 10
+    options['output_level'] = 1
+
+    return options
+
+
+def minimize(opt_prob, options, x_start=None):
+    '''
+    Optimization methods for unconstrainted optimization
+
+    This is an implementation of the gradient descent method for unconstrained
[…]

### 2024-11-03 src/seismic_model.py

@@ -1,0 +1,152 @@
+'''
+Name: Victor Agaba
+
+Date: 2nd November 2024
+
+The goal of this file is to create a model of the nonlinear inverse
+problem to which we will apply the optimization algorithms.
+'''
+
+
+import numpy as np
+from numpy import linalg
+from matplotlib import pyplot as plt
+import functions as fn
+
+
+class InvProblem:
+    '''
+    Base class for an unconstrained optimzation problem.
+    Catch anything whose method is not implemented.
+    '''
+
+    def __call__(self, p):
+        raise Exception('Class is not callable')
+
+    def gradient(self, p):
+        raise Exception('Method "gradient" is not implemented')
+
+    def hessian(self, p):
+        raise Exception('Method "hessian" not is implemented')
+
+    def starting_point(self):
+        raise Exception('Method "starting_point" not implemented')
+
+
+class SeismicModel(InvProblem):
+    '''
+    This class computes the misfit function and its gradient
+    for a given model.
+    Uses the same notation as in the paper.
+    '''
+    def __init__(self, phi, i, j, alpha, beta, Ao):
+        '''
+        Initialize with the parts of the model that don't change.
+        ALL ANGLES ARE IN RADIANS!!!
+        
+        Args:
+        phi = raypath's azimuth
+        i, j = P and S wave takeoff angles
+        alpha, beta = P and S wave velocities
+        Ao = observed amplitudes
+        '''
+        # initialize misfits iterates for tracking
+        self.misfits = []
+        self.iterates = []
+        
+        # initialize constant Jacobian matrix
+        factors = np.array([alpha, beta, beta])**3
+        self.J_A = np.array(
+            [[(3*np.cos(i)**2 - 1), -np.sin(2*i), -np.sin(i)**2, 0, 0],
+             [1.5*np.sin(2*j), np.cos(2*j), 0.5*np.sin(2*j), 0, 0],
+             [0, 0, 0, np.cos(j), np.sin(j)]])
+        
+        # scale each row of J_A by the corresponding factor
+        self.J_A /= factors[:,np.newaxis]
+        
+        # also sture J and nabla_E for later use
+        self.J_eta = np.zeros((5,3))
+        self.nabla_Phi = np.zeros(3)
+        
[…]

### 2024-11-04 src/functions.py
functions

@@ -3,6 +3,7 @@
 import numpy as np
 from numpy import linalg
 
+tol = 1e-6
 
 def unit_vector(vector: list):
     """ Returns the unit vector of a vector. """
@@ -15,17 +16,46 @@ def boundary_points(Ao: list, Uo: list, As: list = [], chi2: float = 1):
     Args:
         Ao (list): Vector of observed amplitudes.
         Uo (list): Vector of measured uncertainties.
-        As (list, optional): _description_. Defaults to [].
+        As (list, optional): Vector of simulated amplitudes. Defaults to [].
         chi2 (float, optional): _description_. Defaults to 1.
     """
     if len(As) == 0:
         As = np.zeros(len(Ao))
         As[0] = 1
+        if True: # TODO:...
+            As[1] = 1
+            
     
+    # adjusted covariance matrix
+    assert all(Uo > 0), "All uncertainties must be positive."
     Sigma = np.diag(Uo)/chi2
     Sigma_inv = linalg.inv(Sigma)
     
-        
+    # constants
+    c1 = Ao@Sigma_inv@Ao
+    c2 = Ao@Sigma_inv@As
+    c3 = As@Sigma_inv@As
+    
+    # conditions
+    assert c1 >= 1, "COnfidence ellipsoid is too big."
+    assert c1*c3 > c2**2, "Cauchy-Schwarz inequality is not satisfied."
+    
+    # boundary points
+    left = (1 - 1/c1)*Ao
+    right = (1/c1)*np.sqrt((c1-1)/(c1*c3 - c2**2))*(c1*As - c2*Ao)
+    Ab_low, Ab_high = left - right, left + right
+    
+    return Ab_low, Ab_high
+    
+
+def get_epsilon(Ao: list, bounds: tuple):
+    """ Returns the epsilon value for the boundary points.
+
+    Args:
+        Ao (list): Vector of observed amplitudes.
+        bounds (tuple): Tuple of boundary points.
+    """
+    pass
 
 
 def main():

### 2024-11-04 src/functions.py
i need obspy

@@ -32,9 +32,9 @@ def boundary_points(Ao: list, Uo: list, As: list = [], chi2: float = 1):
     Sigma_inv = linalg.inv(Sigma)
     
     # constants
-    c1 = Ao@Sigma_inv@Ao
-    c2 = Ao@Sigma_inv@As
-    c3 = As@Sigma_inv@As
+    c1 = Ao @ Sigma_inv @ Ao
+    c2 = Ao @ Sigma_inv @ As
+    c3 = As @ Sigma_inv @ As
     
     # conditions
     assert c1 >= 1, "COnfidence ellipsoid is too big."

### 2024-11-04 src/optimizer.py

@@ -1,0 +1,159 @@
+'''
+Name: Victor Agaba
+
+Date: 4th November 2024
+
+The goal of this module is to design optimization algorithms to be used
+for inversion of seismic data.
+'''
+
+
+import numpy as np
+from numpy import linalg
+
+
+def get_config():
+    '''
+    Get the default configuration for the optimization algorithm.
+    
+    Output:
+        config (dict): Default configuration for the optimization algorithm.
+        method (str): Optimization method to use: 'SD', 'Newton', 'BFGS'.
+        c_decrease (float): Sufficient decrease factor.
+        c_increase (float): Curvature condition factor.
+        k_max (int): Maximum number of iterations.
+        alpha (float): Initial step size.
+        rho (float): Backtracking line search parameter.
+        tolerance (float): Stopping criterion.
+        print_every (int): Print output every n iterations.
+    '''
+    config = {
+        'method': 'SD',
+        'c_decrease':1e-4,
+        'c_increase': 1-1e-4,
+        'k_max': 20000,
+        'alpha': 1,
+        'rho': 0.5,
+        'tolerance': 1e-6,
+        'print_every': 1
+    }
+    return config
+
+
+def minimize(objective, config, start=None):
+    '''
+    Minimize an objective function starting from a given point.
+    
+    Args:
+        function (callable): The objective function to minimize.
+        config (dict): Hyperparameters for the optimization.
+        start (list): Starting point for the optimization.
+    
+    Output:
+        x_k (list): The optimal point found by the optimization.
+    '''
+        
+    # extract config parameters
+    method = config['method']
+    c_decrease = config['c_decrease']
+    c_increase = config['c_increase']
+    k_max = config['k_max']
+    alpha_init = config['alpha']
+    rho = config['rho']
+    tolerance = config['tolerance']
+    print_every = config['print_every']
+    
+    # get initial guess
+    if start is None:
+        start = function.get_start()
+    x_k = start
+    
+    
+    # print output header
+    if len(x_k) > 5:
+        print(f'{method} optimization with backtracking linesearch from start = {x_k[:5]}...')
+    else:
+        print(f'{method} optimization with backtracking linesearch from start = {x_k}')
+    print(f'iter                f       ||p_k||       alpha     #func     ||grad_f||')
+    
[…]

### 2024-11-04 src/seismic_model.py

@@ -3,23 +3,22 @@ Name: Victor Agaba
 
 Date: 2nd November 2024
 
-The goal of this file is to create a model of the nonlinear inverse
+The goal of this module is to create a model of the nonlinear inverse
 problem to which we will apply the optimization algorithms.
 '''
 
 
 import numpy as np
 from numpy import linalg
-from matplotlib import pyplot as plt
 import functions as fn
 
 
-class InvProblem:
+class Model:
     '''
     Base class for an unconstrained optimzation problem.
     Catch anything whose method is not implemented.
     '''
-
+    
     def __call__(self, p):
         raise Exception('Class is not callable')
 
@@ -33,7 +32,7 @@ class InvProblem:
         raise Exception('Method "starting_point" not implemented')
 
 
-class SeismicModel(InvProblem):
+class SeismicModel(Model):
     '''
     This class computes the misfit function and its gradient
     for a given model.

### 2024-11-05 paper/main.tex
.
Finished bulk of clean codebase + heavy-duty algos

@@
 % import packages
-\usepackage{amssymb, nicefrac, bm, tikz}
+\usepackage{amssymb, nicefrac, bm, tikz, subfigure}
 
@@
       \item [i)]
-        Find estimates $\widehat{\psi}, \widehat{\delta}, \widehat{\lambda}$ for the strike, dip
-        and rake that \textit{best} describe the collected data.
+        Find estimates $\widehat{\psi}, \widehat{\delta}, \widehat{\lambda} \in [0^\circ,360^\circ) \times 
+        [0^\circ,90^\circ] \times [-180^\circ,180^\circ)$ for the strike, dip and rake that \textit{best} describe
+        the collected data.
 
@@
     \begin{align} \label{eq:3}
-      \Phi(\bm{p}) = \frac{-\bm{A}_o^\top\bm{A}_s}{\|\bm{A}_o\|\|\bm{A}_s\|}
+      \Phi(\bm{m}) = \frac{-\bm{A}_o^\top\bm{A}_s}{\|\bm{A}_o\|\|\bm{A}_s\|}
     \end{align}
@@
     (A_s^P, A_s^{SV}, A_s^{SH})^\top$ are synthetic amplitudes, and
-    $\bm{p} = (\psi, \delta, \lambda)^\top$ are the underlying parameters.
+    $\bm{m} = (\psi, \delta, \lambda)^\top$ are the underlying parameters.
 
@@
     \begin{align} \label{eq:4}
-      \Phi(\bm{p}) = \sqrt{\left( \Delta \frac{A^{P}}{A^{SV}}\right)^2 + \left( \Delta \frac{A^{SV}}{A^{SH}}\right)^2}
+      \Phi(\bm{m}) = \sqrt{\left( \Delta \frac{A^{P}}{A^{SV}}\right)^2 + \left( \Delta \frac{A^{SV}}{A^{SH}}\right)^2}
     \end{align}
@@
     \begin{align} \label{eq:5}
-      \Phi(\bm{p}) = \sqrt{\log^2 \left( \Delta \frac{A^P}{A^{SV}} \right) + \log^2 \left( \Delta \frac{A^{SV}}{A^{SH}} \right)}
+      \Phi(\bm{m}) = \sqrt{\log^2 \left( \Delta \frac{A^P}{A^{SV}} \right) + \log^2 \left( \Delta \frac{A^{SV}}{A^{SH}} \right)}
     \end{align}
@@
     \begin{align} \label{eq:6}
-      &\Phi(\bm{p}) \sim |\Sigma|^{-1/2} \exp \left( - \frac{1}{2} (\bm{A} - \bm{A}_o)^\top \bm{\Sigma}^{-1} (\bm{A} - \bm{A}_o) \right)
+      &\Phi(\bm{m}) \sim |\Sigma|^{-1/2} \exp \left( - \frac{1}{2} (\bm{A} - \bm{A}_o)^\top \bm{\Sigma}^{-1} (\bm{A} - \bm{A}_o) \right)
     \end{align}
@@
     \begin{align} \label{eq:7}
-      E(\bm{p}) = (\bm{A} - \bm{A}_o)^\top \tilde{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = 1
+      E(\bm{m}) = (\bm{A} - \bm{A}_o)^\top \tilde{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = 1
     \end{align}
@@
       \nonumber \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} ((1 - \nicefrac{1}{c_1})\bm{A}_o - \bm{A}_o)  = -1\\
-      \Rightarrow c_1 = \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_o = \|\bm{A}_o^\top, \bm{A}_o\|_{\tilde{\bm{\Sigma}}^{-1}}^2
+      \Rightarrow c_1 = \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_o = \|\bm{A}_o\|_{\tilde{\bm{\Sigma}}^{-1}}^2
     \end{align}
@@
 
-    Denote $c_3 = \bm{A}_s^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s^\top, \bm{A}_s\|_{\tilde{\bm{\Sigma}}^{-1}}^2$. Then by
+    Denote $c_3 = \bm{A}_s^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s\|_{\tilde{\bm{\Sigma}}^{-1}}^2$. Then by
     substitution we have
@@
     The vector $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ can be treated as a continuously differentiable 
-    nonlinear function of parameters $\bm{p} = (\psi, \delta, \lambda)^\top$, defined by equation 
+    nonlinear function of parameters $\bm{m} = (\psi, \delta, \lambda)^\top$, defined by equation 
     \ref{eq:1}. We can thus apply directed search methods to minimize the following cost function:
     \begin{align} \label{eq:17}
-      \Phi(\bm{p}) = -\hat{\bm{A}}_o^\top \hat{\bm{A}}
+      \Phi(\bm{m}) = -\hat{\bm{A}}_o^\top \hat{\bm{A}}
     \end{align}
@@
     To calculate the gradient, first note that $\bm{A}$ is a linear combination of nonlinear functions
-    of $\bm{p}$ (equation \ref{eq:2}).
+    of $\bm{m}$ (equation \ref{eq:2}).
 
@@
     \begin{align} \label{eq:18}
-      \Phi(\bm{p}) = \Phi(\bm{A}(\bm{\eta}(\bm{p})))
+      \Phi(\bm{m}) = \Phi(\bm{A}(\bm{\eta}(\bm{m})))
     \end{align}
@@
 
-    i) From $\bm{p}$ to $\bm{\eta}$, we can use equation \ref{eq:2} to define the Jacobian
+    i) From $\bm{m}$ to $\bm{\eta}$, we can use equation \ref{eq:2} to define the Jacobian
     as follows:
@@
       &\begin{bmatrix}
-        \frac{3\cos^2(i)-1}{\alpha^3}    &\frac{-\sin(2i)}{\alpha^3}    &\frac{-sin^2(i)}{\alpha^3}   &0    &0\\
-        \frac{1.5\sin(2j)}{\beta^3}      &\frac{\cos(2j)}{\beta^3}      &\frac{0.5\sin(2j)}{\beta^3}  &0    &0\\
-        0         &0           &0        &\frac{\cos(j)}{\beta^3}       &\frac{\sin(j)}{\beta^3}
+        \frac{3\cos^2(i)-1}{\alpha_h^3}    &\frac{-\sin(2i)}{\alpha_h^3}    &\frac{-sin^2(i)}{\alpha_h^3}   &0    &0\\
+        \frac{1.5\sin(2j)}{\beta_h^3}      &\frac{\cos(2j)}{\beta_h^3}      &\frac{0.5\sin(2j)}{\beta_h^3}  &0    &0\\
+        0         &0           &0        &\frac{\cos(j)}{\beta_h^3}       &\frac{\sin(j)}{\beta_h^3}
       \end{bmatrix}
@@
     \begin{align} \label{eq:21}
-      \nonumber\nabla_{\bm{A}}\Phi &= \frac{\|\bm{A}\|\nabla_{\bm{A}}(-\hat{\bm{A}}_o^\top\bm{A}) + 
-      \hat{\bm{A}}_o^\top \bm{A} \nabla_{\bm{A}}(\|\bm{A}\|)}{\|\bm{A}\|^2}\\
-            &= \frac{1}{\|\bm{A}\|}(\Phi\hat{\bm{A}} - \hat{\bm{A}}_o)
+      \nonumber\nabla_{\bm{A}}\Phi &= \frac{\|\bm{A}\|\nabla_{\bm{A}}(-\hat{\bm{A}}_o^\top\bm{A}) - 
+      (-\hat{\bm{A}}_o^\top \bm{A}) \nabla_{\bm{A}}(\|\bm{A}\|)}{\|\bm{A}\|^2}\\
+            &= \frac{-1}{\|\bm{A}\|}(\hat{\bm{A}}_o + \Phi\hat{\bm{A}})
     \end{align}
@@
     \begin{align} \label{eq:22}
-      \nabla_{\bm{p}}\Phi = \bm{J_p}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\Phi
+      \nabla_{\bm{m}}\Phi = \bm{J_p}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\Phi
     \end{align}
@@
   \subsection{Optimization algorithms}
-    Something about gradient-based methods. Try different algorithms, they may produce
-    interesting results >> IEMS 450-2.
-
-    \noindent \textbf{TODO:} Create one 3D image with the results of gradient descent from random points in space.
+    \begin{figure}[ht!]
+      \includegraphics[width=8.6cm]{seed_2024} 
+      \caption{Results of steepest descent algorithm after a systematic sweep of starting points.}
+      \label{fig:2}
+    \end{figure}
 
 \section{Uncertainty Quantification} \label{sec:uncertainty}
+
+    \subsection{Kagan angle distribution}
+      An alternative way of quantifying model deviations.
 
@@
       Why hybrid search was used, why systematic over random.
-
-    \subsection{Kagan angle}
-      An alternative way of quantifying model deviations.
 

### 2024-11-05 src/function_repo.py

@@ -1,1273 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-import scipy.stats as stats
-from obspy.taup import TauPyModel
-import collections
-
-from obspy.imaging.beachball import beachball
-from obspy.imaging.beachball import beach
-
-eps = 1e-10; halfpi = np.pi/2; twopi = 2*np.pi
-i_hat = np.array([1,0,0]); j_hat = np.array([0,1,0]); k_hat = np.array([0,0,1])
-np.random.seed(1029) # I hope this carries over to other files
-# It only does for randomness associated with fr
-
-def set_axes_equal(ax): # might be useful later
-    """
-    Make axes of 3D plot have equal scale so that spheres appear as spheres,
-    cubes as cubes, etc.
-
-    Input
-      ax: a matplotlib axis, e.g., as output from plt.gca().
-    """
-
-    x_limits = ax.get_xlim3d()
-    y_limits = ax.get_ylim3d()
-    z_limits = ax.get_zlim3d()
-
-    x_range = abs(x_limits[1] - x_limits[0])
-    x_middle = np.mean(x_limits)
-    y_range = abs(y_limits[1] - y_limits[0])
-    y_middle = np.mean(y_limits)
-    z_range = abs(z_limits[1] - z_limits[0])
-    z_middle = np.mean(z_limits)
-
-    # The plot bounding box is a sphere in the sense of the infinity
-    # norm, hence I call half the max range the plot radius.
-    plot_radius = 0.5*max([x_range, y_range, z_range])
-
-    ax.set_xlim3d([x_middle - plot_radius, x_middle + plot_radius])
-    ax.set_ylim3d([y_middle - plot_radius, y_middle + plot_radius])
-    ax.set_zlim3d([z_middle - plot_radius, z_middle + plot_radius])
-
-def uniform_samples(n: int, bounds: list) -> list:
-    """
-    Divide bounds into n uniformly distributed points
-
-    Args:
-        n (int): # divisions
-        bounds (list[float]): array bounds [start, end]
-        
-    Returns:
-        list: n-size array of floats
-    """
-    start = bounds[0]
-    end = bounds[1]
-    
-    return stats.uniform.rvs(loc=start, scale=end-start, size=n)
-
-def fill_square_randomly(n: int, bounds: list) -> list:
-    """
-    Fill 2D grid with n uniformly distributed points
-    Return array of coordinates w.r.t grid boundaries
[…]

### 2024-11-05 src/functions.py

@@ -1,40 +1,55 @@
-# every necessary function for the paper is defined here
+'''
+Name: Victor Agaba
+
+Date: 4th November 2024
+
+The goal of this module is to provide all helper functions needed for the
+seismic inversion project.
+
+'''
+
 
 import numpy as np
 from numpy import linalg
+import pandas as pd
+from matplotlib import pyplot as plt
+from obspy.imaging.beachball import beachball
+from obspy.imaging.beachball import beach
 
-tol = 1e-6
 
-def unit_vector(vector: list):
-    """ Returns the unit vector of a vector. """
+eps = 1e-10; halfpi = np.pi/2; twopi = 2*np.pi
+i_hat = np.array([1,0,0]); j_hat = np.array([0,1,0]); k_hat = np.array([0,0,1])
+
+
+def unit_vec(vector: list) -> list:
+    ''' Returns the unit vector of a vector. '''
     return vector / linalg.norm(vector)
 
 
-def boundary_points(Ao: list, Uo: list, As: list = [], chi2: float = 1):
-    """ Returns the boundary points of the model.
+def boundary_points(Ao: list, Uo: list, As: list = [], chi2: float = 1) -> tuple:
+    '''
+    Returns the boundary points of the model.
 
     Args:
         Ao (list): Vector of observed amplitudes.
         Uo (list): Vector of measured uncertainties.
         As (list, optional): Vector of simulated amplitudes. Defaults to [].
         chi2 (float, optional): _description_. Defaults to 1.
-    """
+    '''
+    # get a simulation for spherical model
     if len(As) == 0:
-        As = np.zeros(len(Ao))
-        As[0] = 1
-        if True: # TODO:...
-            As[1] = 1
-            
+        As = i_hat
+        if abs(Ao @ j_hat) < eps: As = As + j_hat 
     
     # adjusted covariance matrix
     assert all(Uo > 0), "All uncertainties must be positive."
-    Sigma = np.diag(Uo)/chi2
-    Sigma_inv = linalg.inv(Sigma)
+    Sigma_tilde = np.diag(Uo)/chi2
+    Sigma_tilde_inv = linalg.inv(Sigma_tilde)
     
     # constants
-    c1 = Ao @ Sigma_inv @ Ao
-    c2 = Ao @ Sigma_inv @ As
-    c3 = As @ Sigma_inv @ As
+    c1 = Ao @ Sigma_tilde_inv @ Ao
[…]

### 2024-11-05 src/minimize_objective.py

@@ -1,333 +1,0 @@
-#!/usr/bin/env python2
-# -*- coding: utf-8 -*-
-
-'''
-Name: Victor Agaba
-
-Date: 2023-20-11
-
-A module with 3 functions for implementing Newton method and gradient descent:
-    init_options()
-    minimize()
-    factorize_with_mod()
-'''
-
-# Import numpy for linear algebra computations
-import numpy as np
-
-# Import scipy to use in Cholesky factorization
-import scipy as sp
-
-
-def init_options():
-    '''
-    Option initialization for the gradient descent algorithm
-
-    Initialize algorithm options with default values
-
-    Return values:
-        options:
-            This is a dictionary with fields that correspond to algorithmic
-            options of our method.  In particular:
-
-            max_iter:
-                Maximum number of iterations
-            tol:
-                Convergence tolerance
-            step_type:
-                Different ways to calculate the search direction:
-                    'gradient_descent'
-                    'Newton'
-            linesearch:
-                Type of line search:
-                    'constant'
-                    'backtracking'
-            alpha_init:
-                First trial step size to attempt during line search:
-                    'constant':     Value of the constant step size
-                    'backtracking': Initial trial step size
-            suff_decrease_factor:
-                Coefficient in sufficient decrease condition for backtracking
-                    line search
-            perturb_init:
-                Initial perturbation of Hessian matrix to ensure positive
-                    definiteness
-            perturb_inc_factor:
-                Increase factor for Hessian perturbation
-            output_level:
-                Amount of output to be printed
-                0: No output
-                1: Only summary information
-                2: One line per iteration (good for debugging)
-                3: Print iterates in each iteration
-    '''
-    options = {}
-    options['max_iter'] = 1000
-    options['tol'] = 1e-6
-    options['step_type'] = 'gradient_descent'
-    options['linesearch'] = 'backtracking'
-    options['alpha_init'] = 1
-    options['suff_decrease_factor'] = 1e-4
-    options['perturb_init'] = 1e-4
-    options['perturb_inc_factor'] = 10
-    options['output_level'] = 1
-
-    return options
-
-
-def minimize(opt_prob, options, x_start=None):
-    '''
-    Optimization methods for unconstrainted optimization
-
-    This is an implementation of the gradient descent method for unconstrained
[…]

### 2024-11-05 src/optimizer.py

@@ -18,7 +18,7 @@ def get_config():
     
     Output:
         config (dict): Default configuration for the optimization algorithm.
-        method (str): Optimization method to use: 'SD', 'Newton', 'BFGS'.
+        method (str): Optimization method to use: 'SD' or 'Newton'.
         c_decrease (float): Sufficient decrease factor.
         c_increase (float): Curvature condition factor.
         k_max (int): Maximum number of iterations.
@@ -30,9 +30,8 @@ def get_config():
     config = {
         'method': 'SD',
         'c_decrease':1e-4,
-        'c_increase': 1-1e-4,
-        'k_max': 20000,
-        'alpha': 1,
+        'k_max': 1000,
+        'alpha': 1e-2,
         'rho': 0.5,
         'tolerance': 1e-6,
         'print_every': 1
@@ -56,7 +55,6 @@ def minimize(objective, config, start=None):
     # extract config parameters
     method = config['method']
     c_decrease = config['c_decrease']
-    c_increase = config['c_increase']
     k_max = config['k_max']
     alpha_init = config['alpha']
     rho = config['rho']
@@ -70,11 +68,12 @@ def minimize(objective, config, start=None):
     
     
     # print output header
-    if len(x_k) > 5:
-        print(f'{method} optimization with backtracking linesearch from start = {x_k[:5]}...')
-    else:
-        print(f'{method} optimization with backtracking linesearch from start = {x_k}')
-    print(f'iter                f       ||p_k||       alpha     #func     ||grad_f||')
+    if print_every > 0:
+        if len(x_k) > 5:
+            print(f'{method} optimization with backtracking linesearch from start = {x_k[:5]}...')
+        else:
+            print(f'{method} optimization with backtracking linesearch from start = {x_k}')
+        print(f'iter                f       ||p_k||       alpha     #func     ||grad_f||')
     
     # evaluate function at start for later comparison
     f_k = objective(x_k)
@@ -84,13 +83,14 @@ def minimize(objective, config, start=None):
     k = 1
     
     # optimization loop
-    while k < k_max:
+    converged = False
+    while not converged:
         
         # reset hyperparameters for line iteration
         num_calls = 0
         alpha = alpha_init
 
-        # compute gradient at x_k, cuts across all methods       
+        # compute gradient at x_k, cuts across all methods
         grad_k = objective.gradient(x_k)
[…]

### 2024-11-05 src/seismic_model.py

@@ -10,6 +10,8 @@ problem to which we will apply the optimization algorithms.
 
 import numpy as np
 from numpy import linalg
+from matplotlib import pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
 import functions as fn
 
 
@@ -19,13 +21,13 @@ class Model:
     Catch anything whose method is not implemented.
     '''
     
-    def __call__(self, p):
+    def __call__(self, params):
         raise Exception('Class is not callable')
 
-    def gradient(self, p):
+    def gradient(self, params):
         raise Exception('Method "gradient" is not implemented')
 
-    def hessian(self, p):
+    def hessian(self, params):
         raise Exception('Method "hessian" not is implemented')
 
     def starting_point(self):
@@ -38,10 +40,10 @@ class SeismicModel(Model):
     for a given model.
     Uses the same notation as in the paper.
     '''
-    def __init__(self, phi, i, j, alpha, beta, Ao):
+    def __init__(self, azimuth, takeoff_angles, velocities, Ao = []):
         '''
         Initialize with the parts of the model that don't change.
-        ALL ANGLES ARE IN RADIANS!!!
+        All angles are in RADIANS!!!
         
         Args:
         phi = raypath's azimuth
@@ -49,12 +51,17 @@ class SeismicModel(Model):
         alpha, beta = P and S wave velocities
         Ao = observed amplitudes
         '''
+        # extract inputs
+        i, j = takeoff_angles
+        i, j = np.deg2rad(i), np.deg2rad(j)
+        alpha_h, beta_h = velocities
+        
         # initialize misfits iterates for tracking
         self.misfits = []
         self.iterates = []
         
         # initialize constant Jacobian matrix
-        factors = np.array([alpha, beta, beta])**3
+        factors = np.array([alpha_h, beta_h, beta_h])**3
         self.J_A = np.array(
             [[(3*np.cos(i)**2 - 1), -np.sin(2*i), -np.sin(i)**2, 0, 0],
              [1.5*np.sin(2*j), np.cos(2*j), 0.5*np.sin(2*j), 0, 0],
@@ -68,14 +75,14 @@ class SeismicModel(Model):
         self.nabla_Phi = np.zeros(3)
         
         # store other inputs for later use
-        self.phi = phi
+        self.phi = azimuth
         self.Ao = Ao
     
     
[…]

### 2024-11-28 src/functions.py
Breakthrough!

@@ -255,6 +255,41 @@ def tp2sdr(t: list, p: list, deg: bool = False) -> tuple:
     return (np.array([strike1, dip1, rake1]), np.array([strike2, dip2, rake2]))
 
 
+def sdr2tp(sdr: list, deg: bool = False) -> tuple:
+    """
+    Converts from strike, dip, and rake to T and P axes.
+
+    Args:
+        sdr (list): strike, dip, and rake
+        deg (bool, optional): input in degrees if True
+
+    Returns:
+        tuple: (t, p) as unit vectors in x, y, z
+    """
+    if deg: sdr = np.deg2rad(sdr)
+    
+    # first normal from strike and dip
+    strike, dip, rake = sdr
+    
+    n1_theta = bearing2angle(strike + halfpi)
+    n1_phi = dip
+    n1 = pol2rect(np.array([1, n1_theta, n1_phi]))
+
+    n2_theta = bearing2angle(strike)
+    n2_init = pol2rect(np.array([1, n2_theta, halfpi]))
+    n2 = rotate_vec(n2_init, n1, rake)
+    
+    # get T and P axes
+    t = (n1 + n2)/np.linalg.norm(n1 + n2)
+    p = (n1 - n2)/np.linalg.norm(n1 - n2)
+    
+    # restrict to upper hemisphere
+    if t[2] < 0: t *= -1
+    if p[2] < 0: p *= -1
+    
+    return t, p
+
+
 def bound(params: list) -> list:
     '''
     Put params in the range [0, 2pi], [0, pi/2], [-pi, pi].
@@ -288,7 +323,7 @@ def bound(params: list) -> list:
     return out
 
 
-def rigid_hemisphere_samples(step_size: float) -> list:
+def systematic_hemisphere_samples(step_size: float) -> list:
     """
     Systematically samples n samples evenly spaced on an upper hemisphere surface.
 
@@ -320,7 +355,7 @@ def systematic_params(dd: float) -> list:
         dd (float): angular distance between samples in DEGREES
     '''
     params = []
-    Ts = rigid_hemisphere_samples(dd)
+    Ts = systematic_hemisphere_samples(dd)
     P_rotations = np.arange(0, np.pi, np.deg2rad(dd))
     for T in Ts:
         P_start = starting_direc(T, i_hat + j_hat + k_hat)
@@ -332,16 +367,68 @@ def systematic_params(dd: float) -> list:
     return np.array(params)
 
 
+def orthogonalize(v: list, u: list) -> list:
+    '''
[…]

### 2024-11-28 src/graveyard.py

@@ -1,0 +1,20 @@
+import functions as fn
+import numpy as np
+
+start0 = np.zeros(3)
+start1 = np.ones(3)
+others = []
+for _ in range(4):
+    # create random array for t, normalised to 1
+    t = fn.unit_vec(np.random.rand(3))
+
+    # create random array for p, normalised to 1
+    direc = fn.unit_vec(np.random.rand(3))
+    p = fn.starting_direc(t, direc)
+
+    # get true params for synthetic test
+    add2start = fn.tp2sdr(t, p)[0]
+    others.append(add2start)
+
+starts = [start0, start1]
+starts.extend(others)

### 2024-11-28 src/optimizer.py

@@ -66,7 +66,6 @@ def minimize(objective, config, start=None):
         start = function.get_start()
     x_k = start
     
-    
     # print output header
     if print_every > 0:
         if len(x_k) > 5:
@@ -89,7 +88,7 @@ def minimize(objective, config, start=None):
         # reset hyperparameters for line iteration
         num_calls = 0
         alpha = alpha_init
-
+        
         # compute gradient at x_k, cuts across all methods
         grad_k = objective.gradient(x_k)
         
@@ -108,18 +107,17 @@ def minimize(objective, config, start=None):
             return x_k
         
         # perform line search to satisfy Wolfe conditions
-        f_k1 = objective(x_k + alpha*p_k)
+        x_k1 = x_k + alpha*p_k
+        f_k1 = objective(x_k1)
         num_calls += 1
         upper_boundary = f_k + c_decrease*alpha*(grad_k @ p_k)
         
         while f_k1 > upper_boundary:
             alpha *= rho
-            f_k1 = objective(x_k + alpha*p_k)
+            x_k1 = x_k + alpha*p_k
+            f_k1 = objective(x_k1)
             num_calls += 1
             upper_boundary = f_k + c_decrease*alpha*(grad_k @ p_k)
-        
-        # make the step
-        x_k1 = x_k + alpha*p_k
         
         # print iteration output
         max_grad_k = linalg.norm(grad_k, ord=np.inf)
@@ -127,7 +125,7 @@ def minimize(objective, config, start=None):
             norm_p_k = linalg.norm(p_k)
             print(f'{k:<5d}     {f_k:<+5.4e}      {norm_p_k:<5.2e}    {alpha:<5.2e}         {num_calls:<5d}   {max_grad_k:<5.2e}')
         
-        # update variables
+        # make the step
         x_k = x_k1
         f_k = f_k1
         k += 1
@@ -144,5 +142,9 @@ def minimize(objective, config, start=None):
                 print('Maximum number of iterations reached!')
                 
             break
+    
+    if converged:
+        objective.set_optimal_iterate(x_k)
+        objective.set_optimal_amplitude()
     
     return x_k, f_k, converged

### 2024-11-28 src/other/beach_plot.py

@@ -1,0 +1,139 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+from obspy.core.trace import Trace, Stats
+from obspy.core.stream import Stream
+from obspy import UTCDateTime
+from obspy.clients.fdsn import Client
+
+from obspy.imaging.beachball import beachball
+from obspy.imaging.beachball import beach
+from obspy.imaging.source import plot_radiation_pattern
+
+import argparse
+from constants import DATA_PATH, EXP_PATH
+import logging 
+from datetime import datetime 
+from pathlib import Path
+import utils 
+import warnings 
+import sys 
+import os 
+
+
+def alphabeach(event, df, n, norm_const, c='b'):
+    fig1, ax1 = plt.subplots(subplot_kw={'aspect': 'equal'})
+    fig2, ax2 = plt.subplots(subplot_kw={'aspect': 'equal'})
+    fig3, ax3 = plt.subplots(subplot_kw={'aspect': 'equal'})
+
+    # hide axes + ticks
+    ax1.axison = False
+    ax2.axison = False
+    ax3.axison = False 
+
+    max_mis = df['Misfit'].max()
+    min_mis = df['Misfit'].min()    
+
+    # Max-min normalization 
+    # df['Normalized'] = (max_mis - df['Misfit'])/(max_mis-min_mis)
+    
+    # Gaussian normalization 
+    sigma = max_mis 
+    # pref = 1/(sigma*np.sqrt(2*np.pi)) 
+    df['Normalized'] = norm_const*np.exp(-(np.square(df['Misfit']))/(2*np.square(sigma)))
+
+    # Plot complete solution set, with n figures in each plot 
+    counter = 1
+    num_figs = 0
+
+    for index, rows in df.iterrows():
+        f = [rows.Strike, rows.Dip, rows.Rake]
+        x = 210 * ((counter-1) % 10)
+        y = 210 * ((counter-1) // 10)
+
+        # Collection1 is for individual solutions 
+        collection1  = beach(f, xy=(x, y), facecolor=c, alpha=rows.Normalized)
+        ax1.add_collection(collection1)
+
+        # Collection2 is for superimposed solutions 
+        collection2 = beach(f, xy=(0, 0), facecolor=c, alpha=rows.Normalized)
+        ax2.add_collection(collection2)
+
+        # Plot set of n solutions per plot 
+        if counter % n == 0 and index not in (0, len(df)-1): 
+            ax1.autoscale_view(tight=False, scalex=True, scaley=True)
[…]

### 2024-11-28 src/other/victor_kagan_angles.py

@@ -1,0 +1,377 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import math
+import pandas as pd
+#### define all the relevant functions #########
+
+def F4R1(q1,q2,icode):
+    """
+    q=q2*(i,j,k,1))/q1 
+    first compute q2*(i,j,k,1) by using boxtest
+    second, get q usin quatD which divides two quaternions
+    """
+    ### compute q2*(i,j,k,1) bu calling boxtest
+    qr2=my_BOXTEST(q2,icode)
+    q=QUATD(q1,qr2)  #first element is the denominator and the second is the numerator
+    qp=QUATP(q1,q)
+    #print(f"q, {q}")
+    rotation_angle,colatitude,azimuth=SPHCOOR(q)
+    #print("anle,theta, phi")
+    #print(rotation_angle, colatitude, azimuth)
+    return q
+
+def my_BOXTEST(q1,icode):
+    """
+    for Icode=0, finds the minimal rotation quaternion
+    for Icode=N, N=1,3, finds the rotation quaternion q2=q1*(i,j,k,1)
+    Note my code is for a quaternion q3*k+q2*j+q1*i+q0
+    """
+    q2 = np.zeros(4)
+    quatt = np.zeros(4)
+    quat=np.eye(4)
+    q2=np.array([q1[i] for i in range(4)])
+    if icode==0:
+        #print("minimum rotation quaternion")
+        quatt=quat[:,3]
+    elif icode==1:
+        quatt=quat[:,0]  #multiplication by i
+    elif icode==2: 
+        quatt=quat[:,1]  # multiplication by j
+    elif icode==3: 
+        quatt=quat[:,2]  #multiplication by k
+    #print(quatt)
+    q2=QUATP(quatt,q1)
+    if q2[-1]<0:
+        q2=np.array([-q2[i] for i in range(4)])
+    #if icode==0:
+     #   print(f"Qmin={q2}")
+    return q2
+
+def SPHCOOR(quat):
+    """
+    for the rotation quaternion quat, this function finds the rotation angle( angl) 
+    of a counterclockwise rotation and spehrical coordinates (colatitude, theta, and azimuth)
[…]

### 2024-11-28 src/seismic_model.py

@@ -56,9 +56,16 @@ class SeismicModel(Model):
         i, j = np.deg2rad(i), np.deg2rad(j)
         alpha_h, beta_h = velocities
         
-        # initialize misfits iterates for tracking
+        # initialize misfits, iterates and optimals for tracking
         self.misfits = []
         self.iterates = []
+        self.optimal_iterates = []
+        self.misfits = []
+        self.tp_axes = []
+        self.amplitudes = []
+        self.optimal_amplitudes = []
+        
+        # TODO: get convergence rate built in
         
         # initialize constant Jacobian matrix
         factors = np.array([alpha_h, beta_h, beta_h])**3
@@ -136,6 +143,7 @@ class SeismicModel(Model):
         if store:
             self.misfits.append(Phi)
             self.iterates.append(params)
+            self.amplitudes.append(self.A)
         
         return grad
     
@@ -169,10 +177,52 @@ class SeismicModel(Model):
         ''' Return misfit values per iterate from optimizaiton. '''
         return self.misfits
     
+    
+    def set_optimal_iterate(self, params):
+        ''' Set the optimal parameters. '''
+        self.optimal_iterates.append(params)
+    
+    
+    def set_optimal_amplitude(self):
+        ''' Set the amplitude vector. '''
+        self.optimal_amplitudes.append(self.A)
+        
+    
+    def get_optimal_iterates(self):
+        ''' Return the optimal parameters. '''
+        return self.optimal_iterates
+    
+    
+    def get_optimal_amplitudes(self):
+        ''' Return the optimal amplitudes. '''
+        return self.optimal_amplitudes
+    
+    
+    def get_tp_axes(self):
+        ''' Return the tp axes. '''
+        return self.tp_axes
+    
+    
+    def filter_outliers(self, threshold=2):
+        ''' Filter out outliers based on Z-score of optimal amplitude norms. '''
+        norms = np.array([linalg.norm(a) for a in self.optimal_amplitudes])
+        z_scores = np.abs((norms - np.mean(norms)) / np.std(norms))
+        self.optimal_iterates = [m for i, m in enumerate(self.optimal_iterates) 
+                                 if z_scores[i] < threshold]
+        self.optimal_amplitudes = [a for i, a in enumerate(self.optimal_amplitudes) 
+                                   if z_scores[i] < threshold]
+    
+    
     def reset(self):
[…]

### 2024-11-28 src/functions.py
LstSq

@@ -91,7 +91,7 @@ def starting_direc(point: list, direc: list) -> list:
     
     proj = (np.dot(direc, point)/np.dot(point, point))
     rem = np.array(direc) - proj*np.array(point)
-    return rem/np.linalg.norm(rem) if np.linalg.norm(rem) > eps else j_hat
+    return rem/linalg.norm(rem) if linalg.norm(rem) > eps else j_hat
 
 
 def rect2pol(rect: list) -> list:
@@ -104,12 +104,12 @@ def rect2pol(rect: list) -> list:
     
     if len(rect) == 2:
         x, y = rect
-        r = np.linalg.norm(rect)
+        r = linalg.norm(rect)
         theta = np.arctan2(y, x) % twopi
         return np.array([r, theta])
     
     x, y, z = rect
-    rho = np.linalg.norm(rect)
+    rho = linalg.norm(rect)
     theta = np.arctan2(y, x) % twopi
     phi = np.arctan2(np.sqrt(x**2 + y**2), z)
     
@@ -241,10 +241,10 @@ def tp2sdr(t: list, p: list, deg: bool = False) -> tuple:
 
     # get rakes from  base and slip vectors
     if not done1:
-        rake1 = np.arccos(np.dot(slip1, base1)/np.linalg.norm(slip1))
+        rake1 = np.arccos(np.dot(slip1, base1)/linalg.norm(slip1))
         rake1 *= np.sign(slip1[2])
     if not done2:
-        rake2 = np.arccos(np.dot(slip2, base2)/np.linalg.norm(slip2))
+        rake2 = np.arccos(np.dot(slip2, base2)/linalg.norm(slip2))
         rake2 *= np.sign(slip2[2])
     
     if deg:
@@ -280,8 +280,8 @@ def sdr2tp(sdr: list, deg: bool = False) -> tuple:
     n2 = rotate_vec(n2_init, n1, rake)
     
     # get T and P axes
-    t = (n1 + n2)/np.linalg.norm(n1 + n2)
-    p = (n1 - n2)/np.linalg.norm(n1 - n2)
+    t = (n1 + n2)/linalg.norm(n1 + n2)
+    p = (n1 - n2)/linalg.norm(n1 - n2)
     
     # restrict to upper hemisphere
     if t[2] < 0: t *= -1
@@ -375,8 +375,8 @@ def orthogonalize(v: list, u: list) -> list:
         v (list): vector to orthogonalize
[…]

### 2024-11-28 src/seismic_model.py

@@ -222,18 +222,6 @@ class SeismicModel(Model):
         self.tp_axes = []
         self.amplitudes = []
         self.optimal_amplitudes = []
-    
-    
-    def plot_misfit(self):
-        '''
-        Plot the misfits in a subplot.
-        '''
-        fig, ax = plt.subplots(1, 1, figsize=(6, 5))
-        ax.plot(self.misfits)
-        ax.set_title('Misfit function')
-        ax.set_xlabel('Iteration')
-        ax.set_ylabel('Misfit')
-        plt.show()
         
     
     def mirror(self, what, index=0):
@@ -255,9 +243,11 @@ class SeismicModel(Model):
                     iterates.append(fn.tp2sdr(t, p)[index])
                 self.iterates = iterates
             elif what == 'axes':
+                tp_axes = []
                 for m in self.optimal_iterates:
                     t, p = fn.sdr2tp(m)
-                    self.tp_axes.append([t, p])
+                    tp_axes.append([t, p])
+                self.tp_axes = fn.filter_axes(tp_axes)
             else:
                 raise ValueError('Invalid keyword for "what"')
         elif len(what) > 1:
@@ -267,6 +257,18 @@ class SeismicModel(Model):
             raise ValueError('Invalid argument for "what"')
     
     
+    def plot_misfit(self):
+        '''
+        Plot the misfits in a subplot.
+        '''
+        fig, ax = plt.subplots(1, 1, figsize=(6, 5))
+        ax.plot(self.misfits)
+        ax.set_title('Misfit function')
+        ax.set_xlabel('Iteration')
+        ax.set_ylabel('Misfit')
+        plt.show()
+        
+        
     def plot_iterates_2D(self, cmap='rainbow', s=10, optimal=True, index=2):
         '''
         Make 3 2D plots of the iterates: psi, delta, lambda.
@@ -316,7 +318,6 @@ class SeismicModel(Model):
         opt_strikes = [np.rad2deg(m[0]) for m in optimal_iterates]
         opt_dips = [np.rad2deg(m[1]) for m in optimal_iterates]
         opt_rakes = [np.rad2deg(m[2]) for m in optimal_iterates]
-        
         
         # make the plots
         fig, (ax1, ax2, ax3) = plt.subplots(1, 3, figsize=(18, 5))
@@ -487,12 +488,14 @@ class SeismicModel(Model):
         Make a 3D plot of the optimal tp axes.
         '''
[…]

### 2024-11-29 src/beachball.py
Parameterization algorithm complete

@@ -1,0 +1,9 @@
+'''
+Name: Victor Agaba
+
+Date: 28th November 2024
+
+The goal of this module is to provide all helper functions needed for
+beachball plotting.
+'''
+

### 2024-11-29 src/functions.py

@@ -483,14 +483,15 @@ def regression_plane(data: list) -> tuple:
         data (list): list of data points
     '''
     centroid = np.mean(data, axis=0)
-    A = np.c_[data, np.ones(data.shape[0])]
-    print(f'Rank of A: {linalg.matrix_rank(A)}')
-    # TODO: why is this not working?
-    normal = np.lin
-    normal = normal[:-1]
-    print(f"Centroid: {centroid}, Normal: {normal}")
+    A = np.c_[data[:,:-1], np.ones(data.shape[0])]
+    b = data[:,-1]
+    x = linalg.solve(A.T @ A, A.T @ b)
+    normal = unit_vec(np.array([x[0], x[1], -1]))
+    if np.dot(normal, centroid) < 0: normal *= -1
+    normal = unit_vec(normal)
     
     return centroid, normal
+
 
 def regression_axes(tp_axes: list) -> tuple:
     '''
@@ -511,8 +512,8 @@ def regression_axes(tp_axes: list) -> tuple:
     T_try, T_normal = unit_vec(T_centroid), unit_vec(T_normal)
     P_try, P_normal = unit_vec(P_centroid), unit_vec(P_normal)
     
-    # return Centroid and normal whose dot product is largest
+    # return centroid and normal whose dot product is largest
     if T_try @ T_normal > P_try @ P_normal:
-        return T_centroid, T_normal
+        return T_centroid, T_normal, 0
     else:
-        return P_centroid, P_normal
+        return P_centroid, P_normal, 1

### 2024-11-29 src/optimizer.py

@@ -30,11 +30,11 @@ def get_config():
     config = {
         'method': 'SD',
         'c_decrease':1e-4,
-        'k_max': 1000,
-        'alpha': 1e-2,
+        'k_max': 3000,
+        'alpha': 5e-2,
         'rho': 0.5,
-        'tolerance': 1e-6,
-        'print_every': 1
+        'tolerance': 1e-5,
+        'print_every': 0
     }
     return config
 
@@ -146,5 +146,7 @@ def minimize(objective, config, start=None):
     if converged:
         objective.set_optimal_iterate(x_k)
         objective.set_optimal_amplitude()
+        
+    objective.update_convergence(converged)
     
-    return x_k, f_k, converged
+    return x_k

### 2024-11-29 src/seismic_model.py

@@ -64,8 +64,8 @@ class SeismicModel(Model):
         self.tp_axes = []
         self.amplitudes = []
         self.optimal_amplitudes = []
-        
-        # TODO: get convergence rate built in
+        self.half_angles = []
+        self.runs, self.converged = 0, 0
         
         # initialize constant Jacobian matrix
         factors = np.array([alpha_h, beta_h, beta_h])**3
@@ -168,6 +168,17 @@ class SeismicModel(Model):
         self.Ao = Ao
     
     
+    def update_convergence(self, converged):
+        ''' Update the number of runs and the number of convergences. '''
+        self.runs += 1
+        self.converged += converged
+    
+    
+    def get_convergence_rate(self):
+        ''' Return the convergence rate. '''
+        return 100*self.converged/self.runs
+    
+    
     def get_iterates(self):
         ''' Return parameter iterates from optimization. '''
         return self.iterates
@@ -203,6 +214,11 @@ class SeismicModel(Model):
         return self.tp_axes
     
     
+    def get_half_angles(self):
+        ''' Return the half angles. '''
+        return self.half_angles
+    
+    
     def filter_outliers(self, threshold=2):
         ''' Filter out outliers based on Z-score of optimal amplitude norms. '''
         norms = np.array([linalg.norm(a) for a in self.optimal_amplitudes])
@@ -222,6 +238,8 @@ class SeismicModel(Model):
         self.tp_axes = []
         self.amplitudes = []
         self.optimal_amplitudes = []
+        self.half_angles = []
+        self.runs, self.converged = 0, 0
         
     
     def mirror(self, what, index=0):
@@ -268,6 +286,18 @@ class SeismicModel(Model):
         ax.set_ylabel('Misfit')
         plt.show()
         
+    
+    def plot_half_angles(self, bins=20):
+        '''
+        Plot the half angles in a subplot.
+        '''
+        fig, ax = plt.subplots(1, 1, figsize=(6, 5))
+        ax.hist(np.rad2deg(self.half_angles), bins=bins)
+        ax.set_title('Half-angles diagnostic histogram')
+        ax.set_xlabel('Half angle (deg)')
+        ax.set_ylabel('Frequency')
+        plt.show()
+        
         
     def plot_iterates_2D(self, cmap='rainbow', s=10, optimal=True, index=2):
         '''
@@ -279,7 +309,6 @@ class SeismicModel(Model):
[…]

### 2024-11-29 src/seismic_model.py
To be contunued...

@@ -174,7 +174,7 @@ class SeismicModel(Model):
         self.converged += converged
     
     
-    def get_convergence_rate(self):
+    def convergence_rate(self):
         ''' Return the convergence rate. '''
         return 100*self.converged/self.runs
     
@@ -293,7 +293,7 @@ class SeismicModel(Model):
         '''
         fig, ax = plt.subplots(1, 1, figsize=(6, 5))
         ax.hist(np.rad2deg(self.half_angles), bins=bins)
-        ax.set_title('Half-angles diagnostic histogram')
+        ax.set_title('Diagnostic histogram of half-angles')
         ax.set_xlabel('Half angle (deg)')
         ax.set_ylabel('Frequency')
         plt.show()
@@ -520,7 +520,7 @@ class SeismicModel(Model):
         '''
         if len(self.tp_axes) == 0: self.mirror(['axes'])
         zero = np.zeros(3)
-        _, normal, _ = fn.regression_axes(self.tp_axes)
+        _, c, _ = fn.regression_axes(self.tp_axes)
         
         fig = plt.figure(figsize=(15, 10))
         ax = fig.add_subplot(111, projection='3d')
@@ -542,7 +542,9 @@ class SeismicModel(Model):
             ax.plot([p[0], p_prime[0]], [p[1], p_prime[1]], [p[2], p_prime[2]],
                     c='red', alpha=0.5)
         
-        ax.plot([normal[0], zero[0]], [normal[1], zero[1]], [normal[2], zero[2]],
+        if half: c_prime = zero
+        else: c_prime = -c
+        ax.plot([c[0], c_prime[0]], [c[1], c_prime[1]], [c[2], c_prime[2]],
                 c='green', alpha=0.5, label="central", linewidth=3)
         
         ax.set_xlabel('x')

### 2024-11-29 src/graveyard.py
.

@@ -1,20 +1,0 @@
-import functions as fn
-import numpy as np
-
-start0 = np.zeros(3)
-start1 = np.ones(3)
-others = []
-for _ in range(4):
-    # create random array for t, normalised to 1
-    t = fn.unit_vec(np.random.rand(3))
-
-    # create random array for p, normalised to 1
-    direc = fn.unit_vec(np.random.rand(3))
-    p = fn.starting_direc(t, direc)
-
-    # get true params for synthetic test
-    add2start = fn.tp2sdr(t, p)[0]
-    others.append(add2start)
-
-starts = [start0, start1]
-starts.extend(others)

### 2024-12-01 src/3D.py
Taking another break, the work is done

@@ -1,0 +1,65 @@
+import numpy as np
+from numpy import linalg
+import functions as fn
+import seismic_model as sm
+import optimizer as opt
+from matplotlib import pyplot as plt
+from obspy.taup import TauPyModel
+import importlib
+
+np.random.seed(2024)
+# np.random.seed(np.random.randint(0, 2024))
+
+# create random array for t, normalised to 1
+t = fn.unit_vec(np.random.rand(3))
+
+# create random array for p, normalised to 1
+direc = fn.unit_vec(np.random.rand(3))
+p = fn.starting_direc(t, direc)
+
+# get truen params for synthetic test
+true_params = fn.tp2sdr(t, p, True)[0]
+print('Original params: ', true_params)
+
+importlib.reload(fn)
+importlib.reload(sm)
+importlib.reload(opt)
+
+# set up parameters for inversion
+model = TauPyModel(model='ak135')  # velocity model
+hdepth = 15  # km - assumed quake depth
+epdist = 10  # degrees - epicentral distance
+azimuth = 200  # degrees - azimuth of seismometer
+arrivals = model.get_travel_times(source_depth_in_km=hdepth,
+                        distance_in_degree=epdist, phase_list=['P', 'S'])
+takeoff_angles = [a.takeoff_angle for a in arrivals]
+velocities = np.array([5.8000, 3.4600])
+
+# initialize model
+inv_model = sm.SeismicModel(np.deg2rad(azimuth), takeoff_angles, velocities)
+best_fit = inv_model(np.deg2rad(true_params), set_Ao=True)
+
+
+
+######################
+
+
+Ao = inv_model.get_Ao()
+inv_model.set_Uo()
+Uo = inv_model.get_Uo()
+chi2 = inv_model.get_chi2()
+normal = inv_model.get_normal()
+dd = 15
+n = 50 # enough to cover a half-revolution
+amps = fn.systematic_elliptical_cone(normal, Ao, Uo, chi2, dd, n)
+
+# plot amps in 3D space
+fig = plt.figure()
+ax = fig.add_subplot(111, projection='3d')
+# ax.plot([0, normal[0]], [0, normal[1]], [0, normal[2]], c='b')
+ax.scatter(amps[0], amps[1], amps[2], c='r', marker='o')
+ax.set_xlabel('AP')
+ax.set_ylabel('ASV')
+ax.set_zlabel('ASH')
+ax.set_title('Simulated amplitude data')
+plt.show()

### 2024-12-01 src/functions.py

@@ -26,7 +26,7 @@ def unit_vec(vector: list) -> list:
     return vector / linalg.norm(vector)
 
 
-def boundary_points(Ao: list, Uo: list, As: list = [], chi2: float = 1) -> tuple:
+def boundary_points(Ao: list, Uo: list, As: list, chi2: float) -> tuple:
     '''
     Returns the boundary points of the model.
 
@@ -36,10 +36,6 @@ def boundary_points(Ao: list, Uo: list, As: list = [], chi2: float = 1) -> tuple
         As (list, optional): Vector of simulated amplitudes. Defaults to [].
         chi2 (float, optional): _description_. Defaults to 1.
     '''
-    # get a simulation for spherical model
-    if len(As) == 0:
-        As = i_hat
-        if abs(Ao @ j_hat) < eps: As = As + j_hat 
     
     # adjusted covariance matrix
     assert all(Uo > 0), "All uncertainties must be positive."
@@ -52,7 +48,7 @@ def boundary_points(Ao: list, Uo: list, As: list = [], chi2: float = 1) -> tuple
     c3 = As @ Sigma_tilde_inv @ As
     
     # conditions
-    assert c1 >= 1, "COnfidence ellipsoid is too big."
+    assert c1 >= 1, f"Confidence ellipsoid is too big: c1 = {c1}"
     assert c1*c3 > c2**2, "Cauchy-Schwarz inequality is not satisfied."
     
     # boundary points
@@ -81,7 +77,7 @@ def get_epsilon(Ao: list, bounds: tuple) -> float:
 
 def starting_direc(point: list, direc: list) -> list:
     '''
-    Find what vector is perpendicular to point vector and lies in the
+    Find what unit vector is perpendicular to point vector and lies in the
     plane that contains point and direc.
 
     Args:
@@ -174,7 +170,7 @@ def rotate_vec(vec: list, axis: list, theta: float) -> list:
     Axis must be a unit vector.
     """
     rotated_vec = vec*np.cos(theta) + (np.cross(axis,vec))*np.sin(theta) + \
-        axis*(np.dot(axis,vec))*(1-np.cos(theta))
+        axis*(axis @ vec)*(1-np.cos(theta))
     
     return rotated_vec
 
@@ -219,17 +215,17 @@ def tp2sdr(t: list, p: list, deg: bool = False) -> tuple:
         if abs(np.dot(check1, t)) < eps: rake1 = 0.
         else: rake1 = np.pi
[…]

### 2024-12-01 src/optimizer.py

@@ -150,3 +150,4 @@ def minimize(objective, config, start=None):
     objective.update_convergence(converged)
     
     return x_k
+

### 2024-12-01 src/seismic_model.py

@@ -13,6 +13,7 @@ from numpy import linalg
 from matplotlib import pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
 import functions as fn
+import optimizer as opt
 
 
 class Model:
@@ -21,13 +22,13 @@ class Model:
     Catch anything whose method is not implemented.
     '''
     
-    def __call__(self, params):
+    def __call__(self):
         raise Exception('Class is not callable')
 
-    def gradient(self, params):
+    def gradient(self):
         raise Exception('Method "gradient" is not implemented')
 
-    def hessian(self, params):
+    def hessian(self):
         raise Exception('Method "hessian" not is implemented')
 
     def starting_point(self):
@@ -56,7 +57,7 @@ class SeismicModel(Model):
         i, j = np.deg2rad(i), np.deg2rad(j)
         alpha_h, beta_h = velocities
         
-        # initialize misfits, iterates and optimals for tracking
+        # initialize variables for tracking
         self.misfits = []
         self.iterates = []
         self.optimal_iterates = []
@@ -66,6 +67,14 @@ class SeismicModel(Model):
         self.optimal_amplitudes = []
         self.half_angles = []
         self.runs, self.converged = 0, 0
+        
+        # initialize variables for hybrid search
+        self.convergence_rates = []
+        self.optimal_parameterizations = []
+        self.optimal_errors = []
+        self.optimal_axes = []
+        self.sampled_amplitudes = []
+        self.sampled_weights = []
         
         # initialize constant Jacobian matrix
         factors = np.array([alpha_h, beta_h, beta_h])**3
@@ -106,7 +115,7 @@ class SeismicModel(Model):
         eta = np.array([sR, qR, pR, qL, pL])
         self.A = self.J_A @ eta
         
-        # optionally set the observed amplitudes
+        # set observed amplitudes for synthetic parameters
         if set_Ao: self.Ao = self.A
         
         # compute and store misfit
@@ -289,7 +298,8 @@ class SeismicModel(Model):
     
     def plot_half_angles(self, bins=20):
         '''
-        Plot the half angles in a subplot.
+        Plot the half angles in a histogram.
+        This is a diagnostic plot.
         '''
         fig, ax = plt.subplots(1, 1, figsize=(6, 5))
         ax.hist(np.rad2deg(self.half_angles), bins=bins)
@@ -309,6 +319,8 @@ class SeismicModel(Model):
[…]

### 2024-12-09 src/beachball.py
salvation

@@ -1,9 +1,0 @@
-'''
-Name: Victor Agaba
-
-Date: 28th November 2024
-
-The goal of this module is to provide all helper functions needed for
-beachball plotting.
-'''
-

### 2024-12-09 src/beachplot.py

@@ -1,0 +1,55 @@
+'''
+Name: Victor Agaba
+
+Date: 8th December 2024
+
+The goal of this module is to provide all helper functions needed for
+beachball plotting.
+'''
+
+import matplotlib.pyplot as plt
+import numpy as np
+from obspy.imaging.beachball import beach
+import functions as fn
+
+
+def grid_beach(solution_set, width, max_plot, facecolor):
+    '''
+    (ref Omkar)
+    Plot complete solution set, with n figures in each plot
+    Solution set is a list of [s, d, r] for each focal mechanism
+    Input in DEGREES!!!
+    '''
+    # create figure
+    fig, ax = plt.subplots(subplot_kw={'aspect': 'equal'})
+    
+    # hide axes and ticks
+    ax.axison = False
+     
+    # plot all solutions
+    for counter, solution in enumerate(solution_set):
+        x = 210 * (counter % width)
+        y = 210 * (counter // width)
+        
+        solution = np.rad2deg(fn.bound(solution))
+        collection = beach(solution, xy=(x, y), facecolor=facecolor)
+        ax.add_collection(collection)
+        
+        if counter == max_plot:
+            break
+    
+    # scale and plot
+    ax.autoscale_view()
+    if facecolor == 'red': ax.set_title("Central mechanism solutions")
+    else: ax.set_title("Best fitting solutions")
+    
+    plt.show()
+    
+
+def overlay_beach(solution_set, facecolor):
+    '''
+    (ref Suzan)
+    Plot overlayed beachballs
+    '''
+    pass
+

### 2024-12-09 src/functions.py

@@ -87,7 +87,7 @@ def starting_direc(point: list, direc: list) -> list:
     
     proj = (np.dot(direc, point)/np.dot(point, point))
     rem = np.array(direc) - proj*np.array(point)
-    return rem/linalg.norm(rem) if linalg.norm(rem) > eps else j_hat
+    return unit_vec(rem) if linalg.norm(rem) > eps else j_hat
 
 
 def rect2pol(rect: list) -> list:

### 2024-12-09 src/garbage.py

@@ -82,3 +82,5 @@ class InversionModel:
             ax.legend()
         
         plt.show()
+        
+        

### 2024-12-09 src/optimizer.py

@@ -10,6 +10,7 @@ for inversion of seismic data.
 
 import numpy as np
 from numpy import linalg
+import functions as fn
 
 
 def get_config():

### 2024-12-09 src/other/omkar_beach.py

@@ -1,15 +1,16 @@
+'''
+Name: Victor Agaba
+
+Date: 28th November 2024
+
+The goal of this module is to provide all helper functions needed for
+beachball plotting.
+'''
+
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
-
-from obspy.core.trace import Trace, Stats
-from obspy.core.stream import Stream
-from obspy import UTCDateTime
-from obspy.clients.fdsn import Client
-
-from obspy.imaging.beachball import beachball
 from obspy.imaging.beachball import beach
-from obspy.imaging.source import plot_radiation_pattern
 
 import argparse
 from constants import DATA_PATH, EXP_PATH
@@ -18,7 +19,6 @@ from datetime import datetime
 from pathlib import Path
 import utils 
 import warnings 
-import sys 
 import os 
 
 
@@ -38,7 +38,7 @@ def alphabeach(event, df, n, norm_const, c='b'):
     # Max-min normalization 
     # df['Normalized'] = (max_mis - df['Misfit'])/(max_mis-min_mis)
     
-    # Gaussian normalization 
+    # Gaussian normalization NOTE ??? probably for transparency, not relevant
     sigma = max_mis 
     # pref = 1/(sigma*np.sqrt(2*np.pi)) 
     df['Normalized'] = norm_const*np.exp(-(np.square(df['Misfit']))/(2*np.square(sigma)))
@@ -47,12 +47,20 @@ def alphabeach(event, df, n, norm_const, c='b'):
     counter = 1
     num_figs = 0
 
+    # NOTE: does not have to be a dataframe
     for index, rows in df.iterrows():
+        
+        # NOTE: i want all strikes, dips and rakes in degrees
+        # NOTE: f is a list of 3 floats: [strike, dip, rake], one per focal mechanism
         f = [rows.Strike, rows.Dip, rows.Rake]
+        
+        # NOTE: default beachball size is 200
+        # NOTE: i want 10 beachballs per row
         x = 210 * ((counter-1) % 10)
         y = 210 * ((counter-1) // 10)
 
         # Collection1 is for individual solutions 
+        # NOTE: collection1 is actually a single beachball, added one at a time
         collection1  = beach(f, xy=(x, y), facecolor=c, alpha=rows.Normalized)
         ax1.add_collection(collection1)
 
@@ -60,8 +68,8 @@ def alphabeach(event, df, n, norm_const, c='b'):
[…]

### 2024-12-09 src/seismic_model.py

@@ -14,6 +14,7 @@ from matplotlib import pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
 import functions as fn
 import optimizer as opt
+import beachplot as bp
 
 
 class Model:
@@ -68,6 +69,7 @@ class SeismicModel(Model):
         self.optimal_amplitudes = []
         self.half_angles = []
         self.runs, self.converged = 0, 0
+        self.filtered_outliers = False
         
         # initialize variables for hybrid search
         self.convergence_rates = []
@@ -247,12 +249,29 @@ class SeismicModel(Model):
     
     def get_tp_axes(self):
         ''' Return the tp axes. '''
+        if len(self.tp_axes) == 0: self.mirror(['axes'])
         return self.tp_axes
     
     
     def get_half_angles(self):
         ''' Return the half angles. '''
         return self.half_angles
+    
+    
+    def get_central_tps(self):
+        ''' Return the central tps for the optimal tp axes. '''
+        if len(self.tp_axes) == 0: self.mirror(['axes'])
+        central, pos = fn.regression_axes(self.tp_axes)
+        output = np.zeros_like(self.tp_axes)
+        
+        for i in range(len(self.tp_axes)):
+            tp = self.tp_axes[i]
+            to_project = tp[pos]
+            projected = fn.starting_direc(central, to_project)
+            to_insert = [central, projected] if pos == 0 else [projected, central]
+            output[i] = to_insert
+        
+        return output
     
     
     def filter_outliers(self, threshold=2):
@@ -265,6 +284,7 @@ class SeismicModel(Model):
                                    if z_scores[i] < threshold]
         self.optimal_laplacians = [l for i, l in enumerate(self.optimal_laplacians)
                                       if z_scores[i] < threshold]
+        self.filtered_outliers = True
     
     
     def reset(self):
@@ -279,13 +299,14 @@ class SeismicModel(Model):
         self.optimal_amplitudes = []
         self.half_angles = []
         self.runs, self.converged = 0, 0
+        self.filtered_outliers = False
         
     
     def mirror(self, what, index=0):
         '''
         Mirror the optimal parameters to show only one fault plane.
         '''
-        self.filter_outliers()
+        if not self.filtered_outliers: self.filter_outliers()
         
         if len(what) == 1:
             what = what[0]
[…]

### 2024-12-09 src/utils.py

@@ -1,0 +1,99 @@
+import pickle 
+import numpy as np 
+from obspy.core.trace import Trace, Stats
+from obspy.core.stream import Stream
+from obspy import read, read_inventory
+from obspy.clients.fdsn import Client
+import re 
+from constants import DATA_PATH 
+import os 
+
+
+def save_file(filename, file):
+    """
+    Save file in pickle format
+    Args:
+        file (any object): Can be any Python object. We would normally use this to save the
+        processed Pytorch dataset
+        filename (str): Name of the file
+    """
+
+    with open(filename, 'wb') as f:
+        pickle.dump(file, f, protocol=pickle.HIGHEST_PROTOCOL)
+
+
+def load_file(filename):
+    """
+    Load a pickle file
+    Args:
+        filename (str): Name of the file
+    Returns (Python obj): Returns the loaded pickle file
+    """
+    with open(filename, 'rb') as f:
+        file = pickle.load(f)
+
+    return file
+
+
+def uvw2enz(st):
+    """
+    Rotate the waveforms 
+    """
+    if len(st) != 3:
+       print('Stream does not contain 3 Traces')
+       return st
+    for trace in st:
+        head = trace.stats
+        channel = head.channel
+        if channel == 'BHU': U = trace.data
+        elif channel == 'BHV': V = trace.data
+        elif channel == 'BHW': W = trace.data
+        else:
+            print('Trace.channel is not BHU, BHV, or BHW')
+            return st
+
+    d = np.radians(-30)
+    aU = np.radians(135)
+    aV = np.radians(15)
+    aW = np.radians(255)
+
+    A = np.array([[np.cos(d)*np.sin(aU), np.cos(d)*np.cos(aU),-np.sin(d)],
+                  [np.cos(d)*np.sin(aV), np.cos(d)*np.cos(aV), -np.sin(d)],
+                  [np.cos(d)*np.sin(aW), np.cos(d)*np.cos(aW), -np.sin(d)]])
+
+    B = np.linalg.inv(A)
+    E,N,Z = np.dot(B,(U,V,W))
+
+    head.channel = 'BHE'; trE = Trace(data=E, header=head)
+    head.channel = 'BHN'; trN = Trace(data=N, header=head)
+    head.channel = 'BHZ'; trZ = Trace(data=Z, header=head)
+    stENZ = Stream(traces=[trE,trN,trZ])
+
+    return stENZ
+
+
+def get_waveforms(start, end, event, args):
[…]

### 2024-12-10 src/model.py
reefactoring code

@@ -1,0 +1,584 @@
+'''
+Name: Victor Agaba
+
+Date: 2nd November 2024
+
+The goal of this module is to create a model of the nonlinear inverse
+problem to which we will apply the optimization algorithms.
+'''
+
+
+import numpy as np
+from numpy import linalg
+import functions as fn
+import optimizer as opt
+
+
+class Model:
+    '''
+    Base class for an unconstrained optimzation problem.
+    Catch anything whose method is not implemented.
+    '''
+    
+    def __call__(self):
+        raise Exception('Class is not callable')
+
+    def gradient(self):
+        raise Exception('Method "gradient" is not implemented')
+
+    def hessian(self):
+        raise Exception('Method "hessian" not is implemented')
+
+    def starting_point(self):
+        raise Exception('Method "starting_point" not implemented')
+
+
+class SeismicModel(Model):
+    '''
+    This class computes the misfit function and its gradient
+    for a given model.
+    Uses the same notation as in the paper.
+    '''
+    def __init__(self, azimuth, takeoff_angles, velocities, Ao = []):
+        '''
+        Initialize with the parts of the model that don't change.
+        All angles are in RADIANS!!!
+        
+        Args:
+        phi = raypath's azimuth
+        i, j = P and S wave takeoff angles
+        alpha, beta = P and S wave velocities
+        Ao = observed amplitudes
+        '''
+        # extract inputs
+        i, j = takeoff_angles
+        i, j = np.deg2rad(i), np.deg2rad(j)
+        alpha_h, beta_h = velocities
+        
+        # initialize variables for tracking
+        self.misfits = []
+        self.iterates = []
+        self.optimal_iterates = []
+        self.optimal_laplacians = []
+        self.misfits = []
+        self.tp_axes = []
+        self.amplitudes = []
+        self.optimal_amplitudes = []
+        self.half_angles = []
+        self.runs, self.converged = 0, 0
+        self.filtered_outliers = False
+        
+        # initialize variables for hybrid search
+        self.convergence_rates = []
+        self.optimal_parameterizations = []
+        self.optimal_errors = []
+        self.optimal_axes = []
+        self.sampled_amplitudes = []
+        self.sampled_weights = []
+        
[…]

### 2024-12-10 src/plots.py

@@ -1,0 +1,472 @@
+'''
+Name: Victor Agaba
+
+Date: 10th December 2024
+
+The goal of this module is to implement plotting functions for
+the data analysis.
+'''
+
+import numpy as np
+from matplotlib import pyplot as plt
+from obspy.imaging.beachball import beach
+import functions as fn
+
+
+def misfit(model):
+    '''
+    Plot the misfits in a subplot.
+    '''
+    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
+    ax.plot(model.misfits)
+    ax.set_title('Misfit function')
+    ax.set_xlabel('Iteration')
+    ax.set_ylabel('Misfit')
+    plt.show()
+    
+
+def half_angles(model, bins=20):
+    '''
+    Plot the half angles in a histogram.
+    This is a diagnostic plot.
+    '''
+    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
+    ax.hist(np.rad2deg(model.half_angles), bins=bins)
+    ax.set_title('Diagnostic histogram of half-angles')
+    ax.set_xlabel('Half angle (deg)')
+    ax.set_ylabel('Frequency')
+    plt.show()
+    
+    
+def iterates_2D(model, cmap='rainbow', s=10, optimal=True, index=2):
+    '''
+    Make 3 2D plots of the iterates: psi, delta, lambda.
+    psi against delta, psi against lambda, delta against lambda.
+    Join the points with a line and color them by iteration number.
+    
+    Args:
+        optimal (bool): If True, plot only the optimal points.
+        index (int): 0 is 1st fault plane, 1 is 2nd fault plane, 2 is both.
+    '''
+    fig, (ax1, ax2, ax3) = plt.subplots(1, 3, figsize=(18, 5))
+    
+    if index == 0 or index == 1:
+        if not optimal:
+            model.mirror(['optimals', 'iterates'], index)
+            iterates = model.iterates
+            optimal_iterates = model.optimal_iterates
+        else:
+            model.mirror(['optimals'], index)
+            optimal_iterates = model.optimal_iterates
+    elif index == 2:
+        iterates, optimal_iterates = [], []
+        if not optimal:
+            model.mirror(['optimals', 'iterates'], 0)
+            iterates.extend(model.iterates)
[…]

### 2024-12-10 src/seismic_model.py

@@ -1,977 +1,0 @@
-'''
-Name: Victor Agaba
-
-Date: 2nd November 2024
-
-The goal of this module is to create a model of the nonlinear inverse
-problem to which we will apply the optimization algorithms.
-'''
-
-
-import numpy as np
-from numpy import linalg
-from matplotlib import pyplot as plt
-from mpl_toolkits.mplot3d import Axes3D
-import functions as fn
-import optimizer as opt
-import beachplot as bp
-
-
-class Model:
-    '''
-    Base class for an unconstrained optimzation problem.
-    Catch anything whose method is not implemented.
-    '''
-    
-    def __call__(self):
-        raise Exception('Class is not callable')
-
-    def gradient(self):
-        raise Exception('Method "gradient" is not implemented')
-
-    def hessian(self):
-        raise Exception('Method "hessian" not is implemented')
-
-    def starting_point(self):
-        raise Exception('Method "starting_point" not implemented')
-
-
-class SeismicModel(Model):
-    '''
-    This class computes the misfit function and its gradient
-    for a given model.
-    Uses the same notation as in the paper.
-    '''
-    def __init__(self, azimuth, takeoff_angles, velocities, Ao = []):
-        '''
-        Initialize with the parts of the model that don't change.
-        All angles are in RADIANS!!!
-        
-        Args:
-        phi = raypath's azimuth
-        i, j = P and S wave takeoff angles
-        alpha, beta = P and S wave velocities
-        Ao = observed amplitudes
-        '''
-        # extract inputs
-        i, j = takeoff_angles
-        i, j = np.deg2rad(i), np.deg2rad(j)
-        alpha_h, beta_h = velocities
-        
-        # initialize variables for tracking
-        self.misfits = []
-        self.iterates = []
-        self.optimal_iterates = []
-        self.optimal_laplacians = []
-        self.misfits = []
-        self.tp_axes = []
-        self.amplitudes = []
-        self.optimal_amplitudes = []
-        self.half_angles = []
-        self.runs, self.converged = 0, 0
-        self.filtered_outliers = False
-        
-        # initialize variables for hybrid search
-        self.convergence_rates = []
-        self.optimal_parameterizations = []
-        self.optimal_errors = []
-        self.optimal_axes = []
[…]

### 2024-12-10 project/paper/paper.tex
organize

--- 2024-11-05 paper/main.tex
@@
 
-    Combining equations \ref{eq:7} and \ref{eq:8} gives the plane containing every point of 
+    \noindent Combining equations \ref{eq:7} and \ref{eq:8} gives the plane containing every point of 
     tangency:
@@
 
-    To find the constrained points of tangency, we can intersect the original ellipsoid with planes
+    \noindent To find the constrained points of tangency, we can intersect the original ellipsoid with planes
     \ref{eq:9} and \ref{eq:10}. The line of intersection is such that
@@
     \noindent where $c_2 = \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \langle\bm{A}_o^\top, \bm{A}_s\rangle_{\tilde{\bm{\Sigma}}^{-1}}$.
+    Equation \ref{eq:13} is the triple product expansion of constituent vectors.
 
-    Combining equations \ref{eq:7} and \ref{eq:11} produces a quadratic parameterized by the scalar $t$,
+    \noindent Combining equations \ref{eq:7} and \ref{eq:11} produces a quadratic parameterized by the scalar $t$,
     from which we can obtain a quadratic equation for the points of tangency:
@@
 
-    Denote $c_3 = \bm{A}_s^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s\|_{\tilde{\bm{\Sigma}}^{-1}}^2$. Then by
+    \noindent Denote $c_3 = \bm{A}_s^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s\|_{\tilde{\bm{\Sigma}}^{-1}}^2$. By
     substitution we have
@@
 
-    Thus, given 2 nonparallel vectors $\bm{A}_o$ and $\bm{A}_s$ and positive definite covariance
-    matrix $\tilde{\bm{\Sigma}}$, we accept the simulation if $\hat{\bm{A}}_o^\top \hat{\bm{A}_s} \geq \text{tol} 
-    = \min_b \hat{\bm{A}}_o^\top \hat{\bm{A}_b}$. The condition, more formally, is
-    \begin{align} \label{eq:16}
-      \max_b \hat{\bm{A}}_o^\top (\hat{\bm{A}}_b - \hat{\bm{A}}_s) \leq 0
-    \end{align}
+    \noindent The calculated boundary points are necessary to design reliable Monte-Carlo methods for uncertainty
+    quantification (see section \ref{sec:uncertainty}).
 
-    \noindent where $\hat{\bm{A}}_o$, $\hat{\bm{A}}_s$ and $\hat{\bm{A}}_b$ are unit vectors.
+    % Thus, given 2 nonparallel vectors $\bm{A}_o$ and $\bm{A}_s$ and positive definite covariance
+    % matrix $\tilde{\bm{\Sigma}}$, we accept the simulation if $\hat{\bm{A}}_o^\top \hat{\bm{A}_s} \geq \text{tol} 
+    % = \min_b \hat{\bm{A}}_o^\top \hat{\bm{A}_b}$. The condition, more formally, is
+    % \begin{align} \label{eq:16}
+    %   \max_b \hat{\bm{A}}_o^\top (\hat{\bm{A}}_b - \hat{\bm{A}}_s) \leq 0
+    % \end{align}
+
+    % \noindent where $\hat{\bm{A}}_o$, $\hat{\bm{A}}_s$ and $\hat{\bm{A}}_b$ are unit vectors.
 
@@
     \ref{eq:1}. We can thus apply directed search methods to minimize the following cost function:
-    \begin{align} \label{eq:17}
+    \begin{align} \label{eq:16}
       \Phi(\bm{m}) = -\hat{\bm{A}}_o^\top \hat{\bm{A}}
@@
     the misfit as
-    \begin{align} \label{eq:18}
-      \Phi(\bm{m}) = \Phi(\bm{A}(\bm{\eta}(\bm{m})))
+    \begin{align} \label{eq:17}
+      \Phi(\bm{m}) = \tilde{\Phi}(\bm{A}(\bm{\eta}(\bm{m})))
     \end{align}
@@
     \end{align*}
-    \begin{align} \label{eq:19}
+    \begin{align} \label{eq:18}
       \nonumber&\bm{J_p}(\bm{\eta}) = \\
@@
     of coefficients from equation \ref{eq:1}:
-    \begin{align} \label{eq:20}
+    \begin{align} \label{eq:19}
       \nonumber&\bm{J_\eta}(\bm{A}) = \\
@@
     transposed.
-    \begin{align} \label{eq:21}
-      \nonumber\nabla_{\bm{A}}\Phi &= \frac{\|\bm{A}\|\nabla_{\bm{A}}(-\hat{\bm{A}}_o^\top\bm{A}) - 
+    \begin{align} \label{eq:20}
+      \nonumber\nabla_{\bm{A}}\tilde{\Phi} &= \frac{\|\bm{A}\|\nabla_{\bm{A}}(-\hat{\bm{A}}_o^\top\bm{A}) - 
       (-\hat{\bm{A}}_o^\top \bm{A}) \nabla_{\bm{A}}(\|\bm{A}\|)}{\|\bm{A}\|^2}\\
@@
 
-    Combining \ref{eq:19}, \ref{eq:20} and \ref{eq:21} gives the overall gradient:
-    \begin{align} \label{eq:22}
-      \nabla_{\bm{m}}\Phi = \bm{J_p}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\Phi
+    Combining \ref{eq:18}, \ref{eq:19} and \ref{eq:20} gives the overall gradient:
+    \begin{align} \label{eq:21}
+      \nabla_{\bm{m}}\Phi = \bm{J_p}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\tilde{\Phi}
     \end{align}
@@
   \subsection{Optimization algorithms}
+
+  Insert description here to be backed up by figure 2...
+
+  Why isn't this doing its thing?
+
+  I know you can do it
+
+  Just a few more paragraphs
+
     \begin{figure}[ht!]
       \includegraphics[width=8.6cm]{seed_2024} 
-      \caption{Results of steepest descent algorithm after a systematic sweep of starting points.}
+      \caption{Results of steepest descent algorithm for randomly sampled starting points.}
       \label{fig:2}
@@
 \section{Uncertainty Quantification} \label{sec:uncertainty}
-
-    \subsection{Kagan angle distribution}
-      An alternative way of quantifying model deviations.
 

### 2024-12-10 project/paper/template/main.tex
organize

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Seismica Submission Template
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Options available: 
%%					anonymous
%%					breakmath
%%					languages
%%					preprint
%%					report

%% anonymous: produces an anonymous PDF for double-blind review. Will NOT print authors information and acknowledgements, but WILL PRINT data availibility section, so be careful!
%% breakmath: for manuscripts with many long formulas, you can specify the breakmath option (loads the package breqn and uses the dmath environment)
%% languages: see below, use to add abstract(s) in additional languages
%% preprint: removes line numbers, switch to two-columns
%% report: if this is a report

%\documentclass[breakmath,report]{seismica}
\documentclass[preprint]{seismica}

\title{Your article title goes here}
%\shorttitle{Your short title goes here} % used for header, not mandatory

%% If this is a REPORT, select a report type within: 
%\reporttype{Null Results Report} % (null-results/failed experiments)
%\reporttype{Software Report}
%\reporttype{Data Report} %  (e.g., Large Community dataset initiatives, Instrument Deployments, and Field Campaigns)
%\reporttype{Fast Report}

%% Will not be printed if anonymous option ON
\author[1]{Name Surname
	\orcid{1111-1111-1111-1111}
	\thanks{Corresponding author: firstauthor@university.edu}
}
\author[2]{Name~Secondauthor
	\orcid{2222-2222-2222-2222}
}
\author[1,3]{Name~Thirdauthor
	\orcid{3333-3333-3333-3333}
}
\affil[1]{Department of Earth Sciences, A University, City, Country}
\affil[2]{School of Earth Sciences, Another University, City, Country}
\affil[3]{Center for Studying Cool Things, University of X, City, Country}

%% Author CRediT roles 
%% Please use the CRediT roles as defined at https://casrai.org/credit
%% Use as many roles as necessary; there is no requirement to use all 14 roles
\credit{Conceptualization}{Name Firstauthor, Name Thirdauthor}
%\credit{Methodology}{people}
%\credit{Software}{people}
%\credit{Validation}{people}
\credit{Formal Analysis}{Name Firstauthor, Name Secondauthor}
%\credit{Investigation}{people}
%\credit{Resources}{people}
\credit{Writing - Original draft}{Name Firstauthor}
%\credit{Writing - Review \& Editing}{people}
%\credit{Visualization}{people}
%\credit{Supervision}{people}
%\credit{Project administration}{people}
%\credit{Funding acquisition}{people}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Abstracts in other languages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If your article includes abstracts in other languages, uncomment the lines below and fill in
%%      the appropriate sections. You will need to use the [languages] option at the top,
%%      and will need to use lualatex instead of pdflatex to compile the document.
%% We will use luatex, polyglossia and fontspec for the compilation of the accepted version. 
%% Feel free to use any polyglossia command.
%\setotherlanguages{french,thai}  % replace with your language(s), per polyglossia
%% if an additional font is needed for the abstract, load it with:
%\newfontfamily\thaifont[Script=Thai]{Noto Serif Thai}

%% If you are using Arabic, do not include it in \setotherlanguages{} as it is not supported 
%% by polyglossia. Instead, with the [languages] option at the top, you can use these commands
%% within the text:
%%\begin{Arabic} and \end{Arabic} around paragraphs in Arabic
%%\n{} to wrap any digits within Arabic text that should read left-to-right
%%\textarabic{} for Arabic text embedded in a left-to-right paragraph

%% also see https://www.overleaf.com/latex/examples/how-to-write-multilingual-text-with-different-scripts-in-latex/wfdxqhcyyjxz for reference
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Several abstracts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% the command \makeseistitle does not allow page breaks in preprint mode. If you have
%%		many abstracts, you can use the command \addsummaries. It will induce a pagebreak.

\begin{document}

	%% Your article can include up to 3 abstracts. The first is the English language abstract.
	%% For other languages in the second and third optional abstracts, you might have to define 
	%%      additional font(s) in preamble above
	%% You can also include a non-technical summary in addition to the abstract(s)
	\makeseistitle{
		\begin{summary}{Abstract}
			The text for the first abstract goes here. This should be in English, no longer than 200 words, and should not include references.
		\end{summary}
		%	 \begin{thai}
			%	 \begin{summary}{นามธรรม}
				%	คอรัปชันจุ๊ยโปรดิวเซอร์ สถาปัตย์จ๊าบ แจ็กพ็อต ม้าหินอ่อน ซากุระคันถธุระ ฟีดสตาร์ท งี้ บอยคอตอิ่มแปร้สังโฆคำสาปแฟนซี ศิลปวัฒนธรรมไฟลท์จิ๊กโก๋กับดัก เจลพล็อตมาม่าซากุระดีลเลอร์ ซีนดัมพ์ แฮปปี้ เอ๊าะอุรังคธาตุซิม ฟินิกซ์เทรลเล่อร์อวอร์ด แคนยอนสมาพันธ์ ครัวซองฮัมอาข่าเอ็กซ์เพรส 
				%	 \end{summary}
			%	 \end{thai}
		%	\begin{summary}{Non-technical summary}
			%	The text goes here. Again, no longer than 200 words.
			%	\end{summary}
	}  %% don't forget this one!

	%%%% In preprint mode, in case you have too many abstracts and need a page break, use this:
	%\addsummaries{
		%	\begin{french}
			%	\begin{summary}{Résumé} 
				%	The text goes here. Again, no longer than 200 words.
				%	\end{summary}
			%	\end{french}
		%	\begin{summary}{Non-technical summary}
			%	The text goes here. Again, no longer than 200 words.
			%	\end{summary}
		%}  %% don't forget this one!

	\section{Section name}

	All articles must include an abstract, author ORCIDs and author contributions (in the preamble of this tex file), a data and code availability statement, and a list of references.

	Section names are at the discretion of the authors. A simple structure for an article would include an Introduction, Methods and Data, Results, Discussion, and Conclusions, but authors are encouraged to choose a structure that best presents their work.

	\subsection{Bibliographic citations}
	In the text of an article, citations may either be in-line, as in the case of citing \citet{metropolis_monte_1949}, or in parentheses \citep[e.g.,][]{metropolis_monte_1949}, as appropriate. All citations in the text must be listed in the references section, and all listed references must be cited at least once in the text.

	\subsection{Headings}
	\subsubsection{Subsubsection}
	Three levels of section headings is the maximum\footnote{Seriously, the maximum} - no subsubsubsections, please! Note that footnotes are permitted, as in the previous sentence, though we encourage authors to really think about whether a footnote is necessary or if the information it contains could be included in the main text.

	\subsection{Figures and Tables}
	Figures should be labeled, captioned, and referenced within the text (e.g., Fig.~\ref{fig:1} and Figs~\ref{fig:1}a, b, \ref{fig:2}c). When an article is accepted, separate full-resolution files must be uploaded for each figure. While Figure \ref{fig:1} is a one-column figure, Figure \ref{fig:2} is a full-width figure.

	%% Accepted articles will be typeset in a 2-column format. 
	%% Figures can be either be 1-column wide or page-wide. One column is 8.6 cm wide.
	%% Here, Figure 1 is a 1-column figure.
	\begin{figure}[ht!]
		\includegraphics[width=8.6cm]{empty} 
		\caption{This is an example of a figure caption.}
		\label{fig:1}
	\end{figure}

	%% This is a page-wide figure (18 cm width). 
	\begin{figure*}[ht!]
		\centering
		\includegraphics[width=\textwidth]{empty} 
		\caption{This is a caption on wider figure.}
		\label{fig:2}
	\end{figure*}

	Tables can also be included, with captions.
	%% Use \begin{table*} for a page-wide table
		\begin{table}[ht!]
			\begin{tabular}{llll}
				Event ID    & Location & Magnitude & A random number \\
				\hline
				1 & Here & 2.5 & 17 \\
				2 & There & 4.1 & 1350 \\
			\end{tabular}
			\caption{Caption}
			\label{tab:1}
		\end{table}

		Tab.~\ref{tab:1} (use Tabs if several tables) is an example of a relatively simple table. We strongly encourage authors to put tables in Supplementary Materials, and/or into a csv or similar format, upload them to a data repository such as zenodo, and reference them in the section on data availability instead of including them in the article itself.

		\subsection{Equations and maths}
		Equations can be included in the text, and should be labeled so they can be referenced. One example is Equation \ref{eq1}:
		\begin{equation} \label{eq1}
			\mathrm{G} = \frac{1}{2}(2\cos z) + (1/2)(2\cos z+j\sin z-j\sin z) + (1/2)(\cos z+j\sin z+\cos z-j\sin z) -  (1/2)(e^{jz}+e^{-jz})  
		\end{equation}

		Please type vectors and matrices in bold: $\mathbf{X} = \left[x_1,x_2,\ldots,x_n \right]^T$.

		\subsection{Code}

		Code examples should be concise and descriptive. They should introduce core functionality or specific syntax and should be included using the \code{lstlisting} environment. Note that lines longer than 45 characters will be broken when using the prepress option. Extended examples or use cases should be uploaded separately. Individual words of code can be written inline, for example:

		To improve stability of the inversion, the \code{Model} object accepts the \code{strict} keyword, which disables piecewise linear approximation of the target function (Listing~\ref{code}).

		\begin{lstlisting}[caption=Example use of \code{Model}, label=code, language=Python]
			#2 4 6 8 0 2 4 6 8 0 2 4 6 8 0 2 4 6 8 0 2 4|
			import mymodule as mm

			model = mm.Model(strict=True)
			mdls = model.perturb()

			for mdl in mdls:
			var = mdl.get_variance()
		\end{lstlisting}

		%% Will not be printed if anonymous option ON
		\begin{acknowledgements}
			Thank all relevant parties and acknowledge funding sources, if any.
		\end{acknowledgements}

		\section*{Data and code availability}
		Authors should direct readers to an open access repositories where data and code used in the study are made available. Zenodo, figshare, and Dryad are examples of repositories where authors can archive their data and code. Citations for datasets and codes should be included in the references, including citations for any seismic networks from which data was used. Github is not considered a persistent repository, and we encourage authors to archive a snapshot of any github-hosted code on zenodo.

		\section*{Competing interests}
		Declare any competing interests, financial or otherwise, pertaining to any of the authors. If there are none, state that the authors have no competing interests.

		%% If the article is accepted, a separate bibfile must be uploaded along with the compiled manuscript, source file, and separate figure files.
		%% When available, DOI numbers must be provided for all references, including datasets and codes. 
		\bibliography{mybibfile}

	\end{document}

### 2024-12-10 project/code/package/__init__.py

@@ -1,0 +1,3 @@
+from .functions import *
+from .model import *
+from .optimizer import *

### 2024-12-11 project/code/package/__init__.py
events

@@ -1,3 +1,0 @@
-from .functions import *
-from .model import *
-from .optimizer import *

### 2024-12-11 project/package/model.py

@@ -10,8 +10,8 @@ problem to which we will apply the optimization algorithms.
 
 import numpy as np
 from numpy import linalg
-import functions as fn
-import optimizer as opt
+from . import functions as fn
+from . import optimizer as opt
 
 
 class Model:
@@ -253,6 +253,11 @@ class SeismicModel(Model):
         return self.misfits
     
     
+    def get_amplitudes(self):
+        ''' Return the amplitudes per iterate from optimization. '''
+        return self.amplitudes
+    
+    
     def get_convergence_rate(self):
         ''' Return the convergence rate. '''
         return 100*self.converged/self.runs
@@ -277,6 +282,11 @@ class SeismicModel(Model):
     def get_half_angles(self):
         ''' Return the half angles. '''
         return self.half_angles
+    
+    
+    def get_optimal_laplacians(self):
+        ''' Return the optimal laplacians. '''
+        return self.optimal_laplacians
     
     
     def get_central_tps(self):
@@ -315,12 +325,12 @@ class SeismicModel(Model):
         return flow
     
     
-    def get_optimal_parameterization(self):
+    def get_optimal_parameterization(self, threshold=2):
         '''
         Compute optimal parameterization of fault planes.
         Returns axis direction, half-angle, error and name.
         '''
-        if not self.filtered_outliers: self.filter_outliers()
+        if not self.filtered_outliers: self.filter_outliers(threshold)
         
         if len(self.tp_axes) == 0: self.mirror(['axes'])
         central, position = fn.regression_axes(self.tp_axes)
@@ -345,6 +355,7 @@ class SeismicModel(Model):
     #
     def filter_outliers(self, threshold=2):
         ''' Filter out outliers based on Z-score of optimal amplitude norms. '''
+        if self.filtered_outliers: return
         norms = np.array([linalg.norm(a) for a in self.optimal_amplitudes])
         z_scores = np.abs((norms - np.mean(norms)) / np.std(norms))
         self.optimal_iterates = [m for i, m in enumerate(self.optimal_iterates) 
@@ -393,7 +404,7 @@ class SeismicModel(Model):
             raise ValueError('Invalid argument for "what"')
         
         
-    def hybrid_search(self, num_runs=30, print_every=0):
+    def hybrid_search(self, num_runs=30, print_every=0, threshold=2):
         '''
         Perform a random hybrid search to find the optimal parameterization.
         '''
@@ -407,7 +418,7 @@ class SeismicModel(Model):
[…]

### 2024-12-11 project/package/optimizer.py

@@ -10,7 +10,7 @@ for inversion of seismic data.
 
 import numpy as np
 from numpy import linalg
-import functions as fn
+from . import functions as fn
 
 
 def get_config():

### 2024-12-11 project/package/plots.py

@@ -10,35 +10,36 @@ the data analysis.
 import numpy as np
 from matplotlib import pyplot as plt
 from obspy.imaging.beachball import beach
-import functions as fn
+from . import functions as fn
+from . import model as sm
 
 
-def misfit(model):
+def misfit(model: sm.SeismicModel):
     '''
     Plot the misfits in a subplot.
     '''
     fig, ax = plt.subplots(1, 1, figsize=(6, 5))
-    ax.plot(model.misfits)
+    ax.plot(model.get_misfits())
     ax.set_title('Misfit function')
     ax.set_xlabel('Iteration')
     ax.set_ylabel('Misfit')
     plt.show()
     
 
-def half_angles(model, bins=20):
+def half_angles(model: sm.SeismicModel, bins=20):
     '''
     Plot the half angles in a histogram.
     This is a diagnostic plot.
     '''
     fig, ax = plt.subplots(1, 1, figsize=(6, 5))
-    ax.hist(np.rad2deg(model.half_angles), bins=bins)
+    ax.hist(np.rad2deg(model.get_half_angles()), bins=bins)
     ax.set_title('Diagnostic histogram of half-angles')
     ax.set_xlabel('Half angle (deg)')
     ax.set_ylabel('Frequency')
     plt.show()
     
     
-def iterates_2D(model, cmap='rainbow', s=10, optimal=True, index=2):
+def iterates_2D(model: sm.SeismicModel, cmap='rainbow', s=10, optimal=True, index=2):
     '''
     Make 3 2D plots of the iterates: psi, delta, lambda.
     psi against delta, psi against lambda, delta against lambda.
@@ -53,25 +54,25 @@ def iterates_2D(model, cmap='rainbow', s=10, optimal=True, index=2):
     if index == 0 or index == 1:
         if not optimal:
             model.mirror(['optimals', 'iterates'], index)
-            iterates = model.iterates
-            optimal_iterates = model.optimal_iterates
+            iterates = model.get_iterates()
+            optimal_iterates = model.get_optimal_iterates()
         else:
             model.mirror(['optimals'], index)
-            optimal_iterates = model.optimal_iterates
+            optimal_iterates = model.get_optimal_iterates()
     elif index == 2:
         iterates, optimal_iterates = [], []
         if not optimal:
             model.mirror(['optimals', 'iterates'], 0)
[…]

### 2024-12-12 project/junk/download_and_plot_station_xml.py
i tried

@@ -1,0 +1,40 @@
+from obspy.clients.fdsn import Client
+from obspy import UTCDateTime
+
+import os
+
+# Define paths and parameters
+path = os.getcwd()
+dst = os.path.join(path, 'Stationxml/')
+os.makedirs(dst, exist_ok=True)
+
+client = Client("IRIS")
+netsclts = {"II": "IRIS"}
+starttime = UTCDateTime(2001, 1, 1, 0, 0, 0)
+endtime = UTCDateTime(2020, 12, 12, 23, 59, 59)
+
+# Loop through networks and retrieve metadata
+for net in list(netsclts.keys()):
+    client = Client(netsclts[net])
+    inventory = client.get_stations(
+        starttime=starttime,
+        endtime=endtime,
+        network=net,
+        sta="MBAR",
+        loc="00",
+        channel="*",
+        level="response"
+    )
+    inventory.write(dst + f"metadata.{net}.xml", format="STATIONXML")
+    print(f"Metadata for {net} written to {dst}")
+
+    # Plot instrument response for displacement
+    for network in inventory:
+        for station in network:
+            for channel in station:
+                if channel in ["BHZ","BHN","BHE"]:
+                    response = channel.response
+                    if response:
+                        print(f"Plotting instrument response for {channel.code}")
+                        response.plot(0.001, output="DISP")  # Frequency 0.001 Hz
+

### 2025-11-11 project/paper/paper.tex
back at it again
Farallon is down

@@
 
-\title{Single-seismometer Focal Mechanism and Uncertainty Estimation from Body Waves}
+% \title{Single-seismometer Focal Mechanism and Uncertainty Estimation from Body Waves}
+\title{An Asymmetry-Aware Method for Focal Mechanism and Uncertainty Estimation in Single-seismometer Settings}
 \shorttitle{Single-seismometer Focal Mechanism and Uncertainty Estimation} % used for header, not mandatory
@@
       \nonumber p_L &= 0.5\sin(\lambda) \sin(2\delta) \sin(2\psi_r) + \cos(\lambda) \sin(\delta) \cos(2\psi_r)\\
-      q_L &= -\cos(\lambda) \cos(\delta) \sin(\psi_r) + \sin(\lambda) \cos(2\delta) \cos(\psi_r)
+      \nonumber q_L &= -\cos(\lambda) \cos(\delta) \sin(\psi_r) + \sin(\lambda) \cos(2\delta) \cos(\psi_r)\\
+      \psi_r &= \psi - \phi
     \end{align}
@@
           \foreach \h in {1,2,3,4,5}
-            \draw (P\p) -- (H\h);
+            \draw[->] (P\p) -- (H\h);
 
@@
         \foreach \a in {1,2,3}
-          \draw (A\a) -- (M);
+          \draw[->] (A\a) -- (M);
 
         % \draw[dashed] (P1) -- (H1);
-        \draw (P1) -- (H2);
-        \draw (P1) -- (H3);
-        \draw (P1) -- (H4);
-        \draw (P1) -- (H5);
+        \draw[->] (P1) -- (H2);
+        \draw[->] (P1) -- (H3);
+        \draw[->] (P1) -- (H4);
+        \draw[->] (P1) -- (H5);
 
-        \draw (H1) -- (A3);
-        \draw (H1) -- (A1);
-        \draw (H1) -- (A2);
-        \draw (H2) -- (A1);
-        \draw (H2) -- (A2);
-        \draw (H3) -- (A1);
-        \draw (H3) -- (A2);
-        \draw (H4) -- (A3);
-        \draw (H5) -- (A3);
+        \draw[->] (H1) -- (A1);
+        \draw[->] (H1) -- (A2);
+        \draw[->] (H2) -- (A1);
+        \draw[->] (H2) -- (A2);
+        \draw[->] (H3) -- (A1);
+        \draw[->] (H3) -- (A2);
+        \draw[->] (H4) -- (A3);
+        \draw[->] (H5) -- (A3);
       \end{tikzpicture}
-      \caption{Computational graph of forward model (edges are non-zero partial derivatives).}
+      \caption{Computation graph of forward model (edges are non-zero partial derivatives).}
       \label{fig:1}
@@
       &d = \sin(2\delta)\;\;\;\;\;\;\;\;\;\;\;\; e = \sin(\delta)\;\;\;\;\;\;\;\;\;\;\;\; f = \cos(\delta)\\
-      &g = \sin(\psi - \phi)\;\;\;\;\;\; h = \cos(\psi - \phi)\;\;\;\;\; p = cos(2(\psi - \phi))\\
-      &q = \sin(2(\psi - \phi))
+      &g = \sin(\psi_r)\;\;\;\;\;\;\;\;\;\;\; h = \cos(\psi_r)\;\;\;\;\;\;\;\;\;\; p = cos(2(\psi_r)\\
+      &q = \sin(2\psi_r)
     \end{align*}
     \begin{align} \label{eq:18}
-      \nonumber&\bm{J_p}(\bm{\eta}) = \\
+      \nonumber&\bm{J_m}(\bm{\eta}) = \\
       &\begin{bmatrix}
@@
     \begin{align} \label{eq:21}
-      \nabla_{\bm{m}}\Phi = \bm{J_p}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\tilde{\Phi}
+      \nabla_{\bm{m}}\Phi = \bm{J_m}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\tilde{\Phi}
     \end{align}

### 2025-11-11 other/graveyard.py

@@ -1,0 +1,30 @@
+import numpy as np
+
+halfpi = np.pi/2
+twopi = np.pi*2
+
+
+def bound(params: list) -> list:
+    
+    turns = [halfpi, np.pi, twopi]
+    
+    # leave input unchanged
+    out = params.copy()
+    
+    # driven by dip: case for every quadrant
+    overturned = out[1] % turns[2] > turns[1]
+    out[1] = out[1] % turns[1]
+    
+    # strike-dip relationship
+    if out[1] > turns[0]:
+        out[1] = turns[1] - out[1]
+        out[0] += turns[1]
+    out[0] = out[0] % turns[2]
+    
+    # bound the rake
+    if overturned: out[2] += turns[1]
+    out[2] = out[2] % turns[2]
+    if out[2] > turns[1]:
+        out[2] -= turns[2]
+        
+    return out

### 2025-11-11 project/notebooks/toVictor.py

@@ -1,0 +1,302 @@
+
+# functions written by Suzan van der Lee (SL) @1993 - 2022
+#
+# to compute focal shpere parameters from Moment Tensor and v.v.
+#
+# T-axis means tension at the source and corresponding upward motion of the 
+# first P arrival at a site. The area of the T-axis on the focal sphere
+# is usually colored (black).
+# The P axis means compression at the source (and downward motion at sites).
+# The P axis area is usually left white. 
+# The T axis is associated with a positive eigenmoment and 
+# the P axis with a negative eigenmoment. 
+# When the T and P eigenmoments are equal (and of opposite sign) the
+# focal mechanism is a shear dislocation and can be modeled with a 
+# double couple moment tensor (or Moment, srike, dip, rake)
+# Adding a CLVD component or a minor double couple adds a parameter.
+# Adding an isotropic component adds another parameter. 
+# The moment tensor is symmetric and thus has at most 6 parameters. 
+
+# Aki & Richards coordinate system: 
+# z (vertical, positive down)
+# x (positive North)
+# y (positive East)
+# Normal mode coordinate system (used by Global CMT):
+#  r (vertical, positive up)
+#  theta (t), or delta (epicentral distance) (positive South)
+#  phi (p or f) (positive East)
+# Therefore,
+# mt(1-6): Mrr, Mtt, Mff, Mrt, Mrf, Mtf
+# mt(1-6): Mzz, Mxx, Myy, Mzx, -Mzy, -Mxy
+
+######################################################
+
+import numpy as np
+
+eps = 1.e-6
+halfpi = 0.5*np.pi 
+twopi = 2*np.pi
+
+def perp(vin):
+    """
+    IN: a vector (e.g. a T axis) 
+    Rotate a vector over 90 degrees around any axis
+    OUT: rotated vector of the same length as the input vector
+    """
+    vinl = np.linalg.norm(vin)
+    l = np.argmin(np.abs(vin))
+    m = np.argmax(np.abs(vin))
+    k = 0; j = 1; i = 2
+    if k != l:
+        if j != l:
+            i = k
[…]

### 2025-11-11 project/package/functions.py

@@ -174,6 +174,7 @@ def rotate_vec(vec: list, axis: list, theta: float) -> list:
     
     return rotated_vec
 
+###################################################################
 
 def tp2sdr(t: list, p: list, deg: bool = False) -> tuple:
     '''
@@ -285,6 +286,7 @@ def sdr2tp(sdr: list, deg: bool = False) -> tuple:
     
     return t, p
 
+#####################################################################
 
 def bound(params: list) -> list:
     '''
@@ -294,29 +296,14 @@ def bound(params: list) -> list:
     Args:
         params (list): [strike, dip, rake] in RADIANS
     '''
+    t, p = sdr2tp(params, deg=False)
+    sdr1, sdr2 = tp2sdr(t, p)
     
-    turns = [halfpi, np.pi, twopi]
+    for idx, param in enumerate(params):
+        if param == sdr1[idx]: return sdr1
+        elif param == sdr2[idx]: return sdr2
     
-    # leave input unchanged
-    out = params.copy()
-    
-    # driven by dip: case for every quadrant
-    overturned = out[1] % turns[2] > turns[1]
-    out[1] = out[1] % turns[1]
-    
-    # strike-dip relationship
-    if out[1] > turns[0]:
-        out[1] = turns[1] - out[1]
-        out[0] += turns[1]
-    out[0] = out[0] % turns[2]
-    
-    # bound the rake
-    if overturned: out[2] += turns[1]
-    out[2] = out[2] % turns[2]
-    if out[2] > turns[1]:
-        out[2] -= turns[2]
-    
-    return out
+    return sdr1
 
 
 def systematic_hemisphere_samples(step_size: float) -> list:
@@ -341,7 +328,8 @@ def systematic_hemisphere_samples(step_size: float) -> list:
     return samples
 
 
-def systematic_params(dd: float) -> list:
+def systematic_params(dd: float, hidden: bool = False, az: float = 0,
+                      in_deg: bool = True) -> list:
     '''
     Returns n systematically sampled parameters for the inversion.
     Sampled from tp space for better coverage.
@@ -357,9 +345,12 @@ def systematic_params(dd: float) -> list:
         P_start = starting_direc(T, i_hat + j_hat + k_hat)
         for theta in P_rotations:
[…]

### 2025-11-11 project/package/model.py

@@ -39,42 +39,51 @@ class RadiationModel(Model):
     for a given model.
     Uses the same notation as in the paper.
     '''
-    def __init__(self, azimuth, takeoff_angles, velocities, Ao = []):
+    def __init__(self, azimuth, takeoff_angles, velocities):
         '''
         Initialize with the parts of the model that don't change.
+        '''
+        # initialize parent class
+        super().__init__()
+        
+        # store the original/main hyperparams
+        self.phi = azimuth
+        self.takeoff_angles = takeoff_angles
+        self.hidden = False
+        
+        # initialize variables for many runs
+        self.super_reset()
+        self.set_hyperparams(azimuth, takeoff_angles, velocities)
+    
+    
+    def set_hidden_params(self, hidden=True):
+        '''
+        Use the hidden parameterization for optimization.
+        '''
+        self.hidden = hidden
+    
+    
+    def is_hidden(self):
+        '''
+        Return whether the model is in hidden parameterization.
+        '''
+        return self.hidden
+    
+    
+    def set_hyperparams(self, azimuth, takeoff_angles, velocities):
+        '''
+        Set the hyperparameters of the model.
         All angles are in RADIANS!!!
         
         Args:
         phi = raypath's azimuth
         i, j = P and S wave takeoff angles
         alpha, beta = P and S wave velocities
-        Ao = observed amplitudes
         '''
         # extract inputs
         i, j = takeoff_angles
         i, j = np.deg2rad(i), np.deg2rad(j)
         alpha_h, beta_h = velocities
-        
-        # initialize variables for tracking
-        self.misfits = []
-        self.iterates = []
-        self.optimal_iterates = []
-        self.optimal_laplacians = []
-        self.misfits = []
-        self.tp_axes = []
-        self.amplitudes = []
-        self.optimal_amplitudes = []
-        self.half_angles = []
-        self.runs, self.converged = 0, 0
-        self.filtered_outliers = False
-        
-        # initialize variables for hybrid search
-        self.convergence_rates = []
-        self.optimal_parameterizations = []
-        self.optimal_errors = []
-        self.optimal_axes = []
-        self.sampled_amplitudes = []
-        self.sampled_weights = []
[…]

### 2025-11-11 project/package/optimizer.py

@@ -27,6 +27,7 @@ def get_config():
         rho (float): Backtracking line search parameter.
         tolerance (float): Stopping criterion.
         print_every (int): Print output every n iterations.
+        misfit_tolerance (float): Tolerance for goodness of fit.
     '''
     config = {
         'method': 'SD',
@@ -35,12 +36,14 @@ def get_config():
         'alpha': 5e-2,
         'rho': 0.5,
         'tolerance': 1e-5,
-        'print_every': 0
+        'print_every': 0,
+        'misfit_tolerance': 1e-5,
+        'global_min': -1.0
     }
     return config
 
 
-def minimize(objective, config, start=None):
+def minimize(objective, config, start=None, laplacian=True, converge_globally=False):
     '''
     Minimize an objective function starting from a given point.
     
@@ -48,6 +51,8 @@ def minimize(objective, config, start=None):
         function (callable): The objective function to minimize.
         config (dict): Hyperparameters for the optimization.
         start (list): Starting point for the optimization.
+        laplacian (bool): Whether to compute and store the Laplacian at the optimal point.
+        global_min (bool): Whether to assume the global minimum is found.
     
     Output:
         x_k (list): The optimal point found by the optimization.
@@ -61,7 +66,9 @@ def minimize(objective, config, start=None):
     rho = config['rho']
     tolerance = config['tolerance']
     print_every = config['print_every']
-    
+    misfit_tolerance = config['misfit_tolerance']
+    global_min = config['global_min']
+
     # get initial guess
     if start is None:
         start = function.get_start()
@@ -145,9 +152,13 @@ def minimize(objective, config, start=None):
             break
     
     if converged:
-        objective.set_optimal_iterate(x_k)
-        objective.set_optimal_amplitude()
-        objective.set_optimal_laplacian(x_k)
+        if (converge_globally and abs(f_k - global_min) < misfit_tolerance) or (not converge_globally):
+            objective.set_optimal_iterate(x_k)
+            objective.set_optimal_amplitude()
+            if laplacian: objective.set_optimal_laplacian(x_k)
+        else:
+            if print_every > 0:
+                print('Warning: Converged to a local minimum!') 
         
     objective.update_convergence(converged)
     

### 2025-11-11 project/package/plots.py

@@ -12,9 +12,13 @@ from matplotlib import pyplot as plt
 from obspy.imaging.beachball import beach
 from . import functions as fn
 from . import model as sm
+from matplotlib.patches import Ellipse
+from scipy.stats import chi2
+from mpl_toolkits.mplot3d import Axes3D  # noqa: F401
+from mpl_toolkits.mplot3d import art3d
 
 
-def misfit(model: sm.SeismicModel):
+def misfit(model: sm.RadiationModel):
     '''
     Plot the misfits in a subplot.
     '''
@@ -26,7 +30,7 @@ def misfit(model: sm.SeismicModel):
     plt.show()
     
 
-def half_angles(model: sm.SeismicModel, bins=20):
+def half_angles(model: sm.RadiationModel, bins=20):
     '''
     Plot the half angles in a histogram.
     This is a diagnostic plot.
@@ -38,8 +42,55 @@ def half_angles(model: sm.SeismicModel, bins=20):
     ax.set_ylabel('Frequency')
     plt.show()
     
-    
-def iterates_2D(model: sm.SeismicModel, cmap='rainbow', s=10, optimal=True, index=2):
+
+def cov_ellipse(ax, x, y, cov_xy, conf=0.95, **kwargs):
+    """
+    Plot a covariance ellipse on a given Matplotlib Axes.
+
+    Parameters
+    ----------
+    ax : matplotlib.axes.Axes
+        The axes to plot the ellipse on.
+    x, y : float
+        Mean coordinates (center of the ellipse).
+    cov_xy : (2, 2) array-like
+        Covariance matrix for (x, y).
+    conf : float, optional (default=0.95)
+        Confidence level for the ellipse (e.g., 0.68, 0.95, 0.99).
+    **kwargs :
+        Additional keyword arguments passed to matplotlib.patches.Ellipse,
+        e.g. edgecolor, facecolor, lw, alpha, linestyle.
+
+    Returns
+    -------
+    ellipse : matplotlib.patches.Ellipse
+        The created ellipse patch (already added to the axes).
+    """
+    # cov_xy = np.asarray(cov_xy)
+    if cov_xy.shape != (2, 2):
+        raise ValueError("cov_xy must be a 2x2 covariance matrix")
+
+    # confidence scaling factor from chi-squared distribution
+    scale = np.sqrt(chi2.ppf(conf, df=2))
+
[…]

### 2025-11-11 suzan/package/__init__.py

@@ -1,0 +1,1 @@
+__all__ = ['functions', 'model', 'optimizer', 'plots']

### 2025-11-11 suzan/package/functions.py

@@ -1,0 +1,549 @@
+'''
+Name: Victor Agaba
+
+Date: 4th November 2024
+
+The goal of this module is to provide all helper functions needed for the
+seismic inversion project.
+
+'''
+
+
+import numpy as np
+from numpy import linalg
+import pandas as pd
+from matplotlib import pyplot as plt
+from obspy.imaging.beachball import beachball
+from obspy.imaging.beachball import beach
+
+
+eps = 1e-10; halfpi = np.pi/2; twopi = 2*np.pi
+i_hat = np.array([1,0,0]); j_hat = np.array([0,1,0]); k_hat = np.array([0,0,1])
+
+
+def unit_vec(vector: list) -> list:
+    ''' Returns the unit vector of a vector. '''
+    return vector / linalg.norm(vector)
+
+
+def boundary_points(Ao: list, Uo: list, As: list, chi2: float) -> tuple:
+    '''
+    Returns the boundary points of the model.
+
+    Args:
+        Ao (list): Vector of observed amplitudes.
+        Uo (list): Vector of measured uncertainties.
+        As (list, optional): Vector of simulated amplitudes. Defaults to [].
+        chi2 (float, optional): _description_. Defaults to 1.
+    '''
+    
+    # adjusted covariance matrix
+    assert all(Uo > 0), "All uncertainties must be positive."
+    Sigma_tilde = np.diag(Uo)/chi2
+    Sigma_tilde_inv = linalg.inv(Sigma_tilde)
+    
+    # constants
+    c1 = Ao @ Sigma_tilde_inv @ Ao
+    c2 = Ao @ Sigma_tilde_inv @ As
+    c3 = As @ Sigma_tilde_inv @ As
+    
+    # conditions
+    assert c1 >= 1, f"Confidence ellipsoid is too big: c1 = {c1}"
+    assert c1*c3 > c2**2, "Cauchy-Schwarz inequality is not satisfied."
+    
+    # boundary points
+    left = (1 - 1/c1)*Ao
+    right = (1/c1)*np.sqrt((c1-1)/(c1*c3 - c2**2))*(c1*As - c2*Ao)
+    Ab_low, Ab_high = left - right, left + right
+    
+    return Ab_low, Ab_high
+    
+
+def get_epsilon(Ao: list, bounds: tuple) -> float:
+    '''
+    Returns the max angle between the observed and simulated amplitudes.
+    Output is in RADIANS.
+
+    Args:
+        Ao (list): Vector of observed amplitudes.
[…]

### 2025-11-11 suzan/package/model.py

@@ -1,0 +1,648 @@
+'''
+Name: Victor Agaba
+
+Date: 2nd November 2024
+
+The goal of this module is to create a model of the nonlinear inverse
+problem to which we will apply the optimization algorithms.
+'''
+
+
+import numpy as np
+from numpy import linalg
+from . import functions as fn
+from . import optimizer as opt
+
+
+class Model:
+    '''
+    Base class for an unconstrained optimzation problem.
+    Catch anything whose method is not implemented.
+    '''
+    
+    def __call__(self):
+        raise Exception('Class is not callable')
+
+    def gradient(self):
+        raise Exception('Method "gradient" is not implemented')
+
+    def hessian(self):
+        raise Exception('Method "hessian" not is implemented')
+
+    def starting_point(self):
+        raise Exception('Method "starting_point" not implemented')
+
+
+class RadiationModel(Model):
+    '''
+    This class computes the misfit function and its gradient
+    for a given model.
+    Uses the same notation as in the paper.
+    '''
+    def __init__(self, azimuth, takeoff_angles, velocities):
+        '''
+        Initialize with the parts of the model that don't change.
+        '''
+        # initialize parent class
+        super().__init__()
+        
+        # store the original/main hyperparams
+        self.azimuth = azimuth
+        self.takeoff_angles = takeoff_angles
+        
+        # initialize variables for many runs
+        self.super_reset()
+        self.set_hyperparams(azimuth, takeoff_angles, velocities)
+    
+    
+    def set_hyperparams(self, azimuth, takeoff_angles, velocities):
+        '''
+        Set the hyperparameters of the model.
+        All angles are in RADIANS!!!
+        
+        Args:
+        phi = raypath's azimuth
+        i, j = P and S wave takeoff angles
+        alpha, beta = P and S wave velocities
+        '''
+        # extract inputs
+        i, j = takeoff_angles
+        i, j = np.deg2rad(i), np.deg2rad(j)
+        alpha_h, beta_h = velocities
+        
+        # initialize constant Jacobian matrix
+        factors = np.array([alpha_h, beta_h, beta_h])**3
+        self.J_A = np.array(
+            [[(3*np.cos(i)**2 - 1), -np.sin(2*i), -np.sin(i)**2, 0, 0],
+             [1.5*np.sin(2*j), np.cos(2*j), 0.5*np.sin(2*j), 0, 0],
[…]

### 2025-11-11 suzan/package/optimizer.py

@@ -1,0 +1,155 @@
+'''
+Name: Victor Agaba
+
+Date: 4th November 2024
+
+The goal of this module is to design optimization algorithms to be used
+for inversion of seismic data.
+'''
+
+
+import numpy as np
+from numpy import linalg
+from . import functions as fn
+
+
+def get_config():
+    '''
+    Get the default configuration for the optimization algorithm.
+    
+    Output:
+        config (dict): Default configuration for the optimization algorithm.
+        method (str): Optimization method to use: 'SD' or 'Newton'.
+        c_decrease (float): Sufficient decrease factor.
+        c_increase (float): Curvature condition factor.
+        k_max (int): Maximum number of iterations.
+        alpha (float): Initial step size.
+        rho (float): Backtracking line search parameter.
+        tolerance (float): Stopping criterion.
+        print_every (int): Print output every n iterations.
+    '''
+    config = {
+        'method': 'SD',
+        'c_decrease':1e-4,
+        'k_max': 3000,
+        'alpha': 5e-2,
+        'rho': 0.5,
+        'tolerance': 1e-5,
+        'print_every': 0
+    }
+    return config
+
+
+def minimize(objective, config, start=None):
+    '''
+    Minimize an objective function starting from a given point.
+    
+    Args:
+        function (callable): The objective function to minimize.
+        config (dict): Hyperparameters for the optimization.
+        start (list): Starting point for the optimization.
+    
+    Output:
+        x_k (list): The optimal point found by the optimization.
+    '''
+        
+    # extract config parameters
+    method = config['method']
+    c_decrease = config['c_decrease']
+    k_max = config['k_max']
+    alpha_init = config['alpha']
+    rho = config['rho']
+    tolerance = config['tolerance']
+    print_every = config['print_every']
+    
+    # get initial guess
+    if start is None:
+        start = function.get_start()
+    x_k = start
+    
+    # print output header
+    if print_every > 0:
+        if len(x_k) > 5:
+            print(f'{method} optimization with backtracking linesearch from start = {x_k[:5]}...')
+        else:
+            print(f'{method} optimization with backtracking linesearch from start = {x_k}')
+        print(f'iter                f       ||p_k||       alpha     #func     ||grad_f||')
+    
+    # evaluate function at start for later comparison
+    f_k = objective(x_k)
+    H_k = np.eye(len(x_k))
+    
+    # initialize iteration counter
+    k = 1
+    
[…]

### 2025-11-11 suzan/package/plots.py

@@ -1,0 +1,655 @@
+'''
+Name: Victor Agaba
+
+Date: 10th December 2024
+
+The goal of this module is to implement plotting functions for
+the data analysis.
+'''
+
+import numpy as np
+from matplotlib import pyplot as plt
+from obspy.imaging.beachball import beach
+from . import functions as fn
+from . import model as sm
+
+
+def misfit(model: sm.RadiationModel):
+    '''
+    Plot the misfits in a subplot.
+    '''
+    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
+    ax.plot(model.get_misfits())
+    ax.set_title('Misfit function')
+    ax.set_xlabel('Iteration')
+    ax.set_ylabel('Misfit')
+    plt.show()
+    
+
+def half_angles(model: sm.RadiationModel, bins=20):
+    '''
+    Plot the half angles in a histogram.
+    This is a diagnostic plot.
+    '''
+    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
+    ax.hist(np.rad2deg(model.get_half_angles()), bins=bins)
+    ax.set_title('Diagnostic histogram of half-angles')
+    ax.set_xlabel('Half angle (deg)')
+    ax.set_ylabel('Frequency')
+    plt.show()
+    
+    
+def iterates_2D(model: sm.RadiationModel, cmap='rainbow', s=10, optimal=True, index=2):
+    '''
+    Make 3 2D plots of the iterates: psi, delta, lambda.
+    psi against delta, psi against lambda, delta against lambda.
+    Join the points with a line and color them by iteration number.
+    
+    Args:
+        optimal (bool): If True, plot only the optimal points.
+        index (int): 0 is 1st fault plane, 1 is 2nd fault plane, 2 is both.
+    '''
+    fig, (ax1, ax2, ax3) = plt.subplots(1, 3, figsize=(18, 5))
+    
+    if index == 0 or index == 1:
+        if not optimal:
+            model.mirror(['optimals', 'iterates'], index)
+            iterates = model.get_iterates()
+            optimal_iterates = model.get_optimal_iterates()
+        else:
+            model.mirror(['optimals'], index)
+            optimal_iterates = model.get_optimal_iterates()
+    elif index == 2:
+        iterates, optimal_iterates = [], []
+        if not optimal:
+            model.mirror(['optimals', 'iterates'], 0)
+            iterates.extend(model.get_iterates())
[…]

### 2025-11-12 project/package/functions.py
Synthetic stage complete

@@ -756,3 +756,56 @@ def covariance_transform(cov: np.array) -> tuple:
     
     return sd, sr, dr
 
+
+def condition_colors(matrices: list, threshold: float = 0.2) -> list:
+    """
+    Given a list of square or full-rank matrices, return a list of RGBA colors
+    transitioning from blue (low log(cond)) to green (high log(cond)).
+    
+    Args:
+        matrices (list of np.ndarray): List of matrices.
+        
+    Returns:
+        colors (list of tuple): RGBA colors for each matrix.
+    """
+    # compute condition numbers
+    Ks = np.array([np.linalg.cond(A) for A in matrices])
+    
+    # take ln to handle large range
+    logK = np.log(Ks)
+    
+    # normalize logK to 0-1
+    logK_norm = (logK - logK.min()) / (logK.max() - logK.min())
+    
+    # map to colors: linear interpolation between blue (0,0,1) and green (0,1,0)
+    # RGBA tuples: (R,G,B,A)
+    def val_map(val): return max(0, val-threshold)
+    colors = [(val_map(norm), val_map(norm), 1, 1) for norm in logK_norm]
+    
+    return colors
+
+
+def extract_velocities(lookup_table, depth):
+    '''
+    Return p and s wave velocities at a given depth
+    '''
+    depths = lookup_table.iloc[:,0].values
+    vp_values = lookup_table.iloc[:,1].values
+    vs_values = lookup_table.iloc[:,2].values
+
+    if depth <= depths.min():
+        return vp_values[0], vs_values[0]
+    elif depth >= depths.max():
+        return vp_values[-1], vs_values[-1]
+    else:
+        idx = np.searchsorted(depths, depth)
+        d0, d1 = depths[idx-1], depths[idx]
+        vp0, vp1 = vp_values[idx-1], vp_values[idx]
+        vs0, vs1 = vs_values[idx-1], vs_values[idx]
+        
+        # interpolate
+        vp = vp0 + (vp1 - vp0) * (depth - d0) / (d1 - d0)
+        vs = vs0 + (vs1 - vs0) * (depth - d0) / (d1 - d0)
+
+        return np.array([vp, vs])
+    

### 2025-11-12 project/package/model.py

@@ -753,7 +753,7 @@ class RadiationModel(Model):
         self.sampled_weights = []
 
 
-    def set_Uo(self, Uo = [], scale=5e-3, sphere=False):
+    def set_Uo(self, Uo = [], scale=1e-1, sphere=False):
         ''' Set the observed uncertainties. '''
         if len(Uo) == 0:
             if sphere:

### 2025-11-12 project/package/plots.py

@@ -188,6 +188,11 @@ def iterates_2D(model: sm.RadiationModel, cmap='rainbow', s=10, optimal=True,
     ax1.set_xlabel('Strike (deg)')
     ax1.set_ylabel('Dip (deg)')
     
+    if uncertainty:
+        plt.suptitle(f'Optimal fault parameters with {conf*100:.0f}% confidence ellipses', fontsize=16)
+    else:
+        plt.suptitle('Optimal fault parameters', fontsize=16)
+    
     if not optimal:
         ax2.scatter(strikes, rakes, c=weights, cmap=cmap, norm=norm, s=s)
         ax2.scatter(opt_strikes, opt_rakes, c='black', marker='.', s=s, label='Optimal')
@@ -366,7 +371,10 @@ def amplitudes(model: sm.RadiationModel, elev=30, azim=45, cmap='rainbow', s=50,
     ax.set_ylabel('ASV')
     ax.set_zlabel('ASH')
     ax.legend()
-    plt.title('Optimal amplitudes', fontsize=15)
+    if uncertainty:
+        plt.title(f'Optimal amplitudes with {conf*100:.0f}% confidence ellipsoid', fontsize=15)
+    else:
+        plt.title('Optimal amplitudes', fontsize=15)
     
     if iterates:
         cbar = fig.colorbar(it_scatter, ax=ax, pad=0.1, shrink=0.6)
@@ -625,7 +633,7 @@ def beachballs(model: sm.RadiationModel, order_by='strike',width=10, max_plot=50
         if original: solution_set = [Ao_comp_mt] + solution_set
     else:
         og_set = model.get_optimal_iterates()
-        solution_set = model.get_optimal_iterates()
+        solution_set = model.get_optimal_iterates() # optimal_covariances
         
         assert len(solution_set) == len(og_set), f'Length mismatch'
         # get sorting order from og_set
@@ -641,6 +649,7 @@ def beachballs(model: sm.RadiationModel, order_by='strike',width=10, max_plot=50
     if overlap:
         fig, ax = plt.subplots(subplot_kw={'aspect': 'equal'}, figsize=figsize)
         ax.axison = False
+        solution_set = [np.rad2deg(s) for s in solution_set] if order_by != 'tensor' else solution_set
         collections = overlap_beach(solution_set)
         for collection in collections:
             ax.add_collection(collection)
@@ -651,8 +660,9 @@ def beachballs(model: sm.RadiationModel, order_by='strike',width=10, max_plot=50
         covariances = []
         if uncertainty:
[…]

### 2025-11-12 suzan/package/__init__.py

@@ -1,1 +1,0 @@
-__all__ = ['functions', 'model', 'optimizer', 'plots']

### 2025-11-12 suzan/package/functions.py

@@ -1,549 +1,0 @@
-'''
-Name: Victor Agaba
-
-Date: 4th November 2024
-
-The goal of this module is to provide all helper functions needed for the
-seismic inversion project.
-
-'''
-
-
-import numpy as np
-from numpy import linalg
-import pandas as pd
-from matplotlib import pyplot as plt
-from obspy.imaging.beachball import beachball
-from obspy.imaging.beachball import beach
-
-
-eps = 1e-10; halfpi = np.pi/2; twopi = 2*np.pi
-i_hat = np.array([1,0,0]); j_hat = np.array([0,1,0]); k_hat = np.array([0,0,1])
-
-
-def unit_vec(vector: list) -> list:
-    ''' Returns the unit vector of a vector. '''
-    return vector / linalg.norm(vector)
-
-
-def boundary_points(Ao: list, Uo: list, As: list, chi2: float) -> tuple:
-    '''
-    Returns the boundary points of the model.
-
-    Args:
-        Ao (list): Vector of observed amplitudes.
-        Uo (list): Vector of measured uncertainties.
-        As (list, optional): Vector of simulated amplitudes. Defaults to [].
-        chi2 (float, optional): _description_. Defaults to 1.
-    '''
-    
-    # adjusted covariance matrix
-    assert all(Uo > 0), "All uncertainties must be positive."
-    Sigma_tilde = np.diag(Uo)/chi2
-    Sigma_tilde_inv = linalg.inv(Sigma_tilde)
-    
-    # constants
-    c1 = Ao @ Sigma_tilde_inv @ Ao
-    c2 = Ao @ Sigma_tilde_inv @ As
-    c3 = As @ Sigma_tilde_inv @ As
-    
-    # conditions
-    assert c1 >= 1, f"Confidence ellipsoid is too big: c1 = {c1}"
-    assert c1*c3 > c2**2, "Cauchy-Schwarz inequality is not satisfied."
-    
-    # boundary points
-    left = (1 - 1/c1)*Ao
-    right = (1/c1)*np.sqrt((c1-1)/(c1*c3 - c2**2))*(c1*As - c2*Ao)
-    Ab_low, Ab_high = left - right, left + right
-    
-    return Ab_low, Ab_high
-    
-
-def get_epsilon(Ao: list, bounds: tuple) -> float:
-    '''
-    Returns the max angle between the observed and simulated amplitudes.
-    Output is in RADIANS.
-
-    Args:
-        Ao (list): Vector of observed amplitudes.
[…]

### 2025-11-12 suzan/package/model.py

@@ -1,648 +1,0 @@
-'''
-Name: Victor Agaba
-
-Date: 2nd November 2024
-
-The goal of this module is to create a model of the nonlinear inverse
-problem to which we will apply the optimization algorithms.
-'''
-
-
-import numpy as np
-from numpy import linalg
-from . import functions as fn
-from . import optimizer as opt
-
-
-class Model:
-    '''
-    Base class for an unconstrained optimzation problem.
-    Catch anything whose method is not implemented.
-    '''
-    
-    def __call__(self):
-        raise Exception('Class is not callable')
-
-    def gradient(self):
-        raise Exception('Method "gradient" is not implemented')
-
-    def hessian(self):
-        raise Exception('Method "hessian" not is implemented')
-
-    def starting_point(self):
-        raise Exception('Method "starting_point" not implemented')
-
-
-class RadiationModel(Model):
-    '''
-    This class computes the misfit function and its gradient
-    for a given model.
-    Uses the same notation as in the paper.
-    '''
-    def __init__(self, azimuth, takeoff_angles, velocities):
-        '''
-        Initialize with the parts of the model that don't change.
-        '''
-        # initialize parent class
-        super().__init__()
-        
-        # store the original/main hyperparams
-        self.azimuth = azimuth
-        self.takeoff_angles = takeoff_angles
-        
-        # initialize variables for many runs
-        self.super_reset()
-        self.set_hyperparams(azimuth, takeoff_angles, velocities)
-    
-    
-    def set_hyperparams(self, azimuth, takeoff_angles, velocities):
-        '''
-        Set the hyperparameters of the model.
-        All angles are in RADIANS!!!
-        
-        Args:
-        phi = raypath's azimuth
-        i, j = P and S wave takeoff angles
-        alpha, beta = P and S wave velocities
-        '''
-        # extract inputs
-        i, j = takeoff_angles
-        i, j = np.deg2rad(i), np.deg2rad(j)
-        alpha_h, beta_h = velocities
-        
-        # initialize constant Jacobian matrix
-        factors = np.array([alpha_h, beta_h, beta_h])**3
-        self.J_A = np.array(
-            [[(3*np.cos(i)**2 - 1), -np.sin(2*i), -np.sin(i)**2, 0, 0],
-             [1.5*np.sin(2*j), np.cos(2*j), 0.5*np.sin(2*j), 0, 0],
[…]

### 2025-11-12 suzan/package/optimizer.py

@@ -1,155 +1,0 @@
-'''
-Name: Victor Agaba
-
-Date: 4th November 2024
-
-The goal of this module is to design optimization algorithms to be used
-for inversion of seismic data.
-'''
-
-
-import numpy as np
-from numpy import linalg
-from . import functions as fn
-
-
-def get_config():
-    '''
-    Get the default configuration for the optimization algorithm.
-    
-    Output:
-        config (dict): Default configuration for the optimization algorithm.
-        method (str): Optimization method to use: 'SD' or 'Newton'.
-        c_decrease (float): Sufficient decrease factor.
-        c_increase (float): Curvature condition factor.
-        k_max (int): Maximum number of iterations.
-        alpha (float): Initial step size.
-        rho (float): Backtracking line search parameter.
-        tolerance (float): Stopping criterion.
-        print_every (int): Print output every n iterations.
-    '''
-    config = {
-        'method': 'SD',
-        'c_decrease':1e-4,
-        'k_max': 3000,
-        'alpha': 5e-2,
-        'rho': 0.5,
-        'tolerance': 1e-5,
-        'print_every': 0
-    }
-    return config
-
-
-def minimize(objective, config, start=None):
-    '''
-    Minimize an objective function starting from a given point.
-    
-    Args:
-        function (callable): The objective function to minimize.
-        config (dict): Hyperparameters for the optimization.
-        start (list): Starting point for the optimization.
-    
-    Output:
-        x_k (list): The optimal point found by the optimization.
-    '''
-        
-    # extract config parameters
-    method = config['method']
-    c_decrease = config['c_decrease']
-    k_max = config['k_max']
-    alpha_init = config['alpha']
-    rho = config['rho']
-    tolerance = config['tolerance']
-    print_every = config['print_every']
-    
-    # get initial guess
-    if start is None:
-        start = function.get_start()
-    x_k = start
-    
-    # print output header
-    if print_every > 0:
-        if len(x_k) > 5:
-            print(f'{method} optimization with backtracking linesearch from start = {x_k[:5]}...')
-        else:
-            print(f'{method} optimization with backtracking linesearch from start = {x_k}')
-        print(f'iter                f       ||p_k||       alpha     #func     ||grad_f||')
-    
-    # evaluate function at start for later comparison
-    f_k = objective(x_k)
-    H_k = np.eye(len(x_k))
-    
-    # initialize iteration counter
-    k = 1
-    
[…]

### 2025-11-12 suzan/package/plots.py

@@ -1,655 +1,0 @@
-'''
-Name: Victor Agaba
-
-Date: 10th December 2024
-
-The goal of this module is to implement plotting functions for
-the data analysis.
-'''
-
-import numpy as np
-from matplotlib import pyplot as plt
-from obspy.imaging.beachball import beach
-from . import functions as fn
-from . import model as sm
-
-
-def misfit(model: sm.RadiationModel):
-    '''
-    Plot the misfits in a subplot.
-    '''
-    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
-    ax.plot(model.get_misfits())
-    ax.set_title('Misfit function')
-    ax.set_xlabel('Iteration')
-    ax.set_ylabel('Misfit')
-    plt.show()
-    
-
-def half_angles(model: sm.RadiationModel, bins=20):
-    '''
-    Plot the half angles in a histogram.
-    This is a diagnostic plot.
-    '''
-    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
-    ax.hist(np.rad2deg(model.get_half_angles()), bins=bins)
-    ax.set_title('Diagnostic histogram of half-angles')
-    ax.set_xlabel('Half angle (deg)')
-    ax.set_ylabel('Frequency')
-    plt.show()
-    
-    
-def iterates_2D(model: sm.RadiationModel, cmap='rainbow', s=10, optimal=True, index=2):
-    '''
-    Make 3 2D plots of the iterates: psi, delta, lambda.
-    psi against delta, psi against lambda, delta against lambda.
-    Join the points with a line and color them by iteration number.
-    
-    Args:
-        optimal (bool): If True, plot only the optimal points.
-        index (int): 0 is 1st fault plane, 1 is 2nd fault plane, 2 is both.
-    '''
-    fig, (ax1, ax2, ax3) = plt.subplots(1, 3, figsize=(18, 5))
-    
-    if index == 0 or index == 1:
-        if not optimal:
-            model.mirror(['optimals', 'iterates'], index)
-            iterates = model.get_iterates()
-            optimal_iterates = model.get_optimal_iterates()
-        else:
-            model.mirror(['optimals'], index)
-            optimal_iterates = model.get_optimal_iterates()
-    elif index == 2:
-        iterates, optimal_iterates = [], []
-        if not optimal:
-            model.mirror(['optimals', 'iterates'], 0)
-            iterates.extend(model.get_iterates())
[…]

### 2025-11-12 synthetic/package/functions.py

@@ -1,0 +1,811 @@
+'''
+Name: Victor Agaba
+
+Date: 4th November 2024
+
+The goal of this module is to provide all helper functions needed for the
+seismic inversion project.
+
+'''
+
+
+import numpy as np
+from numpy import linalg
+import pandas as pd
+from matplotlib import pyplot as plt
+from obspy.imaging.beachball import beachball
+from obspy.imaging.beachball import beach
+
+
+eps = 1e-10; halfpi = np.pi/2; twopi = 2*np.pi
+i_hat = np.array([1,0,0]); j_hat = np.array([0,1,0]); k_hat = np.array([0,0,1])
+
+
+def unit_vec(vector: list) -> list:
+    ''' Returns the unit vector of a vector. '''
+    return vector / linalg.norm(vector)
+
+
+def boundary_points(Ao: list, Uo: list, As: list, chi2: float) -> tuple:
+    '''
+    Returns the boundary points of the model.
+
+    Args:
+        Ao (list): Vector of observed amplitudes.
+        Uo (list): Vector of measured uncertainties.
+        As (list, optional): Vector of simulated amplitudes. Defaults to [].
+        chi2 (float, optional): _description_. Defaults to 1.
+    '''
+    
+    # adjusted covariance matrix
+    assert all(Uo > 0), "All uncertainties must be positive."
+    Sigma_tilde = np.diag(Uo)/chi2
+    Sigma_tilde_inv = linalg.inv(Sigma_tilde)
+    
+    # constants
+    c1 = Ao @ Sigma_tilde_inv @ Ao
+    c2 = Ao @ Sigma_tilde_inv @ As
+    c3 = As @ Sigma_tilde_inv @ As
+    
+    # conditions
+    assert c1 >= 1, f"Confidence ellipsoid is too big: c1 = {c1}"
+    assert c1*c3 > c2**2, "Cauchy-Schwarz inequality is not satisfied."
+    
+    # boundary points
+    left = (1 - 1/c1)*Ao
+    right = (1/c1)*np.sqrt((c1-1)/(c1*c3 - c2**2))*(c1*As - c2*Ao)
+    Ab_low, Ab_high = left - right, left + right
+    
+    return Ab_low, Ab_high
+    
+
+def get_epsilon(Ao: list, bounds: tuple) -> float:
+    '''
+    Returns the max angle between the observed and simulated amplitudes.
+    Output is in RADIANS.
+
+    Args:
[…]

### 2025-11-12 synthetic/package/model.py

@@ -1,0 +1,920 @@
+'''
+Name: Victor Agaba
+
+Date: 2nd November 2024
+
+The goal of this module is to create a model of the nonlinear inverse
+problem to which we will apply the optimization algorithms.
+'''
+
+
+import numpy as np
+from numpy import linalg
+from . import functions as fn
+from . import optimizer as opt
+
+
+class Model:
+    '''
+    Base class for an unconstrained optimzation problem.
+    Catch anything whose method is not implemented.
+    '''
+    
+    def __call__(self):
+        raise Exception('Class is not callable')
+
+    def gradient(self):
+        raise Exception('Method "gradient" is not implemented')
+
+    def hessian(self):
+        raise Exception('Method "hessian" not is implemented')
+
+    def starting_point(self):
+        raise Exception('Method "starting_point" not implemented')
+
+
+class RadiationModel(Model):
+    '''
+    This class computes the misfit function and its gradient
+    for a given model.
+    Uses the same notation as in the paper.
+    '''
+    def __init__(self, azimuth, takeoff_angles, velocities):
+        '''
+        Initialize with the parts of the model that don't change.
+        '''
+        # initialize parent class
+        super().__init__()
+        
+        # store the original/main hyperparams
+        self.phi = azimuth
+        self.takeoff_angles = takeoff_angles
+        self.hidden = False
+        
+        # initialize variables for many runs
+        self.super_reset()
+        self.set_hyperparams(azimuth, takeoff_angles, velocities)
+    
+    
+    def set_hidden_params(self, hidden=True):
+        '''
+        Use the hidden parameterization for optimization.
+        '''
+        self.hidden = hidden
+    
+    
+    def is_hidden(self):
+        '''
+        Return whether the model is in hidden parameterization.
+        '''
+        return self.hidden
+    
+    
+    def set_hyperparams(self, azimuth, takeoff_angles, velocities):
+        '''
+        Set the hyperparameters of the model.
+        All angles are in RADIANS!!!
+        
+        Args:
+        phi = raypath's azimuth
+        i, j = P and S wave takeoff angles
+        alpha, beta = P and S wave velocities
+        '''
+        # extract inputs
+        i, j = takeoff_angles
+        i, j = np.deg2rad(i), np.deg2rad(j)
[…]

### 2025-11-12 synthetic/package/plots.py

@@ -1,0 +1,983 @@
+'''
+Name: Victor Agaba
+
+Date: 10th December 2024
+
+The goal of this module is to implement plotting functions for
+the data analysis.
+'''
+
+import numpy as np
+from matplotlib import pyplot as plt
+from obspy.imaging.beachball import beach
+from . import functions as fn
+from . import model as sm
+from matplotlib.patches import Ellipse
+from scipy.stats import chi2
+from mpl_toolkits.mplot3d import Axes3D  # noqa: F401
+from mpl_toolkits.mplot3d import art3d
+
+
+def misfit(model: sm.RadiationModel):
+    '''
+    Plot the misfits in a subplot.
+    '''
+    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
+    ax.plot(model.get_misfits())
+    ax.set_title('Misfit function')
+    ax.set_xlabel('Iteration')
+    ax.set_ylabel('Misfit')
+    plt.show()
+    
+
+def half_angles(model: sm.RadiationModel, bins=20):
+    '''
+    Plot the half angles in a histogram.
+    This is a diagnostic plot.
+    '''
+    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
+    ax.hist(np.rad2deg(model.get_half_angles()), bins=bins)
+    ax.set_title('Diagnostic histogram of half-angles')
+    ax.set_xlabel('Half angle (deg)')
+    ax.set_ylabel('Frequency')
+    plt.show()
+    
+
+def cov_ellipse(ax, x, y, cov_xy, conf=0.95, **kwargs):
+    """
+    Plot a covariance ellipse on a given Matplotlib Axes.
+
+    Parameters
+    ----------
+    ax : matplotlib.axes.Axes
+        The axes to plot the ellipse on.
+    x, y : float
+        Mean coordinates (center of the ellipse).
+    cov_xy : (2, 2) array-like
+        Covariance matrix for (x, y).
+    conf : float, optional (default=0.95)
+        Confidence level for the ellipse (e.g., 0.68, 0.95, 0.99).
+    **kwargs :
+        Additional keyword arguments passed to matplotlib.patches.Ellipse,
+        e.g. edgecolor, facecolor, lw, alpha, linestyle.
+
+    Returns
+    -------
+    ellipse : matplotlib.patches.Ellipse
+        The created ellipse patch (already added to the axes).
+    """
+    # cov_xy = np.asarray(cov_xy)
[…]
